\section{Open and Closed Sets} 

  The first thing to define is a topology. 

  \begin{definition}[Topology]
    Let $X$ be a set and $\T$ be a family of subsets of $X$. Then $\T$ is a \textbf{topology} on $X$\footnote{I will use script letters to denote topologies and capital letters to denote sets.} if it satisfies the following properties. 
    \begin{enumerate}
      \item \textit{Contains Empty and Whole Set}: 
      \begin{equation}
        \emptyset, X \in \T
      \end{equation}

      \item \textit{Closure Under Union}. If $\{U_\alpha\}_{\alpha \in A}$ is a class of sets in $\T$, then 
      \begin{equation}
        \bigcup_{\alpha \in A} U_\alpha \in \T
      \end{equation}

      \item \textit{Closure Under Finite Intersection}: If $U_1, \ldots, U_n$ is a finite class\footnote{Note that we restrict property 3 to be a \textit{finite} intersection because it turns out that the finiteness of intersection allows us to prove many nice properties about topologies, which we will mention later. Another reason is that if we remove this finite restriction, the open ball topology on $\mathbb{R}$ would imply that $\cap_{i = 1}^{\infty} ( - 1/i, +1/i ) = 0$ is an open set $\implies$ all points are open sets too, which is generally not what we want in analysis. 
      } of sets in $\T$, then 
      \begin{equation}
       \bigcap_{i=1}^{n} U_i \in \T
      \end{equation}
    \end{enumerate}
    A \textbf{topological space} is denoted $(X, \T)$. 
  \end{definition}

  This leads to the most general definition of an open set. Note that an open set doesn't really mean anything without talking about with respect to its topology. 

  \begin{definition}[Open Set]
    The elements of $\T$ are called \textbf{open sets} in $X$.\footnote{As implied from the definition of a topology, the arbitrary union and finite intersection of any number of open sets is an open set.} 
    \begin{enumerate}
      \item An open set $U$ which contains a point $x$ is called an \textbf{open neighborhood} of $x$, denoted $U_x$. 
      \item Given an open neighborhood $U_x$ of $x$, the set $U_x \setminus \{x\}$ is called the \textbf{punctured open neighborhood} of $x$. 
    \end{enumerate}
  \end{definition}

  For the sake of giving at least one nontrivial example, here is an example of a finite topology. 

  \begin{example}[Topologies of a Set of Cardinality 3]
    There are a total of 29 topologies that we can construct on $\{1, 2, 3\}$. Two such examples are 
    \begin{enumerate}
      \item $\{\emptyset, \{1, 2\}, \{1, 2, 3\}\}$ 
      \item $\{\emptyset, \{3\}, \{2, 3\}, \{1, 2, 3\}\}$
    \end{enumerate}
  \end{example} 

  When we define a new topology, we must first prove that they are topologies, and so these definitions are really theorems. However, I will introduce them as definitions and reserve the theorem environment for actual theorems. 

  \begin{definition}[Discrete, Indiscrete Topologies]
    Given a set $X$, 
    \begin{enumerate}
      \item $2^X$ is a topology, called the \textbf{discrete topology}. 
      \item $\{\emptyset, X \}$ is a topology, called the \textbf{indiscrete topology}. 
    \end{enumerate}
  \end{definition}
  \begin{proof}
    Listed. 
    \begin{enumerate}
      \item The first property is trivially proven. From the theorems of set theory, $U_\alpha \subset X \implies \cup U_\alpha \subset X \implies \cup U_\alpha \in 2^X$. Finally the same logic holds for intersection as well. 
      \item The first property is trivially proven. We can check for the 4 combinations of unions and intersections and see that they all result in either $\emptyset$ or $X$. 
    \end{enumerate}
  \end{proof}

  \begin{definition}[Finer, Coarser Topologies]
    Suppose that $\T$ and $\T^\prime$ are two topologies on a given set $X$. If $\T \subset \T^\prime$, we say that $\T^\prime$ is \textbf{finer} than $\T$, or equivalently, we say that $\T$ is \textbf{coarser} than $\T^\prime$. 
  \end{definition}

  We can think of the topology of a set $X$ as a truck full of gravel as the open sets. If the gravel is smashed into smaller, finer pieces, then the amount of stuff that we can make from the finer gravel increases, which corresponds to a bigger topology. Clearly, the indiscrete topology is the coarsest topology and the discrete topology is the finest. 

  \begin{theorem}[Intersection of Topologies]
    Given a family of topologies $\{\T\}_{\alpha \in A}$, the set 
    \begin{equation}
      \T = \bigcap_{\alpha \in A} \T_\alpha
    \end{equation}
    is a topology. 
  \end{theorem}

  \begin{corollary}[Unique Coarsest and Finest Topology]
    Given a family of topologies $\{\T\}_{\alpha \in A}$, there exists 
    \begin{enumerate}
      \item a unique smallest topology on $X$ containing all the collections $\T_\alpha$. 
      \item a unique largest topology on $X$ contained in each $\T_\alpha$. 
    \end{enumerate}
  \end{corollary} 

  \begin{example}
    Let $X = \{a, b, c\}$, and let 
    \begin{align}
      \T_1 & = \{\emptyset, X, \{a\}, \{a, b\}\} \\
      \T_2 & = \{\emptyset, X, \{a\}, \{b, c\}\}
    \end{align}
    We claim that the 
    \begin{enumerate}
      \item smallest topology containing $\T_1, \T_2$ is 
      \begin{equation}
        \T_{1 \cup 2} = \{\emptyset, X \{a\}, \{b\}, \{a, b\}, \{b, c\}\}
      \end{equation} 
      Note that this is not simply the union of topologies. The union wouldn't have $\{b\}$, making it not a topology. 

      \item largest topology contained in $\T_1, \T_2$ is 
      \begin{equation}
        \T_{1 \cap 2} = \{\emptyset, X, \{a\}\}
      \end{equation}
      Note that this is simply the intersection of the two topologies. 
    \end{enumerate}
  \end{example}

\subsection{Basis and Fineness} 

  So far so good. We want to continue analyzing the properties of a topology, but sometimes working with the entire topology is a bit thorny. There is a tamer representation of a topology, which can also give us the starting point to \textit{construct} topologies. 

  \begin{definition}[Basis]
    If $X$ is a set, a \textbf{basis} on $X$ is a collection $\B$ of subsets of $X$ (called \textbf{basis elements}) such that
    \begin{enumerate}
      \item For each $x \in X$, there is at least one basis element $B \in \B$ containing $x$. That is, the elements of $\B$ covers $X$. 
      \item If $x$ belongs to the intersection of two basis elements $B_1$ and $B_2$, then there is a basis element $B_3$ containing $x$ such that $B_3 \subset (B_1 \cap B_2)$. 
    \end{enumerate}
  \end{definition} 

  The name gives away the clue that a topology may be created from this basis.  

  \begin{theorem}[Basis to Topology]
    Given a basis $\B$ on a set $X$, we can define a topology $\T$, called the \textbf{topology generated by $\B$}, in the following equivalent ways. 
    \begin{enumerate}
      \item $\T$ consists of subsets $U$ of $X$ satisfying the property that for each $x \in U$, there exists a basis element $B \in \B$ such that $x \in B \subset U$.\footnote{Note that since we can always set $U = \emptyset$, the basis doesn't need to contain $\emptyset$. }
      \begin{center}
        \begin{tikzpicture}
        \draw[dashed] plot [smooth cycle] coordinates {(0,0) (1,1) (3,1) (3,0) (1,-1)};
        \node [right] at (0.5,0) {U};
        \draw[fill=lightgray,dashed] (2,0.7) circle [radius=0.35];
        \draw[fill] (2, 0.7) circle [radius=0.05];
        \node [right] at (1.95,0.7) {x};
        \node [right] at (4,-0.5) {B};
        \draw (1.95,0.6)--(4,-0.5);
        \end{tikzpicture}
      \end{center}

      \item $\T$ consists of all possible unions of elements in $\B$. 
      \begin{equation}
        \T \equiv \Big\{ \bigcup_i b_i \; \Big| \; b_i \in \B\Big\}
      \end{equation}
    \end{enumerate}
  \end{theorem} 
  \begin{proof}
    We prove that the 2 methods generate a topology, and then finally prove that it they are the same topology. 
    \begin{enumerate}
      \item Clearly, $\emptyset$ and $X$ itself are in $\T$. To prove property 2, given a certain indexed family of subsets $\{U_\alpha\}_{\alpha \in I}$ of $\T$, we must show that 
      \begin{equation}
        U = \bigcup_{\alpha \in I} U_\alpha \in \T
      \end{equation}
      Given $x \in U$, there exists at least one index $\alpha$ such that $x \in U_\alpha$. Since $U_\alpha \in \T$ already, there exists a basis element $b \in \B$ such that $x \in b \subset U_\alpha$. But 
      \begin{equation}
        U_\alpha \subseteq U \implies b \subset U
      \end{equation}
      So, by definition, any arbitrary union of $U$ of these subsets is also in $\T$. 

      To prove property 3, we must show that 
      \begin{equation}
        W = \bigcap_{\alpha \in I} U_\alpha \in \T
      \end{equation}
      Given $x \in W$, by definition of a basis element, there exists a $b \in \B$ such that 
      \begin{equation}
        x \in b \subset (U_\beta \cap U_\gamma) \forall \beta, \gamma \in I \implies \text{ there exists } \Tilde{b} \in \B \text{ s.t. } x \in \Tilde{b} \subset \bigcap_{\alpha \in I} U_\alpha
      \end{equation}
      By definition, $W$ is also open. Since this arbitrary set of subsets $\T$ suffices the 3 properties, it is a topology of $X$ by definition. 

      \item $(\rightarrow)$ Given a collection of elements in $\B$, they are also elements of $\T$. Since $\T$ is a topology, their union in also in $\T$. 

      $(\leftarrow)$ Given an open set $U \in \T$, for every point $x \in U$, by definition we can choose a basis element $b \in \B$ such that $x \in b \subset U$. Then, the union of all these basis elements is by definition $b$. 
        
    \end{enumerate}
  \end{proof}

  We have learned how to go from a basis to a topology. The following lemma tells us how to identify a basis within a topology. 

  \begin{theorem}[Topology to Basis]
    Let $X$ be a topological space, and let $\mathscr{C}$ be a collection of subsets of $X$ such that for every open set $U$ and each $x \in U$, there exists an element $C \in \mathscr{C}$ such that
    \begin{equation}
      x \in C \subset U
    \end{equation}
    Then, $\mathscr{C}$ is a basis for the topology of $X$. 
  \end{theorem}
  \begin{proof}
    
  \end{proof} 

  Characterizing topologies in terms of basis is quite effective since we can work with more manageable sets. 

  \begin{lemma}[Fineness w.r.t. Basis]
    Given two topologies $\T$ and $\T^\prime$ with their bases $\B$ and $\B^\prime$, respectively, the following are equivalent. 
    \begin{enumerate}
      \item $\T^\prime$ is finer than $\T$. 
      \item For each $x \in X$ and basis element $B \in \B$ containing $x$, there exists a basis element $B^\prime \in \B^\prime$ such that $x \in B^\prime \subset B$. 
    \end{enumerate}
  \end{lemma}
 
\subsection{Limit Points and Closed Sets} 

  First, we need to learn what it generally means for a point to be infinitesimally close to a set. 

  \begin{definition}[Limit Point]
    Given a topological space $(X, \T)$, let $x \in X$ be a point and $S \subset X$ a subset. $x$ is a \textbf{limit point of $S$} if every punctured neighborhood of $x$ intersects $S$.\footnote{Note that limit point are generally used to talk about points that are infinitesimally close to a set $S$. A limit point may not necessarily be in $S$, and a point of $S$ may not necessarily be a limit point. This is why we use a punctured neighborhood, rather than an open neighborhood. For continuity as we will see later, we just talk about neighborhoods since we also claim that the limit exists and the function value is the limit.} The set of all limit points of a set $S$ is denoted $S^\prime$.  
  \end{definition} 

  \begin{example}[Examples of Limit Points]
    What about the limit points that are not in $S$? Generally, there are two instances. 
    \begin{enumerate}
      \item Let $S$ represent the gray area. $B$ is in the ``interior'' of $S$ and therefore is a limit point. $A$ and $C$ are on the ``boundary'' of $S$ yet not in $S$, and we can show that they are limit points as well. 
        
      \begin{figure}[H]
        \centering 
        \begin{tikzpicture}
          \draw[fill=lightgray, dashed] plot [smooth cycle] coordinates {(0,0) (1,1) (3,1) (3,0) (1,-1)};
          \draw [fill] (3, 1) circle [radius=0.05];
          \node [above right] at (3,1) {A};
          \draw [fill] (2,0) circle [radius=0.05];
          \node [above] at (2,0) {B};
          \draw [fill] (0,0) circle [radius=0.05];
          \node [left] at (0,0) {C};
          \draw [fill] (5,0) circle [radius=0.05];
          \node [right] at (5, 0) {D};
        \end{tikzpicture}
        \caption{Points $A, B, C$ are limit points of the open set. }
        \label{fig:limit_boundary}
      \end{figure}

      \item A point can be at the ``convergence point'' of a sequence. 
      \begin{figure}[H]
        \centering 
        \begin{tikzpicture}
          \draw [fill] (0, 2.2) circle [radius=0.05];
          \draw [fill] (1, 2.4) circle [radius=0.05];
          \draw [fill] (2, 2) circle [radius=0.05];
          \draw [fill] (2.5, 1.8) circle [radius=0.05];
          \draw [fill] (2.6, 1.6) circle [radius=0.05];
          \draw [fill] (2.65, 1.67) circle [radius=0.05];
          \draw [fill] (2.654, 1.64) circle [radius=0.05];
          \draw [fill] (2.6543, 1.63) circle [radius=0.05];
          \node [right] at (2.6543, 1.63) {p};
        \end{tikzpicture}
        \caption{Note that if $S$ is a sequence of points in $\mathbb{R}^{2}$ that converges to $p$ without ever hitting it, we can say that $p \not\in S$ is a limit point of $S$.}
        \label{fig:limit_sequence}
      \end{figure}
    \end{enumerate}
  \end{example}

  \begin{example}[Examples of Non-Limit Points]
    There are generally two instances of non-limit points. Let $X = \mathbb{R}$ and $S = (0, 1) \cup \{2\}$. 
    \begin{enumerate}
      \item $5$ is clearly not a limit point. 
      \item $2$, although in $S$, is not a limit point since we are talking about the punctured neighborhood. A point in $S$ that is not a limit point is called an \textbf{isolated point}. 
    \end{enumerate}
  \end{example}

  \begin{definition}[Closed Set]
    A set $S \subset X$ is \textbf{closed} if its complement $X \setminus S$ is open in $\T$.\footnote{Note that open and closed sets are not mutually exclusive. A set might be open, closed, both, or neither. A set that is both open and closed is called \textbf{clopen}.}
  \end{definition}
  
  Another property, which is often used as the definition of a closed set, is that it contains all of its limit points. 

  \begin{lemma}[Closed Sets Contain Limit Points]
    A set $S \subset X$ is closed iff it contains all of its limit points. 
  \end{lemma}
  \begin{proof}
    We prove bidirectionally. 
  \end{proof}

  \begin{theorem}[Topological Space wrt Closed Sets]
    Let $X$ be a topological space. Then, the following conditions hold
    \begin{enumerate}
      \item $\emptyset$ and $X$ are clopen.
      \item Arbitrary intersections of closed sets are closed. 
      \item Finite unions of closed sets are closed. 
    \end{enumerate}
  \end{theorem}

  \begin{definition}[Dense Subsets]
    Let $S \subset (X, \tau_X)$. $S$ is \textbf{dense} in $X$ if every point $p \in X$ is a limit point of $S$. In other words, for any point $p \in X$ and any open neighborhood $U_p$ of $p$, $U_p \cap S$ is nontrivial. Otherwise, $p$ is a point of $S$. 
  \end{definition}

  The following example is a crucial fact for proving further properties of topological spaces. 

  \begin{example}
    $\mathbb{Q}^{n}$ is a dense set of $\mathbb{R}^{n}$ with the open ball topology. If we have the discrete topology of $\mathbb{R}^{2}$, an open neighborhood of a point is the point itself, so no limit points would exist beyond the points in $S$ itself. So $\mathbb{Q}^{n}$ is not dense in $\mathbb{R}^{n}$ with this topology. 
  \end{example}

\subsection{Interiors and Closures}

  Now that we've determined limit points, we would like to extend sets into their limit points. The process of doing this is called the \textit{closure} of a set. 

  \begin{definition}[Closure]
    The \textbf{closure} of set $S$ is $\overline{S}$ is defined in the following equivalent ways. 
    \begin{enumerate}
      \item $\overline{S} = S \cup S^\prime$, i.e. the union of itself and its limit points. 
      \item $\overline{S}$ is the intersection of all closed sets $C$ containing $S$. 
    \end{enumerate}
  \end{definition}
  \begin{proof}
    
  \end{proof}

  \begin{example}
    If $S$ is an open ball, $\Bar{S}$ is the closed ball. 
  \end{example}

  From semantics, it may seem like the interior and exterior (defined later) are related, but from a mathematical point of view, the interior and closure are dual notions. 

  \begin{definition}[Interior]
    Let $S \subset X$. Then, the following definitions of the \textbf{interior} of $S$, denoted $S^\circ$, are equivalent. 
    \begin{enumerate}
      \item $x \in S^\circ$ if $\exists U_x \ni x$ s.t. $U_x \subset S$. 
      \item $S^\circ$ is the union of all open sets contained in $S$. 
      \item $S^{o}$ is the complement of the closure of the complement of S. 
      \begin{equation}
        S^{o} = \big(\overline{S^{c}}\big)^{c}
      \end{equation}
    \end{enumerate}
  \end{definition}
  \begin{proof}
    
  \end{proof}

  \begin{lemma}[Open and Closed in Terms of Interiors and Closures]
    Let $S$ be a subset of some topological space $X$. 
    \begin{enumerate}
      \item $S$ is open iff $S = S^{o}$. $S^{o}$ is always open.
      \item $S$ is closed iff $S = \overline{S}$. $\overline{S}$ is always closed. 
    \end{enumerate}
  \end{lemma}

  \begin{theorem}[Clopen sets in Reals]
    There are no proper clopen sets in $\mathbb{R}$. 
  \end{theorem}

\subsection{Exteriors and Boundaries}

  \begin{definition}[Exteriors]
    Let $S \subset X$. The \textbf{exterior} of $S$, denoted $S^e$, is defined in the following equivalent ways.\footnote{We can informally think of the exterior being strictly outside of $S$ and its boundary.}
    \begin{enumerate}
      \item $S^e$ is the complement of the closure of $S$. 
      \item $S^e$ is the interior of the complement of $S$. 
    \end{enumerate}
  \end{definition}
  \begin{proof}
    
  \end{proof}

  \begin{definition}[Boundary]
    Let $S \subset X$. The \textbf{boundary} of $S$, denoted $\partial S$, is defined in the following equivalent ways. 
    \begin{enumerate}
      \item $\partial S$ is the closure minus the interior of $S$ in $X$. 
      \item $\partial S$ is the intersection of the closure of $S$ with the closure of its complement, i.e the set of all points $x$ such that every neighborhood $U_x$ intersects both the interior and exterior. 
      \item $\partial S$ is the set of points that are neither in the exterior nor the interior. 
      \item $x \in \partial S$ if every neighborhood of $x$ intersects both the interior and exterior of $S$. 
    \end{enumerate}
  \end{definition}
  \begin{proof}
    
  \end{proof}

  From the above, we get the intuitive notion that these three parts divide up the whole space. 

  \begin{theorem}[Partitioning of Space]
    Given $S \subset X$, $X$ is partitioned into the interior, boundary, and exterior of $S$. 
    \begin{equation}
      X = S^\circ \sqcup \partial S \sqcup S^e
    \end{equation}
  \end{theorem}
  \begin{proof}
    The fact that 
  \end{proof}

  One counterintuitive result is the \href{https://en.wikipedia.org/wiki/Lakes_of_Wada}{Lakes of Wada}, which are three disjoint connected open sets of the open unit square $(0, 1)^2$ with the property that they \textit{all} have the same boundary. In other words, for any point selected on the boundary of one of the lakes, the other two lakes' boundaries also contain that point. 

