\section{The Real Numbers}

  By constructing $\mathbb{Q}$ and its topology in my algebra and topology notes, we can talk about convergence. The first question to ask (if you were the first person inventing the reals) is ``how do I know that there exists some other numbers at all?'' The first clue is trying to find the side length of a square with area $2$. As we see, this number is not rational. 

  \begin{theorem}[$\sqrt{2}$ is Not Rational]
    \label{thm:sqrt2-irrational}
    There exists no $x \in \mathbb{Q}$ s.t. $x^2 = 2$. 
  \end{theorem}
  \begin{proof}
    Assume such a number $x = p/q$ exists, where $\mathrm{gcd}(p, q) = 1$. Then, by the field axioms of $\mathbb{Q}$, we can deduce that 
    \begin{equation}
      2 = \frac{p^2}{q^2} \implies 2 q^2 = p^2
    \end{equation}
    This implies that $2 \mid p$, so we have $p = 2p_0$, and we can write $2 q^2 = 4 p_0^2$. Dividing both sides by $2$, we get $q^2 = 2p_0^2$, which implies that $2 \mid q$. This contradicts the fact that $p$ and $q$ are coprime. 
  \end{proof} 

  We can ``imagine'' that a square with area $2$ certainly exists, but the fact that its side length is undefined is certainly unsettling. I don't know about you, but I would try to ``invent'' $\sqrt{2}$. We can do this in 4 distinct ways, though some may be more similar than others. 

  \begin{enumerate}
    \item \textit{Dedekind Completeness}. I would try and generate new elements by taking a specific \textit{cut}---a partition into two sets such that elements of one set is always greater than that of the other---and seeing which elements lie right in between these cuts. We will often see that we can always find a cut for every rational, but there are additional cuts for each irrational number. This is the idea behind \textit{Dedekind completeness}. 

    \item \textit{Cauchy Completeness}. I write out the decimal expansion one by one, which gives our first exposure to sequences. 
    \begin{equation}
      1, 1.4, 1.41, 1.414, \ldots
    \end{equation} 
    It is clear that on $\mathbb{Q}$, this sequence does not converge. Our intuition tells that that if the terms get closer and closer to each other, they must be getting closer and closer to \textit{something}, though that something is not in $\mathbb{Q}$. This motivates the definition for \textit{Cauchy completeness}. 

    \item \textit{Nested Interval Completeness}. I would write out maybe some nested intervals so that $\sqrt{2}$ \textit{must} lie within each interval. 
    \begin{equation}
      [1, 2] \supset [1.4, 1.5] \supset [1.41, 1.42] \supset \ldots 
    \end{equation}
    This motivates the definition of \textit{nested-interval completeness}. 

    \item \textit{Least Upper Bound Completeness}. I would define the set of all rationals such that $x^2 < 2$, and try to define $\sqrt{2}$ as the max or supremum of this set. We will quickly find that neither the max nor the supremum exists in $\mathbb{Q}$, and this motivates the definition for \textit{least upper bound completeness}. 
  \end{enumerate}

  All four of these methods points at the same intuition that there should not be any ``gaps'' or ``missing points'' in the set that we will construct to be $\mathbb{R}$, which is the general notion of \textit{completeness}. This contrasts with the rational numbers, whose corresponding number line has a "gap" at each irrational value. Even though constructing the reals with one method is sufficient, knowing the different flavors in which completeness manifests is very useful. It allows us to view properties of $\mathbb{R}$ through different lens. 

  The main division between these four properties is that the first two are methods to directly \textit{construct} the reals from $\mathbb{Q}$, while the latter two are more \textit{axiomatic properties} that we use to verify completeness for a given set. In the next two sections, we will take the rationals and add on extra elements using Dedekind cuts and Cauchy sequences. However, it isn't as conventional (though possible) to construct them as the single point contained in a sequence of nested intervals nor as a supremum of all upper-bounded sets. In fact, an alternative way to construct the reals is to define it axiomatically---as a totally ordered field with either the LUB property or the nested interval property.\footnote{In fact, you also need the Archimidean principle, but we'll talk about this later.} 

  Therefore, in the following sections, we will 
  \begin{enumerate}
    \item first define the relevant notion of completeness, 
    \item show that the rationals are not complete 
    \item and then construct the completed version of the rationals as a version of the reals. 
  \end{enumerate}
  Once we have done this for all three versions, we will unify them by proving they are all equivalent. 

  There is a second essential property of the reals that is not talked about as often is the Archimidean principle. 

  \begin{definition}[Archimidean Principle]
    An ordered ring $(X, +, \cdot, \leq)$ that embeds the naturals $\mathbb{N}$\footnote{as in, there exists an ordered ring homomorphism $\iota: \mathbb{N} \rightarrow X$} is said to obey the \textbf{Archimedean principle} if given any $x, y \in X$ s.t. $x, y > 0$, there exists an $n \in \mathbb{N}$ s.t. $\iota(n) \cdot x > y$. Usually, we don't care about the canonical injection and write $nx > y$. 
  \end{definition} 
  
  \begin{lemma}[Rationals are Archimidean]
    $\mathbb{Q}$ satisfies the Archimidean principle. 
  \end{lemma}
  \begin{proof}
    Take any $x = p_1/q_1, y = p_2 / q_2 \in \mathbb{Q}$. Then, take $n = q_1 p_2$, and we get 
    \begin{equation}
      n x = q_1 p_2 \frac{p_1}{q_1} = p_1 p_2 > \frac{p_2}{q_2} = y \iff p_1 > \frac{1}{q_2}
    \end{equation}
    which must be true since $p_1 \geq 1 \geq \frac{1}{q_2}$. 
  \end{proof}

  Usually, we just construct $\mathbb{R}$ right out of $\mathbb{Q}$, and it turns out that the Archimidean principle just trivially follows. However, if we construct $\mathbb{R}$ axiomatically (without the rationals), it needs to be stated. In this axiomatic formulation, we will find that certain types of completeness---like Dedekind completeness---actually \textit{implies} the Archimidean principle, while others---like Cauchy and nested-intervals completeness---does not. Therefore, in a sense, Dedekind completeness is the ``strongest'' form of completeness. 

\subsection{Dedekind Completeness}  

  This is an explicit construction from the rationals. 

  \begin{definition}[Dedekind Cut] 
    A \textbf{Dedekind cut} is a partition of the rationals $\mathbb{Q} = A \sqcup A^\prime$ satisfying the three properties.\footnote{This can really be defined for any totally ordered set. } 
    \begin{enumerate}
      \item $A \neq \emptyset$ and $A \neq \mathbb{Q}$.\footnote{By relaxing this property, we can actually complete $\mathbb{Q}$ to the extended real number line. }
      \item $x < y$ for all $x \in A, y \in A^\prime$. 
      \item The maximum element of $A$ does not exist in $\mathbb{Q}$. 
    \end{enumerate}
    The minimum of $A^\prime$ may exist in $\mathbb{Q}$, and if it does, the cut is said to be \textbf{generated} by $\min A^\prime$. 
  \end{definition}

  Note that in $\mathbb{Q}$, there will be two types of cuts: 
  \begin{enumerate}
    \item ones that are generated by rational numbers, such as 
    \begin{equation}
      A = \{x \in \mathbb{Q} \mid x < 2/3 \}, A^\prime = \{ x \in \mathbb{Q} \mid x \geq 2/3 \} 
    \end{equation}
    \item and the ones that are not 
    \begin{equation}
      A = \{x \in \mathbb{Q} \mid x^2 < 2 \}, A^\prime = \{x \in \mathbb{Q} \mid x^2 \geq 2 \}
    \end{equation}
  \end{enumerate}
  We can intuitively see that the set of all Dedekind cuts $(A, A^\prime)$ will ``extend'' the rationals into a bigger set. We can then define some operations and an order to construct this into an ordered field, and finally it will have the property that we call ``completeness.''

  \begin{definition}[Dedekind Completeness]
    A totally ordered algebraic field $\mathbb{F}$ is \textbf{Dedekind-complete} if every Dedekind cut of $\mathbb{F}$ is generated by an element of $\mathbb{F}$. 
  \end{definition}

  \begin{lemma}[Rationals are Not Dedekind-Complete]
    $\mathbb{Q}$ is not Dedekind-complete. 
  \end{lemma}
  \begin{proof}
    Take a look at the cut
    \begin{equation}
      A = \{x \in \mathbb{Q} \mid x^2 < 2 \}, A^\prime = \{x \in \mathbb{Q} \mid x^2 \geq 2 \}
    \end{equation}
    We wish to show that $A^\prime$ has no minimum. Assume that it does, and call it $p$. Then, define 
    \begin{equation}
      q \coloneqq p - \frac{p^2 - 2}{p + 2} = \frac{2p + 2}{p + 2} 
    \end{equation}
    We can see that $p > 2 \implies p^2 - 2 > 0 \implies q < p$. But we also see that 
    \begin{equation}
      q^2 - 2 = \frac{2 (p^2 - 2)}{(p + 2)^2} \implies q^2 > 2
    \end{equation}
    Therefore, we have found a $q \in A^\prime$ that is smaller than $p$, a contradiction. 
  \end{proof}

  These Dedekind cuts are simply subsets of the power set of $\mathbb{Q}$, which always exists due to the \hyperref[st-power-set-axiom]{power set axiom}. Therefore, we can simply use the axiom of restricted comprehension (?) to create a well-defined set of Dedekind cuts.    

  \begin{definition}[Reals as the Dedekind-Completion of Rationals]
    Let $\mathbb{R}_D$ be the set of all Dedekind cuts $A$\footnote{For convenience we can uniquely represent $(A, A^\prime)$ with just $A$ since $A^\prime = \mathbb{Q} \setminus A$. }  of $\mathbb{Q}$. 

    \begin{equation}
      \mathbb{R}_D \coloneqq \{ A \in 2^{\mathbb{Q}} \mid (A, A^c) \text{ is a Dedekind cut}\}
    \end{equation}
    By doing this we can intuitively think of a real number as being represented by the set of all smaller rational numbers. Let $A, B \in \mathbb{R}_D$ two Dedekind cuts. Then, we define the following order and operations. 
    \begin{enumerate}
      \item \textit{Order}. $A \leq_{\mathbb{R}} B \iff A \subset B$. 
      \item \textit{Addition}. $A +_{\mathbb{R}} B \coloneqq \{ a +_{\mathbb{Q}} b \mid a \in A, b \in B \}$. 
      \item \textit{Additive Identity}. $0_{\mathbb{R}} \coloneqq \{x \in \mathbb{Q} \mid x < 0 \}$. 
      \item \textit{Additive Inverse}. $-B \coloneqq \{ a - b \mid a < 0 , b \in (\mathbb{Q} \setminus B) \}$.
      \item \textit{Multiplication}. If $A, B \geq 0$, then we define $A \times_{\mathbb{R}} B \coloneqq \{ a \times_{\mathbb{Q}} b \mid a \in A, b \in B, a, b \geq 0\} \cup 0_{\mathbb{R}}$. If $A$ or $B$ is negative, then we use the identity $A \times B = -(A \times_{\mathbb{R}} -B) = -(-A \times_{\mathbb{R}} B) = (-A \times_{\mathbb{R}} -B)$ to convert $A, B$ to both positives and apply the previous definition. 
      \item \textit{Multiplicative Identity}. $1_{\mathbb{R}} = \{x \in \mathbb{Q} \mid x < 1 \}$. 
      \item \textit{Multiplicative Inverse}. If $B > 0$, $B^{-1} \coloneqq \{ a \times_{\mathbb{Q}} b^{-1} \mid a \in 1_{\mathbb{R}}, b \in (\mathbb{Q} \setminus B) \}$. If $B$ is negative, then we compute $B^{-1} = -((-B)^{-1})$ by first converting to a positive number and then applying the definition above. 
    \end{enumerate}

    We claim that $(\mathbb{R}, +_{\mathbb{R}}, \times_{\mathbb{R}}, \leq_{\mathbb{R}})$ is a Dedekind-complete totally ordered field, and the canonical injection $\iota: \mathbb{Q} \rightarrow \mathbb{R}$ defined 
    \begin{equation}
      \iota(q) = \{x \in \mathbb{Q} \mid x < q \}
    \end{equation}
    is an ordered field isomorphism. 
  \end{definition} 
  \begin{proof}
    
  \end{proof}

  By the canonical injections $\mathbb{N} \rightarrow \mathbb{Z} \rightarrow \mathbb{Q} \rightarrow \mathbb{R}_D$, we can talk about whether this set has the Archimedean property. By construction, Archimidean is trivial since $\mathbb{R}_D$ contains $\mathbb{Q}$ which is Archimidean. 

  \begin{theorem}[Dedekind Reals is Archimedean]
    $\mathbb{R}_D$ satisfies the Archimedean principle. 
  \end{theorem}
  \begin{proof} 
    $\mathbb{R}_D$ contains $\mathbb{Q}$. 
  \end{proof}

  \begin{definition}[Axiomatic Construction of Dedekind-Reals]
    $\mathbb{R}_D^\prime$ is a totally ordered field that is Dedekind complete.\footnote{Just for this section, I will denote $\mathbb{R}^\prime$ as reals constructed axiomatically.}
  \end{definition}

  \begin{theorem}[Axiomatic Dedekind Reals is Archimedean]
    $\mathbb{R}_D^\prime$ satisfies the Archimedean principle. 
  \end{theorem}
  \begin{proof}
    Assume that this property doesn't hold. Then for any fixed $x$, $nx < y$ for all $n \in \mathbb{N}$. Consider the set 
    \begin{equation}
      A = \bigcup_{n \in \mathbb{N}} (-\infty, nx), \qquad B = \mathbb{R} \setminus A
    \end{equation}
    $A$ by definition is nonempty, and $B$ is nonempty since it contains $y$. Then, we can show that $a \in A, b \in B \implies a < b$ using proof by contradiction. Assume that there exists $a^\prime \in A, b^\prime \in B$ s.t. $a^\prime > b^\prime$. Since $a^\prime \in A$, there exists a $n^\prime \in \mathbb{N}$ s.t. $a^\prime \in (-\infty, n^\prime x) \iff a^\prime < n^\prime x$. But by transitivity of order, this means $b^\prime < n^\prime x \iff b^\prime \in (-\infty, n^\prime x) \implies b^\prime \in A$. 

    Going back to the main proof, we see that $A$ is upper bounded by $y$, and so by the least upper bound property it has a supremum $z = \sup{A}$. 
    \begin{enumerate}
      \item If $z \in A$, then by the induction principle\footnote{Note that $\mathbb{N}$ is defined recursively as $1 \in \mathbb{N}$ and if $n \in \mathbb{N}$, then $n+1 \in \mathbb{N}$. } $z + x \in A$, contradicting that $z$ is an upper bound. 
      \item If $z \not\in A$, then by the induction principle\footnote{The contrapositive of the recursive definition of $\mathbb{N}$ is: if $n \not\in \mathbb{N}$, then $n-1 \not\in \mathbb{N}$.} $z-x \not\in A \implies z-x \in B$. Since every element of $B$ upper bounds $A$ and since $x > 0$, this means that $z-x < z$ is a smaller upper bound of $A$, contradicting that $z$ is a least upper bound. 
    \end{enumerate}
    Therefore, it must be the case that $nx > y$ for some $n \in \mathbb{N}$. 
  \end{proof}

\subsection{Cauchy Completeness} 

  In many cases we are not working with ordered fields, and so different types of completeness may be more useful. In actual practice, you tend to use Cauchy completeness which only assumes a metric structure. 

  \begin{definition}[Cauchy Sequence]
    A sequence $(x_n)_n$ in a metric space $(X, d)$ is a \textbf{Cauchy sequence} if $\forall \epsilon > 0$, $\exists N \in \mathbb{N}$ s.t.  
    \begin{equation}
      n, m \geq N \implies d(x_n, x_m) < \epsilon
    \end{equation}
  \end{definition}

  To motivate this definition, note that in a general topological space $X$, we can define convergence of a sequence $x_n \to x$ perfectly fine. However, take some subset $U \subset X$, and let $x$ be a limit point of $U$. In this case, $x_n$ does not converge in $U$, but it does converge to something outside of $U$---namely, $x \in X$. This is similar to $\mathbb{Q} \subset \mathbb{R}$, where $x$ is an irrational point. However, we are trying to \textit{construct} $\mathbb{R}$, so Cauchy convergence allows us to speak of convergence without actually referring to \textit{what} a sequence is converging to. 

  Note that it is not sufficient to say that a sequence is Cauchy by claiming that each term becomes arbitrarily close to the preceding term. That is, 
  \begin{equation}
    \lim_{n \rightarrow \infty} d(x_{n+1}, x_{n}) = 0
  \end{equation}

  \begin{example}[Adjacent Terms Converging Doesn't Imply Sequence is Cauchy]
    For example, look at the sequence 
    \begin{equation}
      a_n = \sqrt{n} \implies a_{n+1} - a_{n} = \frac{1}{\sqrt{n+1} + \sqrt{n}} < \frac{1}{2\sqrt{n}}
    \end{equation}
    However, it is clear that $a_n$ gets arbitrarily large, meaning that a finite interval can contain at most a finite number of terms in $\{a_n\}$. 
  \end{example}

  It is often more convenient to think of the limit of the \textit{diameter} of rest of the sequence. That is, a sequence is Cauchy if 
  \begin{equation}
    \lim_{n \rightarrow \infty} \mathrm{diam}\{x_{m}\}_{m \geq n} = 0
  \end{equation}

  It is trivial that convergence implies Cauchy convergence, but the other direction is not true. Therefore, we would like to work in a space where these two are equivalent, and this is called completeness. Therefore, we can construct the reals as equivalence classes over Cauchy sequences. Rather than using the order, we take advantage of the metric. 

  \begin{definition}[Cauchy Completeness]
    A metric space $(X, d)$ is \textbf{Cauchy complete} if every Cauchy sequence in that space converges to an element in $X$. 
  \end{definition} 
  
  $\mathbb{Q}$ is not Cauchy-complete. Let $a_n$ be the largest number $x$ up to the $n$th decimal expansion such that $x^2$ does not exceed $2$. The first few terms are 
  \begin{equation}
    1.4, 1.41, 1.414, \ldots
  \end{equation}
  In this case, we can see that this is Cauchy since at the $n$th element and on, the first $n$ decimal places are kept fixed and so the most that the rest of the sequence can deviate by is $10^{-n}$. 

  \begin{definition}[Reals as the Cauchy-Completion of the Rationals]
    Let $\mathbb{R}_C$ be the quotient space of all Cauchy (under the Euclidean metric) sequences $(x_n)$ of rational numbers with the equivalence relation $(x_n) = (y_n)$ iff their difference tends to $0$.\footnote{This equivalence class reflects that the same real number can be approximated in many different sequences. In fact, this shows \textit{by definition} that $1, 1, \ldots$ and $0.9, 0.99, 0.999, \ldots$ are the same number!} That is, for every rational $\epsilon > 0$, there exists an integer $N$ s.t. for all naturals $n > N$, $|x_n - y_n| < \epsilon$. 
    \begin{enumerate}
      \item \textit{Order}. $(x_n) \leq_{\mathbb{R}} (y_n)$ iff $x = y$ or there exists $N \in \mathbb{N}$ s.t. $x_n \leq_{\mathbb{Q}} y_n$ for all $n > N$. 
      \item \textit{Addition}. $(x_n) + (y_n) \coloneqq (x_n + y_n)$. 
      \item \textit{Additive Identity}. $0_{\mathbb{R}} \coloneqq (0_{\mathbb{Q}})$. 
      \item \textit{Additive Inverse}. $-(x_n) \coloneqq (-x_n)$. 
      \item \textit{Multiplication}. $(x_n) \times_{\mathbb{R}} (y_n) = (x_n \times_{\mathbb{Q}} y_n)$. 
      \item \textit{Multiplicative Identity}. $1_{\mathbb{R}} \coloneqq (1)$. 
      \item \textit{Multiplicative Inverse}. $(x_n)^{-1} \coloneqq (x_n^{-1})$. 
    \end{enumerate}
    We claim that $(\mathbb{R}, +_{\mathbb{R}}, \times_{\mathbb{R}}, \leq_{\mathbb{R}})$ is a totally ordered field, and the canonical injection $\iota: \mathbb{Q} \rightarrow \mathbb{R}$ defined 
    \begin{equation}
      \iota(q) = (q)
    \end{equation} 
    is an ordered field isomorphism. Finally, by construction $\mathbb{R}$ is Cauchy-complete. 
  \end{definition}
  \begin{proof}
    
  \end{proof}

  \begin{theorem}[Cauchy Reals is Archimidean]
    $\mathbb{R}_C$ satisfies the Archimedean principle. 
  \end{theorem}
  \begin{proof}
    $\mathbb{R}_C$ contains $\mathbb{Q}$, which is Archimidean. 
  \end{proof}

  The best thing about Cauchy completeness is that we can just take $\mathbb{Q}^n$ to create $\mathbb{R}^n$. It becomes quite general. However, note that first, Cauchy completion depends on \textit{which} metric you use to complete it. 

  \begin{example}[P-adic Numbers]
    Let $p$ be a prime number. For a non-zero rational $x = p^k \cdot \frac{a}{b}$ where $p \nmid a, b$, define the \textbf{p-adic norm}\footnote{This measures divisibility by $p$: the more $p$ divides $x$, the smaller $|x|_p$. For example, $|8|_2 = 2^{-3} = 1/8$ and $|3|_2 = 1$.} as 
    \begin{equation}
      |x|_p \coloneqq p^{-k}, \qquad |0|_p \coloneqq 0
    \end{equation}
    The \textbf{p-adic numbers} $\mathbb{Q}_p$ are the Cauchy completion of $\mathbb{Q}$ with respect to the p-adic metric $d_p(x,y) = |x - y|_p$. This set not only does not satisfy the Archimidean principle; it doesn't even have a natural ordering! 
  \end{example} 

  \begin{definition}[Axiomatic Construction of Cauchy-Reals]
    $\mathbb{R}_D^\prime$ is a totally ordered field that is Cauchy complete and that satisfies the Archimidean principle.
  \end{definition}

  Note that we require the extra Archimidean assumption in the axiomatic construction. 

  \begin{example}[Ordered Cauchy-Complete Fields that are Not Archimidean]
    Provide examples of ordered, Cauchy-complete fields that are not Archimedean. Hyperreals? 
  \end{example}

\subsection{Least Upper Bound Completeness}

  \begin{definition}[Least Upper Bound Property]
    A totally ordered algebraic field $\mathbb{F}$ (must it be a field?) is \textbf{least-upper-bound complete}, or is said to satisfy the \textbf{least upper bound (LUB) property}, if every nonempty set of $\mathbb{F}$ having an upper bound must have a least upper bound (supremum) in $\mathbb{F}$. 
  \end{definition} 

  \begin{theorem}[LUB is Equivalent to GLB]
    A set $(\mathbb{F}, \leq)$ has the least upper bound property iff it has the \textit{greatest lower bound property}, which states that every set bounded below has a greatest lower bound. 
  \end{theorem}
  \begin{proof}
    We will prove one direction since the other is the same logic. Let $S \subset X$ be a nonempty set that is bounded below by some $l \in X$. Let $L \subset X$ be the set of all lower bounds of $S$. Since $l$ exists, it is nonempty. Furthermore, $L$ is bounded above by any element of $S$. Due to LUB property $L$ has a least upper bound, call it $z = \sup{L}$. We claim that $z = \inf{S}$. 
    \begin{enumerate}
      \item $z$ is a lower bound of $S$. Assume that it is not. Then there exists $s \in S$ s.t. $s < z$. But by construction $s$ is an upper bound for $L$ and so $z$ s not the \textit{least} upper bound, a contradiction. 
      \item $z$ is a \textit{greatest} lower bound. Assume that $z$ is not. Then there exists a $z^\prime \in X$ s.t. $z < z^\prime \leq s$ for all $s \in S$. But since $z^\prime, z$ are lower bounds, this means $z, z^\prime \in L$ by definition and $z < z^\prime$ contradicts the fact that $z$ is an upper bound of $L$. 
    \end{enumerate}
    We are done. 
  \end{proof}

  $\mathbb{Q}$ does not satisfy the least upper bound property, but proving this can be tricky for the first time. We state this as a lemma. 

  \begin{theorem}[Rationals Doesn't Satisfy LUB Property]
    $\mathbb{Q}$ does not satisfy the LUB property. 
  \end{theorem}
  \begin{proof}
    Assume it does, and let us denote 
    \begin{equation}
      p \coloneqq \sup \{x \mid \mathbb{Q} \mid x^2 < 2\} \in \mathbb{Q}
    \end{equation}
    The key here is to find another rational that we can always ``squeeze'' in between $p$ and $2$. This can be done with the Archimidean principle, which is already satsified in $\mathbb{Q}$. Since we have \href{thm:sqrt2-irrational}{proved that there exists no rational that squares to $2$}, we only need to consider the two cases. 
    \begin{enumerate}
      \item $p^2 < 2$. Take $\epsilon \in \mathbb{Q}$ so small that 
      \begin{equation}
        p^2 + 2 p \epsilon + \epsilon^2 = (p + \epsilon)^2 < 2
      \end{equation}
      To show complete steps, we can see that by density of reals, there exists some rational $r$ s.t. $0 < r < 2 - p^2$. Therefore, we can invoke Archimidean principle to find a $n \in \mathbb{N}$ s.t. $\epsilon = p/n < 2 p/n < r$. Therefore, $p$ is not an upper bound, so this cannot be true. 

      \item $p^2 > 2$. We can again take an $\epsilon \in \mathbb{Q}$ so small that 
      \begin{equation}
        p^2 - 2 p \epsilon + \epsilon^2 = (p - \epsilon)^2 > 2
      \end{equation}
      Therefore, $p$ is not least, so this cannot be true. 
    \end{enumerate}
  \end{proof}

  \begin{definition}[Axiomatic Construction of Reals with LUB Property]
    $\mathbb{R}_I^\prime$ is a totally ordered field that satisfies the least upper bound property.
  \end{definition}

  Note that we don't need to explicitly assume Archimidean principle here. The LUB property is strong enough that it implies Archimidean!

  \begin{theorem}[LUB Property Implies Archimidean]
    $\mathbb{R}_I^\prime$ is Archimidean. 
  \end{theorem}

  Now let's see how our previous constructions of the reals relate to the LUB property. 

  \begin{theorem}[Dedekind Completed Reals Satisfies LUB Property]
    $\mathbb{R}_D$ satisfies the least upper bound property. 
  \end{theorem}
  \begin{proof}
    Let $\mathcal{A}$ be a nonempty subset of $\mathbb{R}_D$ bounded from above by $T$. Then, $\forall A \in \mathcal{A}$, $A \coloneqq (A, A^c)$ is a Dedekind cut, and we can define
    \begin{equation}
      (B, B^\prime) \coloneqq \bigg( \bigcup_{A \in \mathcal{A}} A, \bigcap_{A \in \mathcal{A}} A^c \bigg)
    \end{equation} 
    We claim that this is a Dedekind cut. 
    \begin{enumerate}
      \item First, it is nonempty set $\mathcal{A} \neq \emptyset$ and so for each $A \in \mathcal{A}$, $A \subset \mathbb{Q}$ is nonempty. It is also not all of $\mathbb{Q}$ since $T$ is an upper bound of $\mathcal{A}$, and so $T \geq a \; \forall a \in A \; \forall A \in \mathcal{A}$, which implies that 
      \begin{equation}
        T \not\in \bigcup_{A \in \mathcal{A}} A
      \end{equation}

      \item Now let $x \in B, y \in B^\prime$. Then, $x \in A_0$ for some $A_0 \in \mathcal{A}$, and since $y$ is in the intersection of all the corresponding $A^c$, it must be in the corresponding $A_0^c$. Therefore we invoke the Dedekind cut property of $(A_0, A_0^c)$ and see $x < y$. 

      \item Finally, for the sake of contradiction, let $m \in \mathbb{Q}$ be the maximum of $B$. Then, $m \in A^\ast$ for some $A^\ast \in \mathcal{A}$. But $m$ is an upper bound for the whole $B$, so this means that $m = \max\{A^\ast\}$, which contradicts the fact that lower cut cannot have a rational maximum. 
    \end{enumerate}
    Now we claim that $B$ is the supremum. It is an upper bound since 
    \begin{equation}
      A \subset B = \bigcup_{A \in \mathcal{A}} A \quad \forall A \in \mathcal{A}
    \end{equation}
    To prove least, we should see that if $(S, S^c)$ is another upper bound of $\mathcal{A}$, then $A \subset S$ for all $A \in \mathcal{A} \implies B = \bigcup_{A \in \mathcal{A}} A \subset S$, which establishes that $B$ is least. 
  \end{proof}

\subsection{Nested Intervals Completeness}

  The next flavor we present is nested-intervals completeness.  This is the least popular way to construct the reals, and it is used more as a post-hoc tool to analyze the reals after you construct it using either of the two previous methods. 

  \begin{definition}[Nested Interval Completeness]
    Let $\mathbb{F}$ be a totally ordered algebraic field. Let $I_n= [a_n, b_n]$ ($a_n < b_n$) be a sequence of decreasing nested intervals that are 
    \begin{enumerate}
      \item closed, 
      \item bounded, 
      \item nested, $I_1 \supset I_2 \supset I_3 \supset \ldots$ 
      \item and decreasing to $0$ in the sense that $\lim_{n \to \infty} b_n - a_n = 0$. 
    \end{enumerate}
    $\mathbb{F}$ is \textbf{nested-interval complete} if the intersection of all of these intervals $I_n$ contains exactly one point. 
    \begin{equation}
      \bigcup_{n=1}^\infty I_n \in \mathbb{F}
    \end{equation}
  \end{definition}

  Note that defining nested intervals requires only an ordered field. One may look at this and try to ask if this is a specific instance of the following conjecture: The intersection of a nested sequence of nonempty closed sets in a topological space has exactly 1 point. This claim may not even make sense, actually. If we define nested in terms of proper subsets, then for a finite topological space a sequence cannot exist since we will run out of open sets and so this claim is vacuously true and false. If we allow $S_n = S_{n+1}$ then we can just select $X \supset X \supset \ldots$, which is obviously not true. However, a slightly weaker claim is that every proper nested non-empty closed sets has a non-empty intersection is a consequence of compactness. 

  \begin{theorem}[Rationals are Not Nested-Interval Complete]
    $\mathbb{Q}$ is not nested-interval complete. 
  \end{theorem}
  \begin{proof}
    This is a nice proof that uses a class method of algorithmically selecting nested intervals that satisfy a following property. This trick will be used many times in analysis. 
    
    For the sake of contradiction, let us assume that $\mathbb{Q}$ satisfies nested intervals completeness. Since $\mathbb{Q}$ is countable, enumerate it as $q_1, q_2, \ldots$. 
    \begin{enumerate}
      \item Choose any closed interval $I_1$ of length $1$ with rational endpoints that doesn't contain $q_1$. 
      \item Now partition $I_1$ into two segments of equal length, and choose $I_2$ to be the segment that doesn't contain $q_2$.  
    \end{enumerate}
    At this point, $I_n$ is a closed interval with rational endpoints of length $2^{-n}$ that cannot contain $q_n$. Therefore, $\cap_n I_n$ cannot contain any $q_n \in \mathbb{Q}$. But this contradicts our assumption of nested interval completeness.  
  \end{proof} 

  One may ask: what is the relationship between LUB property and nested-intervals completeness? It turns out that they are equivalent. 

  \begin{theorem}[LUB is Equivalent to Nested Interval Completeness in Reals]
    Listed. 
    \begin{enumerate}
      \item $\mathbb{R}_L^\prime$ satisfies nested intervals completeness. 
      \item $\mathbb{R}_I^\prime$ satisfies least upper bound completeness. 
    \end{enumerate}
  \end{theorem}
  \begin{proof}
    Listed. 
    \begin{enumerate}
      \item Note that $\{a_n\}$ is bounded above by $b_1$. Therefore by LUB property it must have a supremum, call it $x = \sup_n \{a_n\}$. Then, we see that $a_n \leq x \leq b_n$ for all $n$, and so $x$ is in the intersection. 
    \end{enumerate}
  \end{proof}

  \begin{theorem}[Cantor's Intersection Theorem]
    $\mathbb{R}$ is nested-interval complete. 
  \end{theorem}
  \begin{proof}
    We prove this by first proving the claim that given nested, closed, and bounded sets $I_n$ (not even necessarily intervals), then 
    \begin{equation}
      \bigcap_{n=1}^\infty I_n \neq \emptyset
    \end{equation}
    Suppose this is not true. For every $x \in \mathbb{R}$, there exists a $n_x$ s.t. $x \not\in I_n$ (and all later $I_m$ for $m > n$). Let $O_n = I_n^c$ open sets. Then, $\mathbb{R} \subset \cup_n O_n$ In particular, $I_1 \subset \cup_n O_n$. But $I_1$ is closed and bounded. So we can extract a finite subcover. $O_{n_1}, O_{n_2}, \ldots, O_{n_m}$ (ordered $n_1 < n_2 < \ldots< n_m$). Then since $O_n$ are increasing, $I_1 \subset O_{n_m} = I_{n_m}^c$. But $I_{n_m} \subset I_1$, a contradiction. 

    Now with this, we know that because the limits of the endpoints of the intervals go to $0$, then there cannot be more than 2 points in the intersection. Thus there must be 1 unique point. 
  \end{proof} 

  \begin{definition}[Axiomatic Construction of Reals with Nested Intervals]
    $\mathbb{R}_I$ is a totally ordered field that 
    \begin{enumerate}
      \item satisfies the nested intervals completeness, and 
      \item satisfies the Archimidean principle.
    \end{enumerate}
  \end{definition}

\subsection{Properties of the Real Line}  

  Perfect, now all that remains is to unite the two constructions of the reals. 

  \begin{theorem}[Dedekind and Cauchy Complete Reals are Isomorphic]
    Given that $\mathbb{R}_D$ is the Dedekind-completed version of the rationals and $\mathbb{R}_C$ is the Cauchy-completed version of the rationals, we claim that the two are isomorphic as ordered fields. 
    \begin{equation}
      \mathbb{R}_D \simeq \mathbb{R}_C
    \end{equation}
  \end{theorem}
  \begin{proof}
    
  \end{proof}

  Therefore, it doesn't really matter which one we talk about, and we can refer to \textit{the} real numbers as a single set. Great! Now we can finally feel satisfied about defining metrics, norms, and inner products as mappings to the codomain $\mathbb{R}$. 

  \begin{definition}[Reals (as Construction from Rationals)]
    The \textbf{reals} $\mathbb{R}$ is the totally ordered complete Archimedean field constructed as the completion\footnote{either one} of $\mathbb{Q}$. 
  \end{definition}

  So far, we have taken the completion of the rationals as our main mode of construction. However, we can take an axiomatic approach, and it turns out that there is only one set, up to isomorphism, that satisfies all these properties. 

  \begin{theorem}[Axiomatic Definition of Reals]
    The \textbf{real numbers}, denoted $\mathbb{R}$, is any totally ordered complete Archimedean field. $\mathbb{R}$ is unique up to field isomorphism. That is, if two individuals construct two ordered complete Archimedean fields $\mathbb{R}_A$ and $\mathbb{R}_B$, then 
    \begin{equation}
      \mathbb{R}_A \simeq \mathbb{R}_B
    \end{equation}
  \end{theorem} 
  \begin{proof}
    The proof is actually much longer than I expected, so I draw a general outline.\footnote{Followed from \href{https://math.ucr.edu/~res/math205A/uniqreals.pdf}{here}.} We want to show how to construct an isomorphism $f: \mathbb{R}_A \rightarrow \mathbb{R}_B$. 
    \begin{enumerate}
      \item Realize that there are unique embeddings of $\mathbb{N}$ in $\mathbb{R}_A$ and $\mathbb{R}_B$ that preserve the inductive principle, the order, closure of addition, and closure of multiplication, the additive identity, and the multiplicative identity. Call these ordered doubly-monoid (since it's a monoid w.r.t. $+$ and $\times$) homomorphisms $\iota_A, \iota_B$. 
      \item Construct an isomorphism $f_1: \iota_A(\mathbb{N}) \rightarrow \iota_B(\mathbb{N})$ that preserves the inductive principle, order, addition, and multiplication. This is easy to do by just constructing $f_1 = \iota_B \circ \iota_A^{-1}$. 
      \item Extend $f_1$ to the ordered ring isomorphism $f_2$ by explicitly defining what it means to map additive inverses, i.e. negative numbers. 
      \item Extend $f_2$ to the ordered field isomorphism $f_3$ by explicitly defining what it means to map multiplicative inverses, i.e. reciprocals. 
      \item Extend $f_3$ to the ordered field isomorphism on the entire domain $\mathbb{R}_A$ and codomain $\mathbb{R}_B$. There is no additional operations that we need to support, but we should explicitly show that this is both injective and surjective, which completes our proof. 
    \end{enumerate}
  \end{proof}

  It seems that the real numbers is \textit{any} set that satisfies the definition above. Therefore, a line $\mathbb{L}$ with $+$ associated with the translation of $\mathbb{L}$ along itself and $\cdot$ associated with the "stretching/compressing" of the line around the additive origin $0$ is a valid representation of the reals. $\mathbb{R}$ can also be represented as an uncountable list of numbers with possibly infinite decimal points, known as the decimal number system. 
  \begin{equation}
    \ldots, -2.583\ldots, \ldots , 0, \ldots, 1.2343\ldots, \ldots, \sqrt{2}, \ldots
  \end{equation}

  The first property we should know is that the reals are uncountable. 

  \begin{theorem}[Cantor's Diagonalization]
    $\mathbb{R}$ is uncountable.
  \end{theorem}
  \begin{proof}
    We proceed by contradiction. Suppose the real numbers are countable. Then there exists a bijection $f: \mathbb{N} \to \mathbb{R}$. This means we can list all real numbers in $[0,1]$ as an infinite sequence.\footnote{This must be explicitly proven, but we can take the set of all Cauchy sequences of rationals in their decimal expansion and construct the reals this way.}
    
    \begin{align}
      f(1) &= 0.a_{11}a_{12}a_{13}\dots \\
      f(2) &= 0.a_{21}a_{22}a_{23}\dots \\
      f(3) &= 0.a_{31}a_{32}a_{33}\dots \\
      &\vdots
    \end{align}
    
    where each $a_{ij}$ is a digit between 0 and 9.
    
    Now construct a new real number $r = 0.r_1r_2r_3\dots$ where:
    \begin{equation}
      r_n = \begin{cases}
        1 & \text{if } a_{nn} \neq 1 \\
        2 & \text{if } a_{nn} = 1
      \end{cases}
    \end{equation}
    This number $r$ is different from $f(n)$ for every $n \in \mathbb{N}$, since $r$ differs from $f(n)$ in the $n$th decimal place. Therefore $r \in [0,1]$ but $r \notin \text{range}(f)$, contradicting that $f$ is surjective. Thus our assumption that the real numbers are countable must be false.
  \end{proof}

  With this, we can add the inner product, metric, and topology. 

\subsection{Exponentials, Roots, and Logarithms} 

  Now we will focus on some other operations that become well-defined in the reals. We know that $x^{n}$ for $n \in \mathbb{N}$ denotes repeated multiplication and $x^{-1}$ denotes the multiplicative inverse. We need to build up on this notation. As a general outline, we will show that $x^{-n}$ is well defined, then $x^q, q \in \mathbb{Q}$ is well-defined, and finally $x^r, r \in \mathbb{R}$ is well-defined. For the naturals, we have defined $x^n$ as the repeated multiplication of $n$. It is trivial that the canonical injection $\iota_0: \mathbb{N} \rightarrow \mathbb{R}$ commutes with the exponential map of naturals. We prove that $\iota_1: \mathbb{Z} \rightarrow \mathbb{R}$ also commutes. 

  \begin{lemma}[Integer Exponents]
    We have 
    \begin{enumerate}
      \item For $x_1, \ldots, x_n \in \mathbb{R}$, $(x_1 \ldots x_n)^{-1} = x_n^{-1} \ldots x_1^{-1}$. 
      \item For $x \in \mathbb{R}$, $x > 0$, $(x^n)^{-1} = (x^{-1})^n$. This value is denoted $x^{-n}$. 
      \item For $x \in \mathbb{R}$ and $w, z \in \mathbb{Z}$, $x^{w + z} = x^w x^z$. 
      \item For $w, z \in \mathbb{Z}$, $x^{wz} = (x^z)^w = (x^w)^z$. 
    \end{enumerate}
  \end{lemma}
  \begin{proof}
    Listed. 
    \begin{enumerate}
      \item The proof is trivial, but for $n = 2$ and $x_1 = x, x_2 = y$, we see that by associativity, $(x^{-1} x^{-1}) (x y) = y^{-1} (x^{-1} x) y = y^{-1} y = 1$ and we know inverses are unique. 
      \item Set $x_i = x$ using (1). 
      \item If $w, z > 0$ this is trivial by the associative property. If either or both are negative, say $w < 0 < z$, then we set $w^\prime = -w > 0$ and using (2) we know that 
      \begin{equation}
        x^{w} x^{z} = (x^{-1})^{w^\prime} x^z = x^{-w^\prime + z} = x^{w + z}
      \end{equation}
      by associativity in the second last equality. 
    \end{enumerate}
  \end{proof}

  Therefore, we have successfully defined $x^z$ for all $z \in \mathbb{Z}$, and if $z$ is negative, we're allowed to ``swap'' the $-1$ and $|z|$ in the exponents. Now we want to extend this into rational exponents, first by proving the existence and uniqueness of $n$th roots for any real. The proof is a little involved, but the general idea is that we want to use the LUB property to define the $n$th root as the supremum of a set.  

  \begin{theorem}[Existence of Nth Roots]
    For any real $x > 0$ and every $n \in \mathbb{N}$ there is one and only one positive real $y \in \mathbb{R}$ s.t. $y^n = x$. This is denoted $x^{1/n}$. 
  \end{theorem}
  \begin{proof}
    Let $E$ be the set consisting of all reals $t \in \mathbb{R}$ s.t. $t^n < x$. We show that 
    \begin{enumerate}
      \item it is nonempty. Consider $t = x/(1+x)$. Then $0 \leq t < 1 \implies t^n \leq t < x$. Thus $t \in E$ and $E$ is nonempty. 
      \item it is bounded. Consider any number $s = 1 + x$. Then $s^n \geq s > x$, so $s \not\in E$, and $s = 1 + x$ is an upper bound of $E$. 
    \end{enumerate}
    Therefore, $E$ is a nonempty set that is upper bounded, so it has a least upper bound, called $y = \sup{E}$. We claim that $y^n = x$, proving by contradiction. For both cases, we use the fact that the identity $b^n - a^n = (b - a) (b^{n-1} + b^{n-2} a + \ldots + a^{n-1})$ gives the inequality 
    \begin{equation}
      b^n - a^n < (b-a) n b^{n-1} \text{ for } 0 < a < b
    \end{equation}
    \begin{enumerate}
      \item Assume $y^n < x$. Then we choose a fixed $0 < h < 1$ s.t. 
      \begin{equation}
        h < \frac{x - y^n}{n(y + 1)^{n-1}}
      \end{equation}
      Then by putting $a = y, b = y + h$, we have 
      \begin{equation}
        (y + h)^n - y^n < hn (y + h)^{n-1} < hn (y + 1)^{n-1} < x - y^n 
      \end{equation}
      and thus $y^n < (y + h)^n < x$. This means that $y + h \in E$, and so $y$ is not an upper bound. 

      \item Assume $y^n > x$. Then we set a fixed number 
      \begin{equation}
        k = \frac{y^n - x}{n y^{n-1}} 
      \end{equation}
      Then $0 < k < y$. If we take any $t \in \mathbb{R}$ s.t. $t \geq y - k$, this implies that $t^n \geq (y -k)^n \implies -t^n \geq -(y - k)^n$, and so 
      \begin{equation}
        y^n - t^n \leq y^n - (y - k)^n < k ny^{n-1} = y^n - x
      \end{equation}
      Thus $t^n > x$ and $t \not\in E$. So it must be the case that $t < y - k$, and so $y - k$ is an upper bound of $E$, contradicting that $y$ is least. 
    \end{enumerate}
  \end{proof}

  At this point, rooting has been introduced as sort of an independent map from exponentiation. We show that they have the nice property of commuting. 

  \begin{lemma}[Rooting and Exponentiation Commute]
    For $p \in \mathbb{Z}, q \in \mathbb{N}$ and $x \in \mathbb{R}$ with $x > 0$, we have 
    \begin{equation}
      (x^{p})^{1/q} = (x^{1/q})^p
    \end{equation}
  \end{lemma}
  \begin{proof}
    If $p > 0$, then let $r = (x^p)^{1/q}$. By definition $r^q = x^p$. Let $s = x^{1/q}$ By definition $s^q = x$. Therefore $r^q = (s^q)^p = s^{qp}$ from the lemma on integer exponents. But since roots are well-defined and unique 
    \begin{equation}
      r = (r^q)^{1/q} = (s^{qp})^{1/q} = s^p \implies (x^p)^{1/q} = (x^{1/q})^p
    \end{equation}
    If $p = 0$, this is trivially $0$, and if $p < 0$ the by the same logic with $p = -p^\prime$ for $p^\prime > 0$ and $y = x^{-1} > 0$. we know 
    \begin{align}
      (x^p)^{1/q} = \big( (y^{-1})^{-p^\prime} \big)^{1/q} = (y^{-(-p^\prime)})^{1/q} & = (y^{p^\prime})^{1/q} \\ 
                         & = (y^{1/q})^{p^\prime} = ((x^{-1})^{1/q})^{p^\prime} = (x^{1/q})^{-p^\prime} = (x^{1/q})^p
    \end{align}
  \end{proof}

  \begin{theorem}[Rational Exponential Function]
    Given $m, p \in \mathbb{Z}$ and $n, q \in \mathbb{N}$, prove that 
    \begin{equation}
      (b^m)^{1/n} = (b^p){1/q}
    \end{equation}
    Hence it makes sens to define $b^r = (b^m)^{1/n}$, since every element of the equivalence class $r$ of each rational number maps to the same value. 
  \end{theorem} 
  \begin{proof}
    Since $m/n = p/q \implies mq = np$, 
    \begin{align}
      b^{mq} = b^{np} & \implies (b^m)^q = (b^p)^n \\
                      & \implies b^m = ((b^m)^q)^{1/q} = ((b^p)^n)^{1/q} \\
                      & \implies b^m = ((b^p)^{1/q})^n \\
                      & \implies (b^m)^{1/n} = (b^p)^{1/q}
    \end{align}
    Therefore we can define for any $r \in \mathbb{Q}$ 
    \begin{equation}
      x^r = x^{m/n} = (x^{m})^{1/n} = (x^{1/n})^m
    \end{equation}
    where the final equality holds from the commutativity of rooting and exponentiation. 
  \end{proof}

  It turns out that this is a homomorphism. 

  \begin{corollary}[Rational Exponential Function is a Homomorphism]
    The rational exponential function is a homomorphism. That is, given $r, s \in \mathbb{Q}$ and $x \in \mathbb{R}$, 
    \begin{equation}
      x^{r + s} = x^r \cdot x^s
    \end{equation}
  \end{corollary}
  \begin{proof}
    Let $r = m/n, s = p/q$. Then 
    \begin{align}
      x^{r+s} = x^{m/n + p/q} & = x^{\frac{mq + np}{nq}} && \tag{addition on $\mathbb{Q}$}\\
                              & = (x^{mq + np})^{1/nq} && \tag{exp and roots commute}\\
                              & = (x^{mq} + x^{np})^{1/nq} && \tag{int exp lemma}\\
                              & = (x^{mq})^{1/nq} (x^{np})^{1/nq} && \tag{int exp lemma}\\
                              & = x^{mq/nq} x^{np/nq} && \tag{exp and roots commute} \\
                              & = x^{m/n} x^{p/q} && \tag{relation from $\mathbb{Q}$}
    \end{align}
  \end{proof}

  With rational exponents defined, we can use the least upper bound property to define a consistent extension of a real exponent. 
  
  \begin{lemma} 
    If $r \in \mathbb{Q}$ with $r \geq 0$, then for $x \in \mathbb{R}$, $x > 1$, $1 \leq b^r$. 
  \end{lemma}
  \begin{proof}
    Let $r = m/n$. Then $x^r = x^{m/n} = (x^m)^{1/n}$. Since $1 < x$, and $m \geq 0$, we have 
    \begin{equation}
      1 \leq x \leq x^2 \leq \ldots \leq x^m \implies 1 \leq b^m
    \end{equation}
    Now set $y = x^{m/n}$ and assume that $y < 1$. Then 
    \begin{equation}
      x^m = y^n < y^{n-1} < \ldots < y < 1
    \end{equation}
    and so $x^m < 1$, which is a contradiction. So it must be the case that $y > 1$. 
  \end{proof}

  \begin{lemma}[Monotonicity of Rational Exponents]
    If $x, y \in \mathbb{R}$, then for any rational $r \in \mathbb{Q}$ with $r < x + y$, there exists a $p, q \in \mathbb{Q}$ s.t. $p < x, q < y$ and $p + q = r$. The converse is true as well. 
  \end{lemma}
  \begin{proof}
    $r < x + y \implies r - y < x$. By density of $\mathbb{Q}$ in $\mathbb{R}$, we can choose $r - y < p < x$. Then $-r + y > -p > x \implies r - r + y > r - p > r - x \implies y > r - p > r - x$, and we set $q = r - p$. We are done. The converse is trivial since given $p, q \in \mathbb{Q}$ with $p < x, q < y$, then by the ordered field properties $p + q < x + y$. 
  \end{proof}

  \begin{corollary}[Real Exponential Function]
    Given $x\in \mathbb{R}$, we define 
    \begin{equation}
      B(x) \coloneqq \{ x^q \in \mathbb{R} \mid q \in \mathbb{Q}, \; q \leq x \}
    \end{equation}
    We claim that given $r \in \mathbb{R}$, 
    \begin{equation}
      x^r \coloneqq \sup B(r)
    \end{equation}
    is well-defined and is a homomorphism extension of the rational exponential function. That is, 
    \begin{equation}
      \sup{B(x + y)} = \sup{B(x)} \cdot \sup{B(y)}
    \end{equation}
  \end{corollary}
  \begin{proof}
    To show that $x^r \coloneqq \sup B(r)$ where $B(r) = \{x^t \in \mathbb{R} \mid t \in \mathbb{Q}, t \leq r \}$, 
    \begin{enumerate}
      \item We show it's an upper bound. Assume it wasn't. Then $x^r < x^t$ for some $t \in \mathbb{Q}$ satisfying $t \leq r$. But $t \leq r \implies 0 \leq r - t$, and by the previous lemma, $1 \leq x^{r - t}$. So $1 \leq x^{r-t} = x^{r} x^{-t} = x^r (x^t)^{-1} \implies x^t \leq x^r$, which is a contradiction. 
      \item We show that it is least. Assume that it is not. Then $\exists r^\prime \in \mathbb{Q}$ s.t. $x^t \leq x^{r^\prime}$ and $r^\prime < r$. Now let $s \in \mathbb{Q}$ be an element between $r^\prime$ and $r$, which is guaranteed to exist due to density of rationals in reals. But $s < r$, so by definition $x^s \in B(r)$, but 
      \begin{align}
        0 < s - r^\prime & \implies 1 < b^{s - r^\prime} \\
                         & \implies b^{r^\prime} (b^{r^\prime})^{-1} < b^s (b^{-r^\prime}) \\
                         & \implies 1 < b^s (b^{r^\prime})^{-1} \\
                         & \implies b^{r^\prime} < b^s
      \end{align}
      and so $b^{r^\prime}$ is not an upper bound for $B(r)$. By contradiction, $b^r$ is least. 
    \end{enumerate}
    Since this is defined, the analogous definition for real numbers is consistent with that of hte rationals, and it is upper bounded by the Archimedean principle, so such a supremum must exist. Note that $t$ is rational. For the second part, from the previous lemma and the homomorphism properties of the rational exponent, 
    \begin{align}
      B(x + y) = B^\prime (x + y) & \coloneqq \{b^{p+q} \in \mathbb{R} \mid p, q \in \mathbb{Q}, p \leq x, q \leq y\} \\
                                  & = \{b^p b^q \in \mathbb{R} \mid p, q \in \mathbb{Q}, p \leq x, q \leq y\} \\
    \end{align}
    Therefore we can treat $B$ and $B^\prime$ as the same set. 
    \begin{enumerate}
      \item Prove upper bound $\sup{B(x + y)} \leq \sup{B(x)} \sup{B(y)}$. Given $\alpha \in B^\prime (x + y)$, there exists $p_\alpha, q_{\alpha} \in \mathbb{Q}$ (with $p_\alpha < x$, $q_\alpha < y$) s.t. $b^{p_{\alpha}} b^{q_{\alpha}} = \alpha$. But 
      \begin{equation}
        b^{p_{\alpha}} b^{q_{\alpha}} \leq \sup_{p_{\alpha}} \{ b^{p_{\alpha}}\} \cdot \sup_{q_{\alpha}} \{b^{q_{\alpha}}\} = \sup{B(x)} \sup{B(y)}
      \end{equation}

    \item To prove least, assume there exists $K \in \mathbb{R}$ s.t. $\sup{B^\prime(x + y)} \leq K < \sup{B(x)} \sup{B(y)}$. Then, since the image of $b^x$ is always positive, we assume $0 < K$. We bound its factors as so: $K < \sup{B(x)} \sup{B(y)} \implies K/\sup{B(x)} < \sup{B(y)}$. By density of the rationals, there exists a $\beta \in \mathbb{Q}$, s.t. 
    \begin{equation}
      \frac{K}{\sup{B(x)}} < \beta < \sup{B(y)}
    \end{equation}
    This means $K/\beta < \sup{B(x)}$ and $\beta < \sup{B(y)}$. But this means that there exists $\phi, \gamma \in B(x), B(y)$ s.t. $K/\beta < \phi, \beta < \gamma \implies K = (K/\beta) \cdot \beta < \phi \gamma \implies \phi \gamma \in B^\prime(x + y)$ by definition. So $K$ is not an upper bound. 
    \end{enumerate}
  \end{proof}

  Furthermore, this is an isomorphism, and the inverse is defined. Let's define this analytically. 

  \begin{theorem}[Logarithm]
    For $b > 1$ and $y > 0$, there is a unique real number $x$ s.t. $b^x = y$. We claim 
    \begin{equation}
      x = \sup\{ w \in \mathbb{R} \mid b^w < y \}
    \end{equation}
    $x$ is called the \textbf{logarithm of $y$ to the base $b$}. 
  \end{theorem}
  \begin{proof}
    We use the inequality $b^n - 1 \leq n (b-1)$ for all $n \in \mathbb{N}$.\footnote{We prove by induction. For $n=1$ $b^1 - 1 \leq 1 (b-1)$. Assume that this holds for some $n$. Then $b^{n+1} - 1 = b^{n+1} - b + b - 1 = b (b^n - 1) + (b-1) \geq bn (b-1) + (b-1) = (bn + 1) (b-1) \geq (n+1) (b-1)$, where the last step follows since $b \geq 1 \implies bn \geq n \implies bn + 1 \geq n + 1$. } By substituting $b = b^{1/n}$ (valid since $b > 1 \iff b^{1/n} > 1$) so $b-1 \geq n(b^{1/n} - 1)$. Now set some $t > 1$, and by Archimidean principle, we can choose some $n \in \mathbb{N}$ s.t. $n > \frac{b-1}{t-1}$. Then $n (t-1) > b-1$, and with the inequality derived we get 
    \begin{equation}
      n (t - 1) > b - 1 \geq n (b^{1/n} - 1) \implies t > b^{1/n}
    \end{equation} 
    This allows us to prove 2 things. 
    \begin{enumerate}
      \item If $w$ satisfies $b^w < y$, then $b^{w + (1/n)} < y$ for sufficiently large $n$. Setting $t = y b^{-w}$ (which is greater than $1$ since $b^w < y$) gives $y \cdot b^{-w} > b^{1/n} \implies b^w b^{1/n} < y \implies b^{w + (1/n)} < y$. 
      \item If $w$ satisfies $b^w > y$, then $b^{w - (1/n)} > y$ for sufficiently large $n$. Setting $t = b^w y^{-1}$ (which is greater than $1$ since $b^w > y$) gives $b^w y^{-1} > b^{1/n} \implies b^{w - (1/n)} > y$. 
    \end{enumerate}
    Now we can prove existence. Let $A$ the set of all $w$ s.t. $b^w < y$. We claim that $x = \sup{A}$. 
    \begin{enumerate}
      \item Assume that $b^x < y$. We know that there exists $n \in \mathbb{N}$ s.t. $b^{x + (1/n)} < y \implies x + (1/n) \in A$, contradicting that $x$ is an upper bound. 
      \item Assume that $b^x > $. We know that there exists $n \in \mathbb{N}$ s.t. $b^{x - (1/n)} > y \implies x - (1/n)$ is also an upper bound for $A$, contradicting that $x$ is least. Therefore $b^x = y$. 
    \end{enumerate}
    We now prove uniqueness. Assume that there are two such $x$'s , call them $x, x^\prime$. By total ordering and $x \neq x^\prime$, WLOG let $x > x^\prime \implies x - x^\prime > 0 \implies b^{x - x^\prime} > 1$. By density of rationals, since we can choose $r \in \mathbb{R}$ s.t. $0 < r < x - x^\prime$, we have $1 < b^r < b^{x - x^\prime}$ and so $B(r) \subset B(x - x^\prime)$. Since $1 < b^{x - x^\prime} \implies 1 \cdot b^{x^\prime} < b^{x - x^\prime} \cdot b^{x^\prime} = b^x$, we have $b^{x^\prime} < b^x$ and they cannot both by $y$. So $x = x^\prime$. 
  \end{proof}

\subsection{Extended Reals} 

  Often, we deal with numbers that are not finite, and we would like to have a system to incorporate $\pm \infty$ into the real line. Most first courses glaze over this, but it's important to see the construction as well. The problem is that we can't really add in these numbers without breaking a lot of the algebraic properties, but let's see for ourselves. It should be pretty obvious that we want (note the strict inequalities)
  \begin{equation}
    -\infty < x < +\infty \quad \forall x \in \mathbb{R}
  \end{equation} 
  To define addition, we can't make $x + \infty$ a finite number since then 
  \begin{equation}
    \infty \leq x + \infty = y 
  \end{equation}
  which is a contradiction. So $x + \infty = +\infty$. We keep doing this but the main problem comes in with trying to define $\infty - \infty$ or $\infty/\infty$, which are known as \textbf{indeterminate forms}. These are particularly bad since we cannot deduce $x = y$ from $x + \infty = y + \infty$ or from $x \cdot \infty = y \cdot \infty$. The solution to this is to \textit{simply avoid them}\footnote{as far as I know} by making these indeterminate terms undefined. 

  \begin{definition}[Extended Real Number Line]
    The \textbf{extended real number line} is the set $\mathbb{R} \cup \{\pm \infty\}$ with the following operations. 
    \begin{enumerate}
      \item \textit{Order}. $-\infty \leq x$ and $x \leq +\infty$ for all $x \in \overline{\mathbb{R}}$. 
      \item \textit{Addition}. 
        \begin{align}
          \forall x \in \mathbb{R}, & x + \infty = \infty + x = +\infty \\
          \forall x \in \mathbb{R}, & x - \infty = \infty - x = -\infty \\ 
          & + \infty + \infty = +\infty \\
          & - \infty - \infty = -\infty \\ 
          & +\infty - \infty, -\infty + \infty \text{ are undefined}
        \end{align}
      \item \textit{Multiplication}.\footnote{The fact that $0 \cdot \infty = 0$ might sound odd. Look at the extension into hyperreals later.} 
      \begin{align} 
        \forall x \in \mathbb{R} \setminus \{0\}, & x \cdot +\infty = +\infty \cdot x = \begin{cases} +\infty \text{ if } x > 0 \\ -\infty \text{ if } x < 0 \end{cases} \\
        \forall x \in \mathbb{R} \setminus \{0\}, & x \cdot -\infty = -\infty \cdot x = \begin{cases} -\infty \text{ if } x < 0 \\ -\infty \text{ if } x > 0 \end{cases} \\
        & 0 \cdot +\infty = +\infty \cdot 0 = 0 \\
        & 0 \cdot -\infty = -\infty \cdot 0 = 0 \\
        & +\infty \cdot +\infty = -\infty \cdot -\infty = +\infty \\ 
        & +\infty \cdot -\infty = -\infty \cdot +\infty = +\infty \\ 
      \end{align}
    \end{enumerate}
  \end{definition} 

  It turns out that this is still Dedekind-complete, which is nice. Unfortunately this is not even a field since the multiplicative inverse of $\pm \infty$ is not defined. Furthermore, we lose the Archimedean property. 

  The general rule of thumb is that if one wishes to use cancellation, this is only safe if one can guarantee that the numbers we work with are all finite. If we must work with infinity, another way is to work with the nonnegative reals. 

  \begin{definition}[Extended Real Number Line]
    The \textbf{extended nonnegative reals} is the set $\mathbb{R}_{\geq 0} \cup \{+\infty\}$ with the following operations. 
    \begin{enumerate}
      \item \textit{Order}. $x \leq +\infty$ for all $x \in \overline{\mathbb{R}}$. 
      \item \textit{Addition}. 
      \begin{align}
        \forall x \in [0, +\infty], +\infty + x = x + \infty = +\infty 
      \end{align}
    \item \textit{Multiplication}.
      \begin{align}
        \forall x \in (0, +\infty], & +\infty \cdot x = x \cdot +\infty = +\infty \\ 
                                    & 0 \cdot +\infty = +\infty \cdot 0 = 0 
      \end{align}
    \end{enumerate}
  \end{definition} 

  There is a tradeoff here: we can work with infinity, or negative numbers, but not both. Also, note that if we define $\infty \cdot 0 = 0$, the multiplication becomes \textit{upward continuous}, but not \textit{downwards continuous}. This leads to an asymmetry when defining integrals, but in univariate analysis we will only work with bounded functions, and this will not hinder us until measure theory. 

\subsection{Hyperreals}

  The loss of the field property of the extended reals is quite bad, and we might want to recover this. Therefore, we can add more elements that serve to be the multiplicative inverse of infinity. We call these inverses \textit{infinitesimals} and the new set the \textit{hyperreal numbers}. 

  \begin{theorem}[Hyperreals]
    The \textbf{hyperreals} 
  \end{theorem}

  In fact, when Newton first invented calculus, the hyperreals were what he worked with, and you can surprisingly build a good chunk of calculus with this. Even though this is a dead topic at this point, a lot of modern notation is based off of this number system, so it's good to see how it works. For example, when we write the integral 
  \begin{equation}
    \int f(x) \,dx
  \end{equation} 
  we are saying that we are taking the uncountable sum of the terms $f(x) \,dx$, the multiplication of the real number $f(x)$ and the infinitesimal number $dx$ living in the hyperreals. Unfortunately, we cannot fully construct a rigorous theory of calculus with only infinitesimals. However, in practice (especially physics) people tend to manipulate and do algebra with infinitesimals, so having a good foundation on what you can and cannot do with them is practical. While the focus won't be on \textit{smooth infinitesimal analysis (SIA)}, I will include some alternate constructions later purely with infinitesimals. 

\subsection{Some Algebraic Inequalities}

  We also introduce various inequalities that may be useful for producing future results. The following lemmas can be proved with elementary algebra on the field of reals. 

  \begin{lemma}[Bernoulli's Inequality]
    \label{thm:bernoullis-inequality}
    For any real $x \geq -1$ and $n \in \mathbb{N}$, we have 
    \begin{equation}
      (1 + x)^n \geq 1 + nx
    \end{equation}
  \end{lemma}
  \begin{proof}
    We prove by induction. For $n = 1$, it is trivial. Now given that the inequality is satisfied for some $n \in \mathbb{N}$, we have 
    \begin{align}
      (1 + x)^{n + 1} & = (1 + x)^n (1 + x) \\ 
                      & \geq (1 + nx) (1 + x) \\ 
                      & = 1 + nx + x + nx^2 \\ 
                      & = 1 + (n+1)x
    \end{align}
  \end{proof}

  \begin{lemma}[Young's Inequalities]
    If $a>0$ and $b>0$, and the numbers $p$ and $p$ are such that $p \neq 0, 1$ and $q \neq 0, 1$, and $\frac{1}{p} + \frac{1}{q} = 1$, then 
    \begin{align}
      a^{\frac{1}{p}} b^{\frac{1}{q}} \leq \frac{1}{p} a + \frac{1}{q} b \text{  if } p > 1 \\
      a^{\frac{1}{p}} b^{\frac{1}{q}} \geq \frac{1}{p} a + \frac{1}{q} b \text{  if } p < 1
    \end{align}
    and equality holds in both cases if and only if $a = b$. 
  \end{lemma}
  \begin{proof}
    
  \end{proof}

  \begin{lemma}[Holder's Inequalities]
    Let $x_i \geq 0, y_i \geq 0$ for $i = 1, 2, ..., n$, and let $\frac{1}{p} + \frac{1}{q} = 1$. Then, 
    \begin{align}
      &\sum_{i=1}^n x_i y_i \leq \bigg( \sum_{i=1} x_i^p \bigg)^{\frac{1}{p}} \, \bigg( \sum_{i=1} y_i^q \bigg)^{\frac{1}{q}} \text{  for } p > 1 \\
      &\sum_{i=1}^n x_i y_i \geq \bigg( \sum_{i=1} x_i^p \bigg)^{\frac{1}{p}} \, \bigg( \sum_{i=1} y_i^q \bigg)^{\frac{1}{q}} \text{  for } p < 1, p \neq 0
    \end{align}
  \end{lemma}
  \begin{proof}
    
  \end{proof}

  \begin{lemma}[Minkowski's Inequalities]
    Let $x_i \geq 0, y_i \geq 0$ for $i = 1, 2, ... ,n$. Then, 
    \begin{align}
      \bigg( \sum_{i=1}^n (x_i + y_i)^p \bigg)^{\frac{1}{p}} & \leq \bigg( \sum_{i=1}^n x_i^p \bigg)^\frac{1}{p} + \bigg( \sum_{i=1}^n y_i^p \bigg)^{\frac{1}{p}} \text{  when } p > 1 \\
      \bigg( \sum_{i=1}^n (x_i + y_i)^p \bigg)^{\frac{1}{p}} & \geq \bigg( \sum_{i=1}^n x_i^p \bigg)^\frac{1}{p} + \bigg( \sum_{i=1}^n y_i^p \bigg)^{\frac{1}{p}} \text{  when } p < 1, p \neq 0
    \end{align}
  \end{lemma}
  \begin{proof}
    
  \end{proof}

\subsection{Exercises}

  \begin{exercise}[Math 531 Spring 2025, PS2.1]
    Prove that the set of all matrices of the form:
    \begin{equation}
      \begin{bmatrix}
        a & -b \\
        b & a
      \end{bmatrix},
    \end{equation}
    with $a,b \in \mathbb{R}$ forms a field with the usual sum and product operations of
    matrices. What does this field resemble? Give extensions to $3 \times 3$ and
    $4 \times 4$ matrices.
  \end{exercise}
  \begin{solution}
    
  \end{solution}

  \begin{exercise}[Math 531 Spring 2025, PS2.2]
    Why can't the field of complex numbers (with its usual operations) be
    made into an ordered field?
  \end{exercise}
  \begin{solution}
    Solution is shown as theorem. 
  \end{solution}

  \begin{exercise}[Math 531 Spring 2025, PS2.3]
    Prove there are no finite ordered fields.
  \end{exercise}
  \begin{solution}
    Solution is shown as theorem. 
  \end{solution}

  \begin{exercise}[Math 531 Spring 2025, PS2.4]
    Prove that if $x$ and $n$ are natural numbers, then
    \begin{equation}
      x^n - 1 = (x-1)(1 + x + x^2 + ... + x^{n-1}).
    \end{equation}
  \end{exercise}
  \begin{solution}
    We use the commutative addition and multiplication, plus distributive property in $\mathbb{Z}$. 
    \begin{align}
      (x - 1) \bigg( \sum_{i=0}^{n-1} x^i \bigg) & = x \sum_{i=0}^{n-1} x^i - \sum_{i=0}^{n-1} x_i \\
                                                 & = \sum_{i=1}^n x^i - \sum_{i=0}^{n-1} x^i \\
                                                 & = x^n + \sum_{i=1}^{n-1} x^i - \sum_{i=1}^{n-1} x^i - 1 \\
                                                 & = x^n - 1
    \end{align}
  \end{solution}

  \begin{exercise}[Math 531 Spring 2025, PS2.7]
    Prove that there is no $q \in \mathbb{Q}$ for which
    \begin{equation}
      q^2 + q = 4.
    \end{equation}
  \end{exercise}
  \begin{solution}
    Assume that such a $q \in \mathbb{Q}$ in canonical form exists. Then by the field properties, since $\frac{1}{4} \in \mathbb{Q}$, 
    \begin{equation}
      q^2 + q + \frac{1}{4} = 4 + \frac{1}{4} = \frac{17}{4} \in \mathbb{Q}
    \end{equation}
    But by distributive properties, $(q + \frac{1}{2})^2 \in \mathbb{Q}$. We claim that there exists no rational $x = a/b$ ($a, b$ coprime) s.t. $x^2 = 17/4$. If there were, then clearly $a \neq 0$ and 
    \begin{align}
      \frac{a^2}{b^2} = \frac{17}{4} & \implies 4a^2 = 17 b^2 \\
                                     & \implies 2 | b, \text{ and so } b = 2b^\prime \text{ for some } b^\prime \in \mathbb{N} \\
                                     & \implies a^2 = 17 (b^\prime)^2 \\ 
                                     & \implies 17 | a \text{ and so } a = 17 a^\prime \text{ for some } a^\prime \in \mathbb{Z} \\
                                     & \implies 17 (a^\prime)^2 = (b^\prime)^2
    \end{align}
    which implies that $17|b^\prime$, but this contradicts the assumption that $a, b$ are coprime. Therefore $q + \frac{1}{2} \not\in \mathbb{Q} \implies q \not\in \mathbb{Q}$. 
  \end{solution}

  \begin{exercise}[Math 531 Spring 2025, PS2.8]
    Let $X$ be an ordered set with the least upper bound property. Prove that $X$ has the greatest lower bound property.
  \end{exercise}
  \begin{solution}
    Shown in theorem above. 
  \end{solution}

  \begin{exercise}[Math 531 Spring 2025, PS2.9]
    Prove that if $x, y \in \mathbb{Q}$ we have that
    \begin{equation}
      ||x| - |y|| \leq |x - y|.
    \end{equation}
  \end{exercise}
  \begin{solution}
    By subadditivity of the norm we have 
    \begin{align}
      |x| \leq |x - y| + |y| & \implies |x| - |y| \leq |x - y| \\
      |y| \leq |y - x| + |x| & \implies |y| - |x| \leq |y - x| 
    \end{align}
    But $|y - x| = |-1 (x - y)| = |-1| \cdot |x - y| = |x - y|$, and so 
    \begin{equation}
      \max\{ |x| - |y|, |y| - |x| \} \leq |x - y|
    \end{equation}
    and the LHS is the definition of the norm $||x| - |y||$ in $\mathbb{Q}$. 
  \end{solution}

  \begin{exercise}[Rudin 1.1]
    If $r$ is rational ($r \neq 0$) and $x$ is irrational, prove that $r + x$ and $rx$ are irrational. 
  \end{exercise}
  \begin{solution}
    If we assume that $r x = t$ and $r + x = s$ are rational, then this violates the field axioms of $\mathbb{Q}$ since then $x = t r^{-1}$ and $x = s + (-r)$ are rational. 
  \end{solution}

  \begin{exercise}[Rudin 1.2]
    Prove that there is no rational number whose square is $12$. 
  \end{exercise}
  \begin{solution}
    Assume that there exists a number $p/q$ such that $p$ and $q$ are both not even. Then, 
    \begin{equation}
      \bigg( \frac{p}{q} \bigg)^2 = 12 \implies p^2 = 12q^2 = 3 (2 q)^2
    \end{equation}
    So $p$ much be even $p = 2 p^\prime$. Therefore, $p^{\prime 2} = 3 q^2$, and $q$ must be odd. This means that $p^\prime$ must be odd. We can rewrite the equation 
    \begin{equation}
      p^{\prime 2} - q^2 = 2 q^2 \implies (p^\prime + q) (p^\prime - q) = 2q^2
    \end{equation}
    where the left hand side is divisible by $4$ but the right hand side is divisible by at most $2$, leading to a contradiction. 
  \end{solution}

  \begin{exercise}[Rudin 1.3]
    Prove that the axioms of multiplication imply the following. 
    \begin{enumerate}
      \item If $x \neq 0$ and $xy = xz$, then $y = z$. 
      \item If $x \neq 0$ and $xy = x$, then $y = 1$. 
      \item If $x \neq 0$ and $xy = 1$, then $y = x^{-1}$. 
      \item If $x \neq 0$, then $(x^{-1})^{-1} = x$. 
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Listed. 
    \begin{enumerate}
      \item $xy = xz \implies \frac{1}{x} \cdot x y = \frac{1}{x} x z \implies y = z$ 
      \item $x y = x \implies \frac{1}{x} x y = \frac{1}{x} x \implies y = 1$ 
      \item $x y = 1 \implies \frac{1}{x} x y = \frac{1}{x} 1 \implies y = \frac{1}{x}$ 
      \item $(x^{-1})^{-1} \cdot x^{-1} = 1 \implies (x^{-1})^{-1} \cdot x^{-1} \cdot x = 1 \cdot x \implies (x^{-1})^{-1} = x$
    \end{enumerate}
  \end{solution}

  \begin{exercise}[Rudin 1.4]
    Let $E$ be a nonempty subset of an ordered set; suppose $\alpha$ is a lower bound of $E$ and $\beta$ is an upper bound of $E$. Prove that $\alpha \leq \beta$. 
  \end{exercise}
  \begin{solution}
    Since $E$ is nonempty, we choose any $x \in E$. By definition, $\alpha \leq x$ and $x \leq \beta$, and by transitive property of orderings, we have $\alpha \leq \beta$. 
  \end{solution}

  \begin{exercise}[Rudin 1.5]
    Let $A$ be a nonempty set of real numbers which is bounded below. Let $-A$ be the set of all numbers $-x$, where $x \in A$. Prove that 
    \begin{equation}
      \inf A = -\sup(-A)
    \end{equation}
  \end{exercise}
  \begin{solution}
    We would like to prove that $\inf A \leq -\sup(-A)$ and $\inf A \geq -\sup(-A)$. For the first part, we start off with the definition of the infimum. 
    \begin{align*}
      \inf A \leq x \; \forall x \in A & \implies - \inf A \geq -x \; \forall x \in A \\
      & \implies - \inf A \geq x \forall x \in -A \\
      & \implies -\inf A \geq \sup(-A) \\
      & \implies \inf A \leq - \sup(-A)
    \end{align*}
    For the second part, we start with the definition of the supremum. 
    \begin{align*}
      \sup(-A) \geq x \forall x \in -A & \implies \sup(-A) \geq -x \forall x \in A \\
      & \implies -\sup(-A) \geq x \forall x \in A \\
      & \implies -\sup(-A) \leq \inf A
    \end{align*}
  \end{solution}

  \begin{exercise}[Rudin 1.6]
    Fix $b > 1$.
    \begin{enumerate}
      \item If $m$, $n$, $p$, $q$ are integers, $n > 0$, $q > 0$, and $r = m/n = p/q$, prove that
      \begin{equation}
        (b^m)^{1/n} = (b^p)^{1/q}.
      \end{equation}
      Hence it makes sense to define $b^r = (b^m)^{1/n}$.
      
      \item Prove that $b^{r+s} = b^rb^s$ if $r$ and $s$ are rational.
      
      \item If $x$ is real, define $B(x)$ to be the set of all numbers $b^t$, where $t$ is rational and $t \leq x$. Prove that 
      \begin{equation}
        b^r = \sup B(r)
      \end{equation}
      when $r$ is rational. Hence it makes sense to define
      \begin{equation}
        b^x = \sup B(x)
      \end{equation}
      for every real $x$.
      
      \item Prove that $b^{x+y} = b^xb^y$ for all real $x$ and $y$.
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Proved in theorem above. 
  \end{solution}

  \begin{exercise}[Rudin 1.7]
    Fix $b > 1$, $y > 0$, and prove that there is a unique real $x$ such that $b^x = y$, by
    completing the following outline. (This $x$ is called the logarithm of $y$ to the base $b$.)
    \begin{enumerate}
      \item For any positive integer $n$, $b^n - 1 \geq n(b-1)$.
      \item Hence $b - 1 \geq n(b^{1/n} - 1)$.
      \item If $t > 1$ and $n > (b-1)/(t-1)$, then $b^{1/n} < t$.
      \item If $w$ is such that $b^w > y$, then $b^{-(1/n)} < y$ for sufficiently large $n$; to see this,
        apply part (c) with $t = y \cdot b^{-w}$.
      \item If $b^w > y$, then $b^{w-(1/n)} > y$ for sufficiently large $n$.
      \item Let $A$ be the set of all $w$ such that $b^w < y$, and show that $x = \sup A$ satisfies
        $b^x = y$.
      \item Prove that this $x$ is unique.
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Proved in theorem above. 
  \end{solution}

  \begin{exercise}[Rudin 1.8]
    Prove that no order can be defined in the complex field that turns it into an ordered field. 
  \end{exercise}
  \begin{solution}
    Note that if $x \geq 0$, then $-x \leq 0$ for all $x$ of any ordered field. Since if $x \geq 0$ and $-x > 0$, then $x -x > 0$, which is absurd. Therefore, one of either $i$ or $-i$ should be greater than $0$. But $i^2 = (-i)^2 = -1$, so this means that $-1 > 0$, which implies that $0 < 1$. But either $1$ or $-1$ must $\geq 0$. 
  \end{solution}

  \begin{exercise}[Rudin 1.9]
    Equip $\mathbb{C}$ with the dictionary order. That is, given $z = a + bi$ and $w = c + di$, $z < w$ if $a < c$, or if $a = c$ and $b < d$. Does this ordered set have a least upper bound property? 
  \end{exercise}
  \begin{solution}
    No it does not. Consider the set $S = \{ a + b i \in \mathbb{C} \mid a \leq 3\}$. $S$ is bounded by $4$, but it doesn't have a least upper bound. Given any $3 + bi$, this is not an upper bound since we can construct $3 + (b + \epsilon) i \in S$. Given any $a + bi$ where $a > 3$, we can always find a lower bound of form $a + (b - \epsilon) i$ that also bounds $S$. 
  \end{solution}

  \begin{exercise}[Rudin 1.10]
    Suppose $z = a + bi$, $w = u + iv$, and 
    \begin{equation}
      a = \bigg( \frac{|w| + u}{2} \bigg)^{1/2} \text{ and } b = \bigg( \frac{|w| - u}{2} \bigg)^{1/2}
    \end{equation}
    Prove that $z^2 = w$ if $v \geq 0$ and that $(\bar{z})^2 = w$ if $v \leq 0$. Conclude that every complex number (with one exception!) has two complex square roots. 
  \end{exercise}
  \begin{solution}
    We can calculate 
    \begin{equation}
      z^2 = (a^2 - b^2) + 2 a b i = u + \sqrt{v^2} i = \begin{cases} u + v i & \text{ if } v \geq 0 \\ u - vi & \text{ if } v \leq 0 \end{cases} 
    \end{equation}
    Since if we assume $v \geq 0$, then we have $z^2 = w$. We also get 
    \begin{equation}
      \bar{z}^2 = (a^2 - b^2) - 2 a b i = u - \sqrt{v^2} i = \begin{cases} u - v i & \text{ if } v \geq 0 \\ u + vi & \text{ if } v \leq 0 \end{cases}
    \end{equation}
    and assuming $v \leq 0$, we have $\bar{z}^2 = w$. Therefore, every complex number $w$ has both $\pm z$ as its square root if $v \geq 0$, $\pm \bar{z}$ if $v \leq 0$, and just one root if $z = 0$. 
  \end{solution}

  \begin{exercise}[Rudin 1.11]
    If $z$ is a complex number, prove that there exists an $r \geq 0$ and a complex number $w$ with $|w| = 1$ s.t. $z = rw$. Are $w$ and $r$ always uniquely determined by $z$? 
  \end{exercise}
  \begin{solution}
    If $z = 0$, then $r = 0$ and there is no unique $w$. If $z = a + bi \neq 0$, then define 
    \begin{equation}
      r = |z| = (a^2 + b^2)^{1/2}, \;\; w = \frac{1}{r} z
    \end{equation}
    which proves existence. As for uniqueness, assume that there are two forms 
    \begin{equation}
      z = r w = r^\prime w^\prime
    \end{equation}
    Then, $w = \frac{r^\prime}{r} w^\prime \implies |w| = \big| \frac{r^\prime}{r} \big| |w^\prime| = 1$, which implies that $r^\prime / r = 1$ and so $r = r^\prime$. This means that $w = w^\prime$. 
  \end{solution}

  \begin{exercise}[Rudin 1.12]
    If $z_1, \ldots, z_n$ are complex, prove that
    \begin{equation}
      |z_1 + z_2 + \ldots + z_n| \leq |z_1| + \ldots + |z_n|
    \end{equation}
  \end{exercise}
  \begin{solution}
    By induction, it suffices to prove $|z_1 + z_2| \leq |z_1| + |z_2|$. We have 
    \begin{align*}
      |z_1 + z_2|^2 & = (z_1 + z_2) (\overline{z_1 + z_2}) \\
      & = (z_1 + z_2) (\bar{z_1} + \bar{z_2}) \\ 
      & = z_1 \bar{z_1} + z_1 \bar{z_2} + z_2 \bar{z_1} + z_2 \bar{z_2} \\
      & = |z_1|^2 + |z_2|^2 + z_1 \bar{z_2} + z_2 \bar{z_1} \\
      & = |z_1|^2 + |z_2|^2 + 2 (a c + bd) \\
      & \leq |z_1|^2 + |z_2|^2 + 2 \sqrt{a^2 + b^2} \sqrt{c^2 + d^2} \;\; (\text{Schwartz}) \\
      & = |z_1|^2 + |z_2|^2 + 2 |z_1| |z_2| \\
      & = (|z_1| + |z_2|)^2
    \end{align*}
    since both sides are positive, we can take their square root to get the desired result. 
  \end{solution}

  \begin{exercise}[Rudin 1.13]
    If $x, y$ are complex, prove that 
    \begin{equation}
      \big| |x| - |y| \big| \leq |x - y|
    \end{equation}
  \end{exercise}
  \begin{solution}
    Since both sides are nonnegative, we can square both sides. Note that due to Cauchy Schwartz inequality, $2|x| |y| \geq x \bar{y} + y \bar{x}$ since expanding them gives 
    \begin{equation}
      2 \sqrt{a^2 + b^2} \sqrt{c^2 + d^2} \geq 2 (ac + bd)
    \end{equation}
    Therefore, the following inequality is true: 
    \begin{equation}
      |x|^2 + |y|^2 - 2 |x| |y| \leq x \bar{x} + y \bar{y} - x \bar{y} - y \bar{x}
    \end{equation}
    which reduces to form $(|x| - |y|)^2 \leq |x - y|^2$. 
  \end{solution}

  \begin{exercise}[Rudin 1.14]
    If $z$ is a complex number s.t. $|z| = 1$, that is such that $z \bar{z} = 1$, compute 
    \begin{equation}
      |1 + z|^2 + |1 - z|^2
    \end{equation}
  \end{exercise}
  \begin{solution}
    Compute. 
    \begin{equation}
      (1 + z) (1 + \bar{z}) + (1 - z) (1 - \bar{z}) = 1 + z + \bar{z} + z \bar{z} + 1 - z - \bar{z} + z \bar{z} = 4
    \end{equation}
  \end{solution}

  \begin{exercise}[Rudin 1.15]
    Under what conditions does equality hold in the Schwarz inequality? 
  \end{exercise}
  \begin{solution}
    If they are antiparallel, since 
    \begin{equation}
      \langle x, y \rangle = ||x || ||y || \cos{\theta}
    \end{equation}
  \end{solution}

  \begin{exercise}[Rudin 1.16]
    Suppose $k \geq 3$, $\mathbf{x}, \mathbf{y} \in \mathbb{R}^k$, $|\mathbf{x} - \mathbf{y}| = d > 0$, and $r > 0$. Prove: 
    \begin{enumerate}
      \item[a)] If $2r > d$, there are infinitely many $\mathbf{z} \in \mathbb{R}^k$ s.t. 
        \begin{equation}
        |\mathbf{z} - \mathbf{x}| = |\mathbf{z} - \mathbf{y}| = r
        \end{equation}
      \item[b)] If $2r = d$, there is exactly one such $\mathbf{z}$. 
      \item[c)] If $2r < d$, there is no such $\mathbf{z}$. 
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    
  \end{solution}

  \begin{exercise}[Rudin 1.17]
    Prove that
    \begin{equation}
      |\mathbf{x} + \mathbf{y}|^2 + |\mathbf{x} - \mathbf{y}|^2 = 2|\mathbf{x}|^2 + 2|\mathbf{y}|^2
    \end{equation}
  \end{exercise}
  \begin{solution}
    This is trivial if we simply expand 
    \begin{align}
      |\mathbf{x} + \mathbf{y}|^2 + |\mathbf{x} - \mathbf{y}|^2 & = \langle \mathbf{x} + \mathbf{y}, \mathbf{x} + \mathbf{y} \rangle + \langle \mathbf{x} - \mathbf{y}, \mathbf{x} - \mathbf{y} \rangle \\
      & = \langle \mathbf{x}, \mathbf{x} \rangle + 2 \langle \mathbf{x}, \mathbf{y} \rangle + \langle \mathbf{y}, \mathbf{y} \rangle + \langle \mathbf{x}, \mathbf{x} \rangle - 2 \langle \mathbf{x}, \mathbf{y} \rangle + \langle \mathbf{y}, \mathbf{y} \rangle \\
      & = 2 \langle \mathbf{x}, \mathbf{x} \rangle + 2 \langle \mathbf{y}, \mathbf{y} \rangle \\
      & = 2|\mathbf{x}|^2 + 2|\mathbf{y}|^2
    \end{align}
  \end{solution}

  \begin{exercise}[Rudin 1.18]
    If $k \geq 2$ and $\mathbf{x} \in \mathbb{R}^k$, prove that there exists $\mathbf{y} \in \mathbb{R}^k$ s.t. $\mathbf{y} \neq \mathbf{0}$, but $\mathbf{x} \cdot \mathbf{y} = 0$. Is this also true if $k = 1$? 
  \end{exercise}
  \begin{solution}
    Let $x \in \mathbb{R}^k$ and $\ell \in \mathbb{R}^{k \ast}$, the dual space. By Riesz representation theorem, we can define the canonical isomorphism $\ell \mapsto y$ between these two spaces as 
    \begin{equation}
      \ell (x) = (x, y)
    \end{equation}
    Since $y \neq 0$ by assumption, $\ell \neq 0$, and so its rank is at least $1$. Since $\ell$ maps to $\mathbb{R}$, the rank has to be $1$. By rank nullity theorem, we have 
    \begin{equation}
      \dim N(\ell) = k - \mathrm{rank}(\ell) = k - 1
    \end{equation}
    and so there exists nontrivial annihilators $\ell$ of $x$, which can be mapped to a nontrivial $y \in \mathbb{R}^k$. 
  \end{solution}

  \begin{exercise}[Rudin 1.19]
    Suppose $\mathbf{a}, \mathbf{b} \in \mathbb{R}^k$. Find $\mathbf{c} \in \mathbb{R}^k$ and $r > 0$ s.t. 
    \begin{equation}
      |\mathbf{x} - \mathbf{a}| = 2 | \mathbf{x} - \mathbf{b}|
    \end{equation}
    if and only if $|\mathbf{x} - \mathbf{c}| = r$. 
  \end{exercise}
  \begin{solution}
    If we draw out the circle, it must contain two points on the line drawn by connecting $A$ and $B$. Since it must be symmetric, its center and radius can then be easily calculated to be 
    \begin{equation}
      r = \frac{2}{3} |b - a|, \;\;\; c = \frac{1}{3} (4b - a)
    \end{equation}
  \end{solution}

  \begin{exercise}[Zorich 2.2.1]
    Using the principle of induction, show that 
    \begin{enumerate}
      \item the sum $x_1 + \ldots + x_n$ of real numbers is defined independently of the insertion of parentheses to specify the order of addition. 
      \item the same is true of the product $x_1 \ldots x_n$ 
      \item $|x_1 + \ldots + x_n| \leq |x_1| + \ldots + |x_n|$ 
      \item $|x_1 \ldots x_n| \leq |x_1| \ldots |x_n|$
      \item For any $n, m \in \mathbb{N}$ such that $m < n$, $(n - m) \in \mathbb{N}$. 
      \item $(1 + x)^n \geq 1 + n x$ for $x > - 1$ and $n \in \mathbb{N}$, equality holding for when $n=1$ or $x=0$. 
      \item $(a + b)^n = a^n + _n C_1 a^{n-1} b^1 + \ldots + b^n$ (aka binomial theorem). 
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Listed. 
    \begin{enumerate}
      \item Let $n$ denote the number of elements in the sum. We prove by strong law of induction. The base case for when $n=1, 2, 3$ is trivially true. 
      \begin{align*}
        x_1 & = x_1 & (\text{identity}) \\
        x_1 + x_2 & = x_1 + x_2 & (\text{identity}) \\
        (x_1 + x_2) + x_3 & = x_1 + (x_2 + x_3) & (\text{associativity}) 
      \end{align*}
      Then, the sum of $n=k$ parameters is defined by $k-2$ pairs of parentheses defining the order of the sum. These parentheses define a sequence of $k-1$ 2-fold additions. Now, assume that the claim is true for 
      \begin{equation}
        S_n \equiv x_1 + \ldots x_n \text{ for } n = 1, 2, \ldots, k
      \end{equation}
      Then, for a specific sum $S_{k+1}$ of $k+1$ elements with $k-1$ parentheses, we can reduce the sum to its final 2-fold addition 
      \begin{equation}
        S_{k+1} \equiv \underbrace{(x_1 + \ldots + x_i)}_{\varphi_1}  + \underbrace{(x_{i+1} + \ldots + x_{k+1})}_{\varphi_2}
      \end{equation}
      Since $i, k-i+1 < k$, by the strong law $\varphi_1$ and $\varphi_2$ are independent of the order of sum. 
      \item Exactly identical to (a). 
      \item By the triangle inequality $|x_1 + x+2| \leq |x_1| + |x_2|$. Now, assume for $n=k$ is true. Then, let $S_k = x_1 + \ldots + x_k$, so 
      \begin{equation}
        |x_1 + \ldots + x_k + x_{k+1}| = |S_k + x_{k+1}| \leq |S_k| + |x_{k+1}| \leq \sum_{i=1}^{k+1} |x_{i}|
      \end{equation}
      \item Same as (c). 
      \item Let us fix $m$ to be any element of $\mathbb{N}$. Then, the base case is for $n = m + 1$ (which is in $\mathbb{N}$ since it is inductive), so 
      \begin{equation}
        n - m = (m + 1) - m = 1 \in \mathbb{N}
      \end{equation}
      Now, given that for some integer $n \geq m+1$, $n - m \in \mathbb{N}$ is true, we have 
      \begin{align*}
        (n + 1) - m & = n + (1 - m ) & (\text{associativity}) \\
        & = n + (-m + 1) & (\text{commutativity}) \\
        & = (n - m) + 1 & (\text{associativity})
      \end{align*}
      where $(n - m) + 1 \in \mathbb{N}$ by inductive property of $\mathbb{N}$. 
      \item We prove by induction. For $n=1$, it is trivial that $(1 + x)^1 \geq 1 + 1 \cdot x$. Now assume that the claim is true for some $k \in \mathbb{N}$. Then, 
      \begin{align*}
        (1 + x)^{k+1} = (1 + x)^k (1 + x) & \geq (1 + k x) (1 + x) \\
        = 1 + (k+1) x + k x^2 \\ 
        \geq 1 + (k+1) x
      \end{align*}
      where equality holds if $x = 0 \implies 1^{k+1} = 1^k \cdot 1 = 1$ or $n=1 \implies $ trivial case. 
      \item The base case for $n=1$ is trivial since $(a + b)^1 = \binom{1}{0} a + \binom{1}{1} b$. We introduce Newton's identity. 
      \begin{align*}
        \binom{k}{j-1} + \binom{k}{j} & = \frac{k!}{(j-1)! (k-j+1)!} + \frac{k!}{j! (k-j)!} \\
        & = k! \bigg( \frac{j}{j! (k-j+1)!} + \frac{k-j+1}{j! (k-j+1)!} \bigg) \\
        & = k! \cdot \frac{k+1}{j! (k-j+1)!} \\
        & = \frac{(k+1)!}{j!(k-j+1)!} = \binom{k+1}{j}
      \end{align*}
      Now assuming that the binomial formula holds for some $n=k$, we have 
      \begin{align}
        (a + b)^{k+1} & = (a + b)^k (a + b) \\
        & = \bigg( \sum_{j=0}^k \binom{k}{j} a^j b^{k-j} \bigg) (a + b) \\
        & = \sum_{j=0}^k \binom{k}{j} a^{j+1} b^{k-j} + \sum_{j=0}^k \binom{k}{j} a^j b^{k-j + 1} \\
        & = \binom{k}{0} a^0 b^{k+1} + \binom{k}{k} a^{j+1} b^0 + \sum_{j=0}^{k-1} \binom{k}{j} a^{j+1} b^{k-j} + \sum_{j=1}^k \binom{k}{j} a^j b^{k-j+1} \\
        & = \binom{k+1}{0} a^0 b^{k+1} + \binom{k+1}{k+1} a^{j+1} b^0 + \sum_{j=1}^k \bigg[ \binom{k}{j-1} + \binom{k}{j} \bigg] \, a^j b^{k-j + 1} \\
        & = \sum_{j=0}^{k+1} \binom{k+1}{j} a^j b^{k-j+1} 
      \end{align}
    \end{enumerate}
  \end{solution}

  \begin{exercise}[Zorich 2.2.3]
    Show that an inductive set is not bounded above. 
  \end{exercise}
  \begin{solution}
    Assume that a $X$ is a nonempty inductive set that is bounded above. By definition, there exists a number $B \in \mathbb{R}$ such that $\max{X} < B$. Then, this means that there exists no numbers in $[B, B + 1)$. Since $X$ is inductive, this means that there cannot exist any elements of $X$ in the interval $[B-1, B)$, and similarly for the interval $[B-2, B)$, and so on, meaning that if $x \in X$, then $x \not\in [B-k, B-k + 1)$ for all $k\in \mathbb{Z}$. By the Archimidean principle, this implies that $X = \emptyset$, contradicting our assumption. 
  \end{solution}

  \begin{exercise}[Zorich 2.2.4]
    Prove the following. 
    \begin{enumerate}
      \item An inductive set is infinite (that is, equipollent with one of its subsets different from itself). 
      \item The set $E_n = \{x \in \mathbb{N}\,|\, x \leq n\}$ is finite. 
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Listed. 
    \begin{enumerate}
      \item Assume that an inductive set $X$ is finite $\implies$ $X$ is bounded above (we can choose upper bound $B = \max{X} + 1$). But from 2.2.3, an inductive set cannot be bounded above, contradicting our assumption. 
      \item It is trivial that $E_1 = \{1\}$ is finite since $\text{card}{E_1} = 1$. Now, if for some $k$, $E_k$ is finite with cardinality $e_k$, then $\text{card}{E_{k+1}} = e_k + 1$, which implies finiteness.  
    \end{enumerate}
  \end{solution}

  \begin{exercise}[Zorich 2.2.5]
    Listed. 
    \begin{enumerate}
      \item Let $m, n \in \mathbb{N}$ and $m >n$. Their greatest common divisor $\gcd(m, n) = d \in \mathbb{N}$ can be found in a finite number of steps using the following algorithm of Euclid involving successive divisions with remainder. 
      \begin{align*}
        m & = q_1 n + r_1 \\
        n & = q_2 r_1 + r_2 \\
        r_1 & = q_3 r_2 + r_3 \\
        \ldots & = \ldots \\
        r_{k-2} & = q_{k} r_{k-1} + r_{k} \\
        r_{k-1} & = q_{k+1} r_k + 0
      \end{align*}
      Then $d=r_k$ 
      \item If $d = \gcd(m, n)$, one can choose numbers $p, q \in \mathbb{Z}$ such that $pm + qn = d$. 
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Listed. 
    \begin{enumerate}
      \item 
      \item Letting $n = r_0$, notice that the equations above satisfy for $i=0, 1, \ldots$
      \begin{equation}
        r_i = q_{i+2} r_{i+1} + r_{i+2} \implies r_{i} - q_{i+2} r_{i+1} = r_{i+2} \tag{1} \label{Euclid}
      \end{equation}
      Note that the second-to-last equation allows us to write $r_k$ as a linear combination of $r_{k-2}$ and $r_{k-1}$: $r_k = r_{k-2} - q_k r_{k-1}$. Now by applying \eqref{Euclid}, we can reduce the above to a linear combination of $r_{k-3}$ and $r_{k-2}$. 
      \begin{align*}
        r_k & = r_{k-2} - q_k r_{k-1} \\
        & = r_{k-2} - q_k (r_{k-3} - q_{k-1} r_{k-3}) \\
        & = (1 + q_{k-1} q_k) r_{k-2} - q_k r_{k-3} 
      \end{align*}
      and repeatedly doing this allows us to reduce $r_k$ to a linear combination $q_0 r_0 + q_1 r_1$. By the ring properties of $\mathbb{Z}$, the new linear coefficients are also in $\mathbb{Z}$. Reducing one last time using the first equation in the Euclidean algorithm gives 
      \begin{align*}
        r_k & = q_0 r_0 + q_1 r_1 \\
        & = q_0 n + q_1 (m - q_1 n) \\
        & = q_1 m + (q_0 - q_1) n \\
        & = p m + q n 
      \end{align*}
    \end{enumerate}
  \end{solution}

  \begin{exercise}[Zorich 2.2.9]
    Show that if the natural number $n$ is not of the form $k^m$, where $k, m \in \mathbb{N}$, then the equation $x^m = n$ has no rational roots. 
  \end{exercise}
  \begin{solution}
    Assume that there is a rational solution $x = p/q$, with $p, q \in \mathbb{N}$ of the equation. Then, 
    \begin{equation}
      \bigg( \frac{p}{q}\bigg)^m = \frac{p^m}{q^m} = n \implies p^m = q^m n
    \end{equation}
    By the fundamental theorem of arithmetic, the exponents of the prime factors of $p^m$ must all be multiples of $m$, and so it must be so for the right hand side $\implies x$ must be of form $x = k^m$ for some $k$. This is a contradiction. 
  \end{solution}

  \begin{exercise}[Zorich 2.2.12]
    Knowing that $\frac{m}{n} \equiv m \cdot n^{-1}$ by definition, where $m \in \mathbb{Z}$ and $n \in \mathbb{N}$, derive the ``rules'' for addition, multiplication, and division of fractions, and also the condition for two fractions to be equal. 
  \end{exercise}
  \begin{solution}
    We can construct a $\mathbb{Q}$ as a quotient space $\mathbb{Z} \times \mathbb{N} / \sim$, where $\sim$ is an equivalent relation where 
    \begin{equation}
      (q_1, p_1) \sim (q_2 , p_2) \text{ iff } q_1 p_2 = p_1 q_2
    \end{equation}
    which is the familiar equivalence relation from ``simplifying'' a fraction. We define addition and multiplication as the following 
    \begin{align*}
      (a, b) + (c, d) & = (ad + bc, bd) \\
      (a, b) \cdot (c, d) & = (ac, bd) 
    \end{align*}
    which turns out to be algebraically closed in $\mathbb{Q}$. The additive identity is the equivalence class $0 = \{(0, c) \,|\, c \in \mathbb{N}\}$, and the multiplicative identity is the equivalence class $1 = \{(c, c) \,|\, c \in \mathbb{N}\}$. It is easy to check that $+$ is commutative, the additive inverse is $-(a, b) = (-a, b)$, and the multiplicative inverse is $(a, b)^{-1} = (b, a)$. We can subtract and divide these elements of $\mathbb{Q}$, called ``fractions,'' as such: 
    \begin{align*}
      (a, b) - (c, d) & = (a, b) + (-(c, d)) = (a, b) + (-c, d) = (ad - bc, bd) \\
      (a, b) \div (c, d) & = (a, b) \cdot (c, d)^{-1} = (a, b) \cdot (d, c) = (ad, bc) 
    \end{align*}
  \end{solution}

  \begin{exercise}[Zorich 2.2.13]
    Verify that the rational numbers $\mathbb{Q}$ satisfy all the axioms for real numbers except for the axiom of completeness. 
  \end{exercise}
  \begin{solution}
    From continuing the steps of 2.2.14, we can prove $\mathbb{Q}$ is an algebraic field (associativity, commutativity of addition and multiplication, along with distributive property).We can actually define the order relation $\leq_\mathbb{Q}$ in two ways: 
    \begin{enumerate}
      \item $(a, b) \leq (c, d)$ iff $ad \leq_{\mathbb{Z}} bc$, where $\leq_{\mathbb{Z}}$ is the order relation on $\mathbb{Z}$ (which can be defined much more simply). 
      \item Recognizing that $\mathbb{Q} \subset \mathbb{R}$, we define the canonical injection map $i: \mathbb{Q} \longrightarrow \mathbb{R}$ and by abuse of language, endow the relation $\leq_{\mathbb{Q}}$ as the restriction of $\leq_{\mathbb{R}}$ onto $\mathbb{Q}$. That is, for $(a, b), (c, d) \in \mathbb{Q}$, 
      \begin{equation}
        (a, b) \leq_{\mathbb{Q}} (c, d) \text{ iff } i(a, b) \leq_{\mathbb{R}} i(c, d)
      \end{equation}
    \end{enumerate}
    The ordering for the 1st step can be checked for consistency. 
    \begin{enumerate}
      \item $(a, b) \leq (a, b)$ since $ab \leq ab$ (true in $\mathbb{Z}$) 
      \item $(a, b) \leq (c, d), (c, d) \leq (a, b)$ means that $ad \leq bc$ and $bc \leq ad \implies ad = bc$ (true in $\mathbb{Z}$) 
      \item $(a, b) \leq (c, d) \leq (e, f)$ implies $ad \leq bc, cf \leq de$. Multiplying positive (important that $f >0$!) to the first inequality gives $adf \leq bcf$, and multiplying positive $b$ to the second gives $bcf \leq bde$, and by interpreting $\leq$ as the ordering defined on $\mathbb{Z}$, we use transitive property of $\leq_\mathbb{Z}$ to get $adf \leq bde \implies af \leq be \iff (a, b) \leq (e, f)$. 
      \item For any $(a, b), (c, d) \in \mathbb{Q}$, $(a, b) \leq (c, d)$ or $(a, b) \geq (c, d)$, which is equivalent to $ad \leq bc$ or $ad \geq bc$, which is true in $\mathbb{Z}$. 
    \end{enumerate}
    It is easy to prove $(a, b) \leq (c, d) \implies (a, b) + (p, q) \leq (c, d) + (p, q)$, and $0_{\mathbb{Q}} \leq (a, b), (c, d) \implies 0_\mathbb{Q} \leq (a, b) \cdot (c, d)$. However, $\mathbb{Q}$ is \textbf{not} complete. We prove this by showing that the subset $X = \{x \in \mathbb{Q} \,|\,x^2 \leq 2 \} \subset \mathbb{Q}$ does not satisfy the least upper bound property. Assume that there is a least upper bound $c \in \mathbb{Q}$. $c \neq \sqrt{2}$ (you should know how to prove irrationality of $\sqrt{2}$!), we have either $c > \sqrt{2}$ or $c < \sqrt{2}$. 
    \begin{enumerate}
      \item Let $c < \sqrt{2} \iff c - \sqrt{2} > 0$. By the Archimidean principle, there exists a $k \in \mathbb{N}$ such that $0 < \frac{1}{k} < c - \sqrt{2}$. Then, $\frac{1}{k} \in \mathbb{Q}$ and $\mathbb{Q}$ is a field, so $c - \frac{1}{k} \in \mathbb{Q}$. 
      \begin{equation}
        c - \frac{1}{k} < c - c + \sqrt{2} = \sqrt{2}
      \end{equation}
      So $c$ is not least and so it must be the case that $c < \sqrt{2}$. 
      \item Let $c < \sqrt{2} \iff \sqrt{2} - c > 0$. By the Archimidean principle, there exists a $k \in \mathbb{N}$ such that $0 < \frac{1}{k} < \sqrt{2} - c$. Then, $c + \frac{1}{k} \in \mathbb{Q}$ and 
      \begin{equation}
        c + \frac{1}{k} < c + (\sqrt{2} - c) = c
      \end{equation}
      So $c$ is not an upper bound. 
    \end{enumerate}
    Note that given a well-defined $c = \sup{X}$ and in the case where $c < \sqrt{2}$, we have $2 - c^2 > 0$, so we can choose a well-defined $\delta$ satisfying (by Archimidean principle) 
    \begin{equation}
      0 < \delta < \min \bigg\{1, \frac{2 - c^2}{2c + 1} \bigg\}
    \end{equation}
    which gives us 
    \begin{align*}
      (c + \delta)^2 & = c^2 + \delta(2c + \delta) & \\
      & < c^2 + \delta (2 c + 1) & (\delta < 1) \\
      & < c^2 + (2 - c^2) = 2 & 
    \end{align*}
    meaning that $c$ is not an upper bound. Similarly for when $c > \sqrt{2}$. 
  \end{solution}

  \begin{exercise}[Zorich 2.2.15]
    Prove the equivalence of these two statements. 
    \begin{enumerate}
      \item If $X$ and $Y$ are nonempty sets of $\mathbb{R}$ having the property that $x \leq y$ for every $x \in X, y \in Y$, then there exists $c \in \mathbb{R}$ such that $x \leq c \leq y$ for all $x \in X$ and $y \in Y$. 
      \item Every set $X \subset \mathbb{R}$ that is bounded above has a least upper bound. 
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Let $S_1$ be the first statement and $S_2$ the second. 
    \begin{enumerate}
      \item ($S_2 \implies S_1$). Let $X \subset \mathbb{R}$ be a set that is bounded above, and $Y$ is a set such that $x \leq y$ for all $x \in X, y \in Y$. Then, by LUB principle, there exists $c = \sup{X} \in \mathbb{R}$. Now, we claim that $c \leq y$ for all $y \in Y$. Assume it doesn't: then there exists $y^\prime \in Y$ such that $y^\prime < c$. But since we assumed $x \leq y$ for all $x \in X, y \in Y$, we have $x \leq y^\prime$ for all $x \in X$, which means that $y^\prime$ is an upper bound of $X$. But $y^\prime < c$, contradicting the given fact that $c$ was the least upper bound. 

      \item ($S_1 \implies S_2$). Given a nonempty set $X \subset \mathbb{R}$, we wish to show the existence of $\sup{X}$. We are guaranteed the existence of nonempty set $Y \subset \mathbb{R}$ such that $x \leq y$ for all $x \in X, y \in Y$, which implies that $X$ must be bounded above. Then, by $S_1$, there must exist a $c \in \mathbb{R}$ such that 
      \begin{equation}
        x \leq c \leq y \text{ for all } x \in X, y \in Y
      \end{equation}
      We claim that $c = \sup{X}$. It is an upper bound of $X$ since $x \leq c$ for all $x \in X$. It is least since the set of all upper bounds of $X$ is $Y$, and $c \leq y$ for all $y \in Y$. 
    \end{enumerate}
  \end{solution}

  \begin{exercise}[Olmsted 1.15]
    Prove \textbf{Dedekind's Theorem}: Let the real numbers be divided into two nonempty sets $A$ and $B$ such that (i) if $x \in A$ and if $y \in B$, then $x < y$ and (ii) if $x \in \mathbb{R}$ then either $x \in A$ or $x \in B$, then there exists a number $c$ (which may belong to either $A$ or $B$) such that any number less than $c$ belongs to $A$ and any number greater than $c$ belongs to $B$. 
  \end{exercise}
  \begin{solution}
    This is really the same statement as Zorich 2.2.15.a, the original statement of completeness, but with the extra condition that the sets $A = X, B = Y$ must be disjoint.
  \end{solution}

  \begin{exercise}[Olmsted 1.7]
    If $x$ is an irrational number, under what conditions on the rational numbers $a, b, c, d$ is $(ax + b)/(cx + d)$ rational? 
  \end{exercise}
  \begin{solution}
    Note that a trivial solution is $a = b = c = d = 1$ which gives $1$. Since 
    \begin{equation}
      \frac{ax + b}{cx + d} = \frac{acx + ad - ad + bc}{cx + d} = a + \frac{bc - ad}{cx + d} 
    \end{equation}
    for the above to be rational it is necessary that $1/(cx + d)$ is rational. But this cannot be the case, which leaves us with the condition that $bc = ad$. 
  \end{solution}

  \begin{exercise}[Olmsted 1.8]
    Prove that the system of integers satisfies the axiom of completeness. 
  \end{exercise}
  \begin{solution}
    Let $S \subset \mathbb{Z}$ be bounded from above. It must have a maximum element (justify?), call it $c$. Then we claim that $c \in \mathbb{Z}$ is the least upper bound. Being the maximum, it is an upper bound, and $c$ is least since the next smallest element is $c-1$, which is less than $c \in S$, and therefore cannot be an upper bound. 
  \end{solution}

  \begin{exercise}[Zorich 2.2.16/Olmsted 1.16]
    Prove the following. 
    \begin{enumerate}
      \item If $A \subset B \subset \mathbb{R}$, then $\sup{A} \leq \sup{B}$ and $\inf{A} \geq \inf{B}$. 
      \item Let $\mathbb{R} \supset X \neq \emptyset$ and $R \supset Y \neq \emptyset$. If $x \leq y$ for all $x \in X, y \in Y$, then $X$ is bounded above , $Y$ is bounded below, and $\sup{X} \leq \inf{Y}$. 
      \item If the sets $X, Y$ in (b), are such that $X \cup Y = \mathbb{R}$, then $\sup{X} = \inf{Y}$. 
      \item If $X$ and $Y$ are the sets defined in (c), then either $X$ has a maximal element or $Y$ as a minimal element. 
      \item Show that Dedekind's theorem is equivalent to the axiom of completeness. 
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Listed. 
    \begin{enumerate}
      \item Let 
      \begin{align*}
        A^\prime & = \{ x \in \mathbb{R} \,|\, x \geq a \; \forall a \in A\} \\
        B^\prime & = \{ x \in \mathbb{R} \,|\, x \geq b \; \forall b \in B\}
      \end{align*}
      where we can easily verify that $B^\prime \subset A^\prime$. By definition, we get $\sup{B} = \min{B^\prime}$ and $\sup{A} = \min{A^\prime}$. But since $B^\prime \subset A^\prime$, for any $b^\prime \in B^\prime$, there exists an $a^\prime \in A^\prime$ such that $a^\prime \leq b^\prime$, which implies that $\sup{B} = \min{B^\prime} \leq \min{A^\prime} = \sup{A}$. 

      \item $X$ is bounded above by any element of $Y$. $Y$ is bounded below by any element of $X$. By the completion axiom, there exists a $c \in \mathbb{R}$ such that
      \begin{equation}
        x \leq c \leq y \text{ for all } x \in X, y \in Y
      \end{equation}
      Since $c$ is an upper bound of $X$, $\sup{X} \leq c$ by definition, and since $c$ is a lower bound of $Y$, $\inf{Y} \geq c$ by definition. Therefore, $\sup{X} \leq c \leq \inf{Y}$. 

      \item From completeness there exists a $c \in \mathbb{R}$ such that $x \leq c \leq y$ for all $x \in X, y \in Y$. $Y$ is, by definition, the set of \textit{all} upper bounds of $X$ (i.e. \textit{every} upper bound of $X$ is in $Y$, unlike $Y$ defined in 2.2.16.b). Since $c \leq y$ for all $y \in Y$, $c$ is minimal and so $c = \sup{X}$. $X$ is the set of all lower bounds of $Y$ by definition, so $c \geq x$ for all $x \in X \implies c = \inf{Y}$. So, $\inf{Y} = c = \sup{X}$. 

      \item We know that there exists $c = \inf{Y} = \sup{X}$. Since $X \cup Y = \mathbb{R}$, $c$ must be in at least $X$ or $Y$. If $c \in X$, then $c = \sup{X} = \max{X}$, and if $c \in Y$, then $c = \inf{Y} = \min{Y}$. 
      \item This is the same statement as Zorich 2.2.15.a (an iff equivalence, not just one way implying). 
    \end{enumerate}
  \end{solution}

  \begin{exercise}[Olmsted 1.13]
    Let $S$ be a nonempty set of numbers bounded above, and let $x$ be the least upper bound of $S$. Prove that $x$ has the two properties corresponding to an arbitrary positive number $\epsilon$: 
    \begin{enumerate}
      \item every element $s \in S$ satisfies the inequality $s < x + \epsilon$
      \item at least one element $s \in S$ satisfies the inequality $s > x - \epsilon$
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Listed. 
    \begin{enumerate}
      \item $x$ is an upper bound $\implies s \leq x$ for all $s \in S$, which implies that $s \leq x < x + \epsilon$. 
      \item By definition, $x - \epsilon$ cannot be an upper bound, so $x - \epsilon \geq s$ for all $s \in S$ is not true. Therefore, there must exist one $s \in S$ such that $s > x - \epsilon$. 
    \end{enumerate}
  \end{solution}

  \begin{exercise}[Zorich 2.2.18]
    Let $-A$ be the set of numbers of the form $-a$, where $a \in A \subset \mathbb{R}$. Show that $\sup(-A) = -\inf(A)$. 
  \end{exercise}
  \begin{solution}
    If $A$ is unbounded below, then $-\inf{A} = \infty$ and $-A$ is unbounded above, implying that $\sup{A} = \infty$. Now assume that $A$ is bounded below, then by completeness, it must have a greatest lower bound. Let us define the set $B = \{b \in \mathbb{R} \,|\, b \leq a \; \forall a \in A\}$. From 2.2.16.b, we have $b \leq \inf{A} \leq a$ for all $a \in A, b \in B$. Multiplying by $-1$ gives $-b \geq -\inf{A} \geq -a$ for all $a \in A, b \in B$, which is equivalent to saying 
    \begin{equation}
      a \leq -\inf{A} \leq b \text{ for all } a \in -A, b \in -B
    \end{equation}
    by definition of $-A, -B$. $-\inf{A}$ is clearly an upper bound of $-A$, and since  
    \begin{align*}
      B & = \{b \in \mathbb{R}\,|\, b \leq a \; \forall a \in A\} \\
      & = \{b \in \mathbb{R}\,|\, -b \geq -a \; \forall a \in A\} \\
      & = \{b \in \mathbb{R}\,|\, -b \geq a \; \forall a \in -A\} 
    \end{align*}
    implies that $-B = \{b \in \mathbb{R}\,|\, b \geq a \; \forall a \in -A\}$ is the set of all upper bounds of $A$. So, $-\inf{A}$ is the least upper bound of $-A$, i.e. $-\inf{A} = \sup(-A)$. 
  \end{solution}

  \begin{exercise}[Zorich 2.2.21]
    Show that the set $\mathbb{Q}(\sqrt{n})$ of numbers of the form $a + b \sqrt{n}$ where $a, b \in \mathbb{Q}$, $n$ is a fixed natural number that is not the square of any integer, is an ordered set satisfying the principle of Archimedes but not the axiom of completeness. 
  \end{exercise}
  \begin{solution}
    The order on $\mathbb{Q}(\sqrt{n})$ can be embedded from the ordering on the reals by defining the canonical injection map $i:\mathbb{Q}(\sqrt{n}) \longrightarrow \mathbb{R}$ and defining for any $x, y \in \mathbb{Q}(\sqrt{n})$, 
    \begin{equation}
      x \leq_{\mathbb{Q}(\sqrt{n})} y \iff i(x) \leq_{\mathbb{R}} i(y)
    \end{equation}
    Now, let $h > 0$ be any fixed real number, and $x = (a,b) = a + b\sqrt{n}$. By the Archimidean principle, we can find a $k \in \mathbb{Z}$ such that
    \begin{equation}
      (k - 1) h \leq x \leq k h \text{ for some } x \in \mathbb{Q}(\sqrt{n}) \subset \mathbb{R}
    \end{equation}
    We now show that $\mathbb{Q}(\sqrt{n})$ is not complete since it doesn't satisfy the LUB property. Since there are infinite prime numbers in $\mathbb{N}$, choose a prime number $p$ that is not a factor of $n$. Then, we are guaranteed that $pn$ is not a perfect square, and can define the set 
    \begin{equation}
      X = \{x \in \mathbb{Q}(\sqrt{n})\,|\, x < \sqrt{pn} \} \subset \mathbb{Q}(\sqrt{n})
    \end{equation}
    and assume that $c = c_1 + c_2 \sqrt{n} = \sup{X}$ exists ($c_1, c_2 \in \mathbb{Q}$). Clearly, $c \neq \sqrt{pn} \not\in \mathbb{Q}(\sqrt{n})$. 
    \begin{enumerate}
      \item Assume $c < \sqrt{pn} \iff 0 < \sqrt{pn} - c \in \mathbb{R}$. By the Archimidean principle, there exists a $k \in \mathbb{N}$ such that $0 < \frac{1}{k} < \sqrt{pn} - c $. Then, we can verify that $c + \frac{1}{k} = (c_1 + \frac{1}{k}) + c_2 \sqrt{n} \in \mathbb{Q}(\sqrt{n})$ and 
      \begin{equation}
        c + \frac{1}{k} < c + \sqrt{pn} - c = \sqrt{pn} \implies c + \frac{1}{k} \in X
      \end{equation}
      implies that $c$ is not an upper bound. So we must turn to case 2. 
      \item Assume $c > \sqrt{pn} \iff c - \sqrt{pn} > 0$. By AP, there exists a $k \in \mathbb{N}$ such that $0 < \frac{1}{k} < c - \sqrt{pn}$. Then, we can verify that $c - \frac{1}{k} \in \mathbb{Q}(\sqrt{n})$ and 
      \begin{equation}
        c - \frac{1}{k} > c - c + \sqrt{pn} = \sqrt{pn}
      \end{equation}
      implies that $c - \frac{1}{k}$ is an upper bound of $X$, so $c$ is not least. 
    \end{enumerate}
    Therefore, by contradiction, $c$ does not exist. 
  \end{solution}

  \begin{exercise}[Zorich 2.2.22]
    Let $n \in \mathbb{N}$ and $n > 1$. In the set $E_n = \{0, 1, \ldots, n-1\}$, we define the sum and product of two elements as the remainders when the usual sum and product in $\mathbb{R}$ are divided by $n$. With these operations on it, the set $E_n$ is denoted $\mathbb{Z}_n$. 
    \begin{enumerate}
      \item Show that if $n$ is not a prime number, then there are nonzero numbers $m, k \in \mathbb{Z}_n$ such that $m \cdot k = 0$, i.e. there exist nonzero zero divisors. 
      \item Show that if $p$ is prime, then there are no zero divisors in $\mathbb{Z}_p$ and $\mathbb{Z}_p$ is a field. 
      \item Show that, no matter what the prime $p$, $\mathbb{Z}_p$ cannot be ordered in a way consistent with the arithmetic operations on it. 
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Listed. 
    \begin{enumerate}
      \item $n$ is composite implies that there exist $1 < m, k < n$ such that $n = m k$. These factors $m, k$ are precisely the zero divisors of $\mathbb{Z}_n$ since $m k = n \equiv 0 \pmod{n}$. 
      \item With $p$ prime, assume that there are nontrivial zero divisors $1 < m, k < p$ in $\mathbb{Z}_p$. Then, $m k \equiv 0 \pmod{n} \implies m k = l p$ for some $l \in \mathbb{N}$. But this implies that $m$ or $k$ must divide $p$, which is impossible since $1 < m, k < p$. Then prove field axioms. 
      \item For any field, we must have $0 \leq 1$, because if not, then 
      \begin{equation}
        0 > 1 \implies 0 < 1^{-1} \cdot 1 = 1^{-1} \implies 0 \cdot 0 < 1^{-1} \cdot 1^{-1} = 1
      \end{equation}
      So, $0 \leq 1$ implies that $0 \leq 1 \leq 2 \leq \ldots \leq p-1$. But 
      \begin{equation}
        0 + 1 \leq (p - 1) + 1 = 0
      \end{equation}
      is false, so any ordering is impossible. 
    \end{enumerate}
  \end{solution}

  \begin{exercise}[Zorich 2.2.23]
    Show that if $\mathbb{R}$ and $\mathbb{R}^\prime$ are two models of the set of real numbers and $f: \mathbb{R} \longrightarrow \mathbb{R}^\prime$ (with $f \not\equiv 0^\prime$) is a mapping such that $f(x + y) = f(x) + f(y)$ and $f(x \cdot y) = f(x) \cdot f(y)$ for any $x, y \in \mathbb{R}$. Prove that $f$ is an order-preserving isomorphism. 
  \end{exercise}
  \begin{solution}
    Let $0, 0^\prime$ be the additive identity of $\mathbb{R}, \mathbb{R}^\prime$, respectively, and $1, 1^\prime$ the multiplicative identity. We claim that $f(0) = 0^\prime$ since
    \begin{align*}
      f(0) & = f(0 + 0) & \text{(definition of additive identity)} \\
      & = f(0) + f(0) & (\text{homomorphism over } + )
    \end{align*}
    which implies that $f(0) + f(0) = f(0) = 0^\prime + f(0) $. Since $f(0)$ lives in field $\mathbb{R}^\prime$, its additive identity $-f(0)$ is well defined, and we get $f(0) = f(0) + f(0) + (-f(0)) = 0^\prime + f(0) + (-f(0)) = 0^\prime$. We also claim that $f(1) = 1^\prime$ since 
    \begin{align*}
      f(1) & = f(1 \cdot 1) & \text{(definition of multiplicative identity)} \\
      & = f(1) \cdot f(1) & (\text{homomorphism over } \cdot ) 
    \end{align*}
    which implies that $f(1) \cdot f(1) = 1^\prime \cdot f(1)$. Since $f(1)$ lives in field $\mathbb{R}^\prime$, its multiplicative identity $f(1)^{-1}$ is well defined, and we get $f(1) = f(1) \cdot f(1) \cdot f(1)^{-1} = 1^\prime \cdot f(1) \cdot f(1)^{-1} = 1^\prime$. Now that we have proved mapping of identities, this implies the mapping of inverses. 
    \begin{align*}
      0^\prime & = f(0) = f(x - x) = f(x) + f(-x) \implies f(-x) = -f(x) \\
      1^\prime & = f(1) = f(x \cdot x^{-1}) = f(x) \cdot f(x^{-1}) \implies f(x^{-1}) = f(x)^{-1}
    \end{align*}
    With these conditions, we have proved that $f$ is a homomorphism of fields. Now we prove that $f$ is a bijection, but first, we claim that $f(x) = 0^\prime \implies x = 0$. Assume that there exists a nonzero $x \in \mathbb{R}$ such that $f(x) = 0^\prime$. Then, $x^{-1}$ is well defined, and 
    \begin{align*}
      f(x) \cdot f(x^{-1}) & = f(x) \cdot f(x)^{-1} = 0^\prime \\
      f(x) \cdot f(x^{-1}) & = f(x \cdot x^{-1}) = f(1) = 1^\prime 
    \end{align*}
    which implies that $0^\prime = 1^\prime$. So, $f(1) = 1^\prime = 0^\prime$, and so for all $k \in \mathbb{R}$, $f(k) =f(k \cdot 1) = f(k) \cdot f(1) = f(k) \cdot 0^\prime = 0^\prime \implies f \equiv 0^\prime$, leading to a contradiction of the assumption that $f^\prime \not\equiv 0^\prime$. 
    \begin{enumerate}
      \item ($f$ injective). Assume $f$ is not injective, i.e. there exists distinct $x_1, x_2 \in \mathbb{R}$ s.t. $f(x_1) = f(x_2)$. Then, using that fact $f(x) = 0 \implies x = 0$, 
      \begin{equation}
        0 = f(x_1) - f(x_2) = f(x_1 - x_2) \implies x_1 - x_2 = 0 \implies x_1 = x_2
      \end{equation}
      \item ($f$ surjective). Let $y$ be any nonzero element in $\mathbb{R}^\prime$ (clearly if $y=0^\prime$ then its preimage is $0$) and $y^{-1}$ its multiplicative inverse. Assume there exists no $x \in \mathbb{R}$ satisfying $f(x) = y$, meaning that there exist no $x$ satisfying
      \begin{equation}
        f(x) \cdot  = y \cdot y^{-1} = 1^\prime
      \end{equation}
      But since $f$ maps inverses to inverses, we can choose $x = (y^{-1})^{-1}$, which leads to 
      \begin{equation}
        f(x) \cdot y^{-1} = (
      \end{equation}
    \end{enumerate}
    Finally, we prove that $f$ is order preserving. Assume that $x \leq y \iff 0 \leq y - x$ , we wish to prove that 
    \begin{equation}
      f(x) \leq f(y) \iff 0 \leq f(y) - f(x) = f(y - x)
    \end{equation}
    Therefore, since this preservation of ordering is really the statement $0 \leq y - x \implies 0 \leq f(y - x)$, it suffices to prove that $0 \leq x \implies 0 \leq f(x)$. Now, assume that we have a $x$ such that $f(x) < 0^\prime$. Adding it with the equation $f(1) = 1^\prime$ gives us 
    \begin{equation}
      f(x + 1) < 1^\prime
    \end{equation}
    It is easy to prove that $0 \leq x \iff 0 \leq x^{-1}$. Now assume that $0 > f(x)$. \textbf{INCOMPLETE}
  \end{solution}

  \begin{exercise}[Density of Rationals in $\mathbb{R}$]
    Prove that for any two distinct $a < b \in \mathbb{R}$, there exists an infinite number of rational numbers between $a$ and $b$. 
  \end{exercise}
  \begin{solution}
    Since $a< b$, then $b - a > 0$ and by the Archimidean principle, there exists a $k \in \mathbb{N}$ such that 
    \begin{equation}
      0 < \frac{1}{k} < b - a \implies 1 < kb - ka
    \end{equation}
    which implies that the length of $[ka, kb)$ greater than $1$. By the inductive property of $\mathbb{Z}$, there must be an integer $p \in [ka, kb)$. If there were not, then this would imply that $[ka+1, kb+1)$ and $[ka-1, kb-1)$ had no integers and repeating would mean that there were no integers in $\mathbb{R}$. Therefore, 
    \begin{equation}
      ka \leq p < kb \implies a \leq \frac{p}{k} < b
    \end{equation}
    for all $a, b \in \mathbb{R}$, with $p/k \in \mathbb{Q}$. If $a$ is irrational we can replace the $\leq$ to $<$, leaving $a \leq \frac{p}{k} < b$, and if $a$ is rational, we can construct another rational $a + \frac{1}{k} \in (a, b)$. 
  \end{solution}

  \begin{exercise}[Nested Interval Lemma]
    With the fact that $\mathbb{R}$ is complete, prove the following. 
    \begin{enumerate}
      \item For a sequence of closed nested intervals $I_1 \supset I_2 \supset \ldots$ of $\mathbb{R}$, there exists a point $c \in \mathbb{R}$ belonging to all these intervals. 
      \item Furthermore, if the hypothesis also satisfies the fact that for any $\epsilon > 0$, there exists a $k \in \mathbb{N}$ such that $|I_k| < \epsilon$ (i.e. the length of the intervals decreases to $0$), then the point $c$ common to all sets is unique. 
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Listed. 
    \begin{enumerate}
      \item Let $I_n = [a_n, b_n]$, with $a_n < b_n$ finite for all $n \in \mathbb{N}$. For all $n\in \mathbb{N}$, we have $I_n = [a_n, b_n]$ and can take the two subsets $X_n = (-\infty, a_n)$ and $Y_n = (b_n, \infty)$, where $x \leq y$ for every $x \in X_n, y \in Y_n$. We also have the fact that $\mathbb{R} = X_n \cup I_n \cup Y_n$. Since $\mathbb{R}$ is complete, there exists a $c$ such that $x \leq c \leq y$ for all $x \in X, y \in Y$. But $x \leq c \iff c \not\in X_n$ and $c \leq y \iff c \not\in Y_n$, so for all $n \in \mathbb{N}$, $c$ must be in $I_n$. 
      
      \item Since we have proved (a), it now suffices to prove uniqueness of $c$. Let there be two distinct points $c_1, c_2 \in \mathbb{R}$ belonging to these intervals. Without loss of generality, assume $c_1 - c_2 > 0$, and choose 
      \begin{equation}
        \epsilon = \frac{c_1 - c_2}{3}
      \end{equation}
      Then, there should exist a $k \in \mathbb{N}$ such that $|I_k| < \epsilon$. Since $I_k$ must contain $c_1$, it must be a subset of $[c_1 - \epsilon, c_1 + \epsilon]$ (should be able to see why) and similarly for $c_2$. 
      \begin{align*}
        I_k & \subset \bigg[ c_1 - \frac{c_1 - c_2}{3}, c_1 + \frac{c_1 - c_2}{3} \bigg] = \bigg[ \frac{2c_1 + c_2}{3}, \frac{4c_1 - c_2}{3} \bigg] = L\\
        I_k & \subset \bigg[ c_2 - \frac{c_1 - c_2}{3}, c_2 + \frac{c_1 - c_2}{3} \bigg] = \bigg[ \frac{-c_1 + 4 c_2}{3}, \frac{c_1 + 2c_2}{3} \bigg] = M
      \end{align*}
      But since $c_1 > c_2 \implies \frac{c_1 + 2c_2}{3} < \frac{2c_1 + c_2}{3}$, $L$ and $M$ are disjoint $\implies I_k$, as a subset of both, leaves us with $I_k = \emptyset$, contradicting that it is a closed interval. 
    \end{enumerate}
  \end{solution}

  \begin{exercise}{Compactness of Closed Interval in $\mathbb{R}$}
    Prove that any system of open intervals covering (i.e. an open cover of) a closed interval contains a finite subsystem that covers the closed interval. Another way to state this is by saying that every closed interval of $\mathbb{R}$ is compact. 
  \end{exercise}
  \begin{solution}
    A closed interval with a finite open covering is trivially compact since any subcovering is also finite. We only need to deal with when a closed interval $I = [a, b]$ has an infinite open covering $\{U_\alpha \}_{\alpha \in A}$, which means that the set of indices $A$ is infinite. Assume that there exists no finite covering of $I$. Then, we divide $I$ into two halves 
    \begin{equation}
      I_1 = \Big[ a, \frac{a + b}{2}\Big], \;\; I_2 = \Big[ \frac{a + b}{2}, b \Big]
    \end{equation}
    and define a subcovering for each of them. That is, we can define $A_1 \subset A$ and $A_2 \subset A$ such that $\{U_\alpha\}_{\alpha \in A_1} \subset \{U_\alpha \}_{\alpha \in A}$ is a covering of $I_1$ and $\{U_\alpha\}_{\alpha \in A_2} \subset \{U_\alpha \}_{\alpha \in A}$ is a covering of $I_2$. At least one of $A_1$ or $A_2$ must be infinite, since if they were both finite, then we can define a finite covering $\{U_\alpha\}_{\alpha \in A_1 \cup A_2}$ of $I$. Choose the interval with the infinite covering and repeat this procedure, which will result in a nested interval that decreases in length by a half. 
    \begin{equation}
      I \supset I_1 \supset I_2 \supset \ldots
    \end{equation}
    By the nested interval lemma, there exists a unique point $c$ common to all these intervals. But since $c \in [a, b]$, the open cover $\{U\}$ should contain an open interval $(c - \delta_1, c + \delta_2)$ containing $c$. We wish to prove that this interval is a superset of some $I_k$ in the sequence above, contradicting the fact that $I_k$ has an infinite cover. Since the length of each $I_i$ decreases arbitrarily (i.e. we can choose any $\epsilon > 0$ and find a $I_k$ with length less than $\epsilon$), we choose $\epsilon = \frac{1}{2} \min\{\delta_1, \delta_2 \}$, and we should be able to find some $I_k$ that is a subinterval of $[c - \epsilon, c + \epsilon]$, which itself is a subinterval of $(c - \delta_1, c + \delta_2)$. 
    \begin{equation}
      I_k \subset \Big[ c - \frac{1}{2} \min\{\delta_1, \delta_2 \}, c + \frac{1}{2} \min\{\delta_1, \delta_2 \} \Big] \subset (c - \delta_1, c + \delta_2)
    \end{equation}
    Therefore, $(c - \delta_1, c + \delta_2)$ is a finite cover of $I_k$, contradicting the fact that all $I_k$'s have infinite covers. 
  \end{solution}

  \begin{exercise}[Bolzano-Weierstrass Theorem]
    Prove that every bounded infinite set of real numbers has at least one limit point. (A limit point $p$ of set $X$ is a point such that every open neighborhood of $p$ contains an infinite number of elements of $X$). 
  \end{exercise}
  \begin{solution}
    Let the set of points be denoted $X$, and let $a$ be the lower bound and $b$ be the upper bound. Then, $X \subset [a, b] = I$. Now divide $[a, b]$ into halves $[a, \frac{a + b}{2}] \cup [\frac{a + b}{2}, b]$. At least one of the halves must have an infinite number of points; choose the interval with infinite points as $I_1$ and doing this repeatedly gives the nested sequence 
    \begin{equation}
      I \supset I_1 \supset I_2 \supset \ldots
    \end{equation}
    By the nested interval lemma, there exists at least one point $c \in \mathbb{R}$ that is in all these intervals. Furthermore, since $|I_i| = \frac{1}{2^i} (b - a)$ decreases to $0$, we can choose a $\epsilon > 0$ and find an interval $I_k$ with $|I_k| < \epsilon$. We claim that $c$ is a limit point of $X$. Given an $\epsilon$, we wish to prove that there are an infinite number of points within the $\epsilon$-neighborhood $(c - \epsilon, c + \epsilon)$ of $c$. Since we can find some $I_k$ with $|I_k| < \epsilon$, we can see that 
    \begin{equation}
      I_k \subset (c - \epsilon, c + \epsilon)
    \end{equation}
    and therefore the $\epsilon$-neighborhood of $c$ contains $I_k$, which contains an infinite number of points in $X$. 

    We can construct another proof that is dependent on the compactness lemma. This construction will be useful for problem 2.3.4. Let $X$ be a given subset of $\mathbb{R}$, and it follows from the definition of boundedness that $X$ is contained in some closed interval $I \subset \mathbb{R}$. We show that at least one point of $I$ is a limit point of $X$. Assume that it is not. Then each point $x \in I$ would have a neighborhood $U(x)$ containing at most a finite number of points from $X$. The totality of such neighborhoods $\{U(x)\}$ constructed for the points $x \in I$ forms an open covering of $X$. Since $I$ is closed, it is compact and therefore we can find a finite subcovering $\{U_i(x)\}_{i}$ of open intervals that cover $I$ and therefore cover $X$. This open cover $\{U_i(x)\}_{i}$ of $X$ is a finite union of sets that each contain at most a finite number of points from $X$, so the covering of $X$ contains a finite number of points from $X$, a contradiction that $X$ contains infinite points. 
  \end{solution}

  \begin{exercise}[Zorich 2.3.1]
    Show that 
    \begin{enumerate}
      \item if $I$ is any system of nested closed intervals, then 
      \[\sup\{ a \in \mathbb{R}\,|\, [a, b] \in I\} = \alpha \leq \beta = \inf\{ b \in \mathbb{R}\,|\,[a, b] \in I\}\]
      and 
      \[[\alpha, \beta] = \bigcap_{[a, b] \in I} [a, b]\]
      \item if $I$ is a system of nested open intervals $(a, b)$, the intersection
      \[\bigcap_{(a, b) \in I} (a, b)\] 
      may happen to be empty. 
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Listed. 
    \begin{enumerate}
      \item (May be tempted to say that $a_1 \leq a_2 \leq \ldots$, but this assumes that the indexing set $I$ is countable). We claim that for any two intervals $[a_n, b_n]$ and $[a_m, b_m]$ in $I$, 
      \[a_n \leq b_m\]
      Assume that $a_n > b_m$. Then $b_n \geq a_n > b_m \geq a_m$ implies that $[a_n, b_n]$ and $[a_m, b_m]$ are disjoint, contradicting the fact that they are nested. Now given that $X$ is the set of $a_n$'s and $Y$ is the set of $b_n$'s, we have $x \leq y$ for all $x \in X, y \in Y$. So by 2.2.16.b, we have $\sup{X} \geq \inf{Y}$. 
      \\
      To prove the second statement, we show that trying to ``expand'' the interval $[\alpha, \beta]$ will lead to a contradiction. Since $\alpha$ is the LUB, given any $\epsilon > 0$, there exists a $(a_l, b_l) \in X$ such that $\alpha - \epsilon < a_l < \alpha$, which implies that $ [\alpha, \beta] \subset [a_l, \beta] \subset [\alpha - \epsilon, \beta]$. Assuming that this extended interval is the intersection, we should be able to choose any point in $[\alpha - \epsilon, \beta]$ and find that it is in every element of $I$. We choose a point in $[\alpha - \epsilon, a_l)$, which is not in the interval $(a_l, b_l)$. We do the same for $\beta \mapsto \beta + \epsilon$. We also check that ``shrinking'' the interval $[\alpha, \beta] \mapsto [\alpha + \epsilon, \beta]$ is no good, since we can find an element in $[\alpha, \alpha + \epsilon)$ that is in every interval in $I$. 
      
      \item Take the system of nested open intervals 
      \[(0, 1) \supset (0, \frac{1}{2}) \supset (0, \frac{1}{3}) \ldots (0, \frac{1}{n}) \supset \ldots\]
      Take their infinite intersection, denote it $S$, and assume that some $\epsilon \in (0, 1)$ is in $S$. Since $\epsilon$ is a real number, by the Archimidean principle there exists a $k \in \mathbb{N}$ such that $\frac{1}{k} < \epsilon$. Therefore, $\epsilon \not\in (0, \frac{1}{k}) \implies \epsilon \not\in S$. 
    \end{enumerate}
  \end{solution}

  \begin{exercise}[Zorich 2.3.2]
    Show that 
    \begin{enumerate}
      \item from a system of closed intervals covering a closed interval it is not always possible to choose a finite subsystem covering the interval. 
      \item from a system of open intervals covering a open interval it is not always possible to choose a finite subsystem covering the interval. 
      \item from a system of closed intervals covering a open interval it is not always possible to choose a finite subsystem covering the interval. 
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    We show with the interval $(0, 1)$ or $[0, 1]$. Using linear transformations it is easy to generalize this to any other interval $(a, b)$ or $[a, b]$. 
    \begin{enumerate}
      \item Consider the infinite covering
      \[[0, 1] = \big[0, \frac{1}{2}\big] \cup \big[\frac{1}{2}, \frac{3}{4}\big] \cup \big[\frac{3}{4}, \frac{7}{8}\big] \cup \ldots \]
      \item Consider the infinite covering 
      \[(0, 1) = \big(0, \frac{1}{2}\big) \cup \big(\frac{1}{2}, \frac{3}{4}\big) \cup \big(\frac{3}{4}, \frac{7}{8}\big) \cup \ldots \]
      \item Consider the infinite covering 
      \[(0, 1) = \big[0, \frac{1}{2}\big] \cup \big[\frac{1}{2}, \frac{3}{4}\big] \cup \big[\frac{3}{4}, \frac{7}{8}\big] \cup \ldots \]
    \end{enumerate}
  \end{solution}

  \begin{exercise}[Zorich 2.3.3]
    Show that if we only take the set $\mathbb{Q}$ of rational numbers instead of the complete set $\mathbb{R}$ of real numbers, with the definitions of closed, open, and neighborhood of a point $r \in \mathbb{Q}$ to mean respectively the corresponding subsets of $\mathbb{Q}$, then none of the three lemmas is true. 
  \end{exercise}
  \begin{solution}
    We prove only for the nested interval lemma. We choose the series of nested intervals 
    \[\bigg( \sqrt{2} - \frac{1}{n}, \sqrt{2} + \frac{1}{n} \bigg)\]
    with $n \in \mathbb{N}$. Assume that there is a $r \in \mathbb{Q}$ such that 
    \[r \in \bigg( \sqrt{2} - \frac{1}{n}, \sqrt{2} + \frac{1}{n} \bigg) \text{ for all } n \in \mathbb{N}\]
    which is equivalent to saying that $\big| r - \sqrt{2}\big| < \frac{1}{n}$ for all $n \in \mathbb{N}$. Clearly, $r \neq \sqrt{2}$, and by the Archimidean principle, there exists a $k \in \mathbb{N}$ such that 
    \[0 < \frac{1}{k} < |r - \sqrt{2}|\]
    which contradicts the above. 
  \end{solution}

  \begin{exercise}[Zorich 2.3.4]
    Show that the three lemmas above are equivalent to the axiom of completeness. 
  \end{exercise}
  \begin{solution}
    Note that from the proofs, completeness implies nested interval lemma, which implies compactness of closed intervals, which implies the Bolzano-Weierstrass theorem. So, it is sufficient to prove that Bolzano-Weierstrass theorem implies completeness to determine equivalence. There are not a lot of direct proofs, so we prove that Weierstrass implies nested interval, which implies completeness. 
    \begin{enumerate}
      \item (Weierstrass $\implies$ Nested) Assume that we have $\mathbb{R}$ with the Bolzano-Weierstrass theorem. Take the series of nested closed intervals 
      \[I = [a, b] \supset I_1 = [a_1, b_1] \supset I_2 = [a_2, b_2] \supset \ldots\]
      We see that $a \leq a_i \leq b$, so the infinite sequence of monotonically nondecreasing values $a_i$ is bounded. Therefore, it must have a limit point, which we will denote as $c$. We claim that $a_i \leq c$ for all $a_i$. Since if it were not, then $c < a_i$ for some $i$, and choosing $\epsilon = 0.5 (a_i - c)$, the $\epsilon$-neighborhood of $c$ will not contain $a_j$ for $j \geq i$ since 
      \[c < a_i \implies 0.5 c < 0.5 a_i \implies c + \epsilon = 0.5c + 0.5 a_i < a_i< a_{i+1} < \ldots\]. 
      With similar reasoning, we can conclude that $b_i \geq c$ for all $b_i$. This implies that $a_i \leq c \leq b_i$ for all $i$ which is equivalent to saying that $c \in [a_i, b_i] = I_i$ for all $i \in \mathbb{N}$. 
      
      \item (Nested $\implies$ LUB Principle) Let $X \subset \mathbb{R}$ be a set that is bounded above, with $b_1$ any upper bound. Since $X$ is nonempty, there exists $a_1 \in X$ that is not an upper bound (otherwise, $X$ would be a singleton set and it trivially has a least upper bound). Consider the well-defined interval $[a_0, b_0]$. Take the mean $m_0 = 0.5 (a_0 + b_0)$, and if $m_0$ is an upper bound, set it to $b_1$ (with $a_1 = a_0$) and $a_1$ if else (with $b_1 = b_0$). Then, we have a sequence of nested intervals 
      \[[a_0, b_0] \supset [a_1, b_1] \supset [a_2, b_2] \supset \ldots \]
      of decreasing lengths $|I_k| = \frac{1}{2^{k}} (b - a)$. All of them must contain a unique common point $c \in \mathbb{R}$ by the nested intervals lemma, which implies that 
      \[a_0 \leq a_1 \leq a_2 \leq \ldots \leq c \leq \ldots \leq b_2 \leq b_1 \leq b_0\]
      I claim two things: 
      \begin{enumerate}
          \item $c$ is an upper bound for $X$. Suppose it were not, then there exists some $x \in X$ such that $c < x$, and let the distance between them be $\epsilon = x - c > 0$. By AP, we can choose $k \in \mathbb{N}$ such that $\frac{1}{k} < \epsilon$. All the $b_n$ are upper bounds of $X$, so we have $x \leq b_n$. Subtracting $c$ on both sides gives 
          \[0 < x - c = \epsilon \leq b_n - c \leq |I_n| = \frac{1}{2^n} (b_0 - a_0)\]
          where the last inequality follows from $c \in I_n = [a_n, b_n]$, so the maximum distance it can be from the endpoint $b_n$ is $|I_n|$. The inequality above holds for all $n \in \mathbb{N}$, so increasing $n$ arbitrarily should decrease $\frac{1}{2^n} (b_0 - a_0)$ past $\epsilon$. To formalize this, we use the inequality
          \[\frac{1}{2^n} < \frac{1}{n} \text{ for all } n \in \mathbb{N}\]
          and so we have 
          \[\epsilon \leq b_n - c < \frac{1}{n} (b_0 - a_0)\]
          We choose the natural number $n = \lceil \frac{2(b_0 - a_0)}{\epsilon} \rceil$, which does not satisfy the inequality above since 
          \[\epsilon < \frac{1}{n} (b_0 - a_0) = \frac{1}{\lceil 2(b_0 - a_0)/\epsilon \rceil} (b_0 - a_0) \leq \frac{\epsilon}{2(b_0 - a_0)} (b_0 - a_0) = \frac{\epsilon}{2}\]
          This leads to a contradiction. 
          \item We now prove that $c$ is least. Assume that $c$ is not least $\implies$ there exists an upper bound $B$ such that $B < c$ and $x \leq B$ for all $x \in X$. \textbf{INCOMPLETE}
      \end{enumerate}
    \end{enumerate}
  \end{solution}

  \begin{exercise}[Zorich 2.4.1]
  Show that the set of real numbers has the same cardinality as the points of the interval $(-1, 1)$. 
  \end{exercise}
  \begin{solution}
    We define the bijective map $\rho: (-1, 1) \longrightarrow \mathbb{R}$ 
    \[p(x) = \begin{cases} 
    0 & \text{ if } x = 0 \\
    \frac{1}{x} & \text{ if } x \neq 0 
    \end{cases}\]
  \end{solution}

  \begin{exercise}[Zorich 2.4.2]
    Give an explicit one-to-one correspondence between 
    \begin{enumerate}
      \item the points of two open intervals 
      \item the points of two closed intervals 
      \item the point of a closed interval and an open interval 
      \item the points of the closed interval $[0, 1]$ and $\mathbb{R}$
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Listed. 
    \begin{enumerate}
      \item $\rho: (a, b) \longrightarrow (c, d)$ defined 
      \[\rho(x) = \frac{d - c}{b - a} (x - a) + c \]
      \item the extension of $\rho$ defined on (a) to $[a, b]$
      \item From (a) and (b), it suffices to prove a bijection from $(0, 1)$ to $[0, 1]$. We extract a countably infinite sequence from $(0, 1)$, say 
      \[x_1 = \frac{1}{3}, \; x_2 = \frac{1}{4}, \ldots, x_i = \frac{1}{i+2}\]
      Then, we define bijection $\rho: (0, 1) \longrightarrow [0, 1]$ as 
      \[\rho (x) = \begin{cases}
      x & \text{ if } x \not\in \{x_i\} \\
      0 & \text{ if } x = x_1 = \frac{1}{2} \\
      1 & \text{ if } x = x_2 = \frac{1}{3} \\
      x_{i-2} & \text{ if } x = x_i \text{ for } i > 2
      \end{cases}\]
      Colloquially, we extract a copy of $\mathbb{N}$ from $(0, 1)$ and use the bijection $\mathbb{N} \simeq \mathbb{N} \cup \{0, -1\}$ to ``shift'' the terms. 
      \item We compose the bijections $\rho_1 : [0, 1] \longrightarrow (0, 1)$ and $\rho_2: (0, 1) \longrightarrow \mathbb{R}$. 
    \end{enumerate}
  \end{solution}

  \begin{exercise}[Zorich 2.4.3]
    Show that 
    \begin{enumerate}
      \item every infinite set contains a countable subset
      \item the set of even integers has the same cardinality as the set of all natural numbers
      \item the union of an infinite set and an at most countable set has the same cardinality as the original infinite set. 
      \item the set of irrational numbers has the cardinality of the continuum 
      \item the set of transcendental numbers has the cardinality of the continuum
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Listed. 
    \begin{enumerate}
      \item Let $A$ be an infinite set. By axiom of choice, choose $a_0 \in A$. Then, $A \setminus \{a_0\} \neq \emptyset$ since $A$ is infinite. By induction, assume you have chosen $a_0, a_1, \ldots, a_k \in A$. Then, since $A$ is infinite, $A \setminus \{a_0, a_1, \ldots, a_k\} \neq \emptyset$, so we can choose $a_{k+1} \in A \setminus \{a_0, \ldots, a_k\}$. Thus, we have constructed a countable subset $\{a_k\}_{k \in \mathbb{N}}$ of $A$. 
      \item Given the quotient ring $2\mathbb{Z}$, define the bijection $\rho: 2\mathbb{Z} \longrightarrow \mathbb{N}$ as 
      \[p(x) = \begin{cases} 
      x + 2 & \text{ if } x \geq 0 \\
      -x - 1 & \text{ if } x < 0 
      \end{cases}\]
      \item From (a), we can extract a countable set from original set $A$, call it $X$. Since the product of countable sets is countable ($\mathbb{N} \cup \mathbb{N}$ is countable), we can define a bijection $\Tilde{\rho}: X \longrightarrow X \cup B$. Therefore, we can define a bijection $\rho: A \longrightarrow A \cup B$ as 
      \[\rho(x) = \begin{cases} 
      x & \text{ if } x \in A \setminus X \\
      \Tilde{\rho}(x) & \text{ if } x \in X
      \end{cases}\]
      \item $\mathbb{Q}$ is countable and $\mathbb{R}$ is uncountable. So, $\mathbb{R} \setminus \mathbb{Q}$ must be uncountable since if it were countable, then the union of the rationals and irrationals, which is $\mathbb{R}$, would be countable. 
      \item It suffices to prove that the set of algebraic numbers (numbers that are possible roots of a polynomial with integer coefficients with leading coefficient nonzero) is countable, since we can apply (d) right after. The set of all $k$th degree polynomials with integer coefficients is isomorphic to $\mathbb{Z}^k$ through the map 
      \[a_k x^k + a_{k-1} x^{k-1} + \ldots + a_2 x^2 + a_1 x^1 + a_0 \mapsto (a^{k-1}, a^{k-2}, \ldots, a_1, a_0)\]
      and the union of these countable sets (minus the $0$ map) 
      \[P = \bigg(\bigcup_{k = 1}^\infty \mathbb{Z}^k\bigg) \setminus \{0\} = \big(\mathbb{Z} \setminus \{0\}\big) \cup \mathbb{Z}^2 \cup \ldots \]
      is countable. For any element in $\mathbb{Z}^k$, there are at most $k$ real roots, and so we can define the set of roots of an element $z \in \mathbb{Z}^k \subset P$ as a $j$-tuple of algebraic numbers, which can have at most $j=k$ roots. 
      \[r(z) = \underbrace{(r_{1z}, r_{2z}, \ldots, r_{jz})}_{j \leq k}\]
      Therefore, the union of all these $j$-tuples for all $z \in P$ 
      \[\bigcup_{z \in P} r(z) = \bigcup_{k = 1}^\infty \bigcup_{z \in \mathbb{Z}^k} r(z)\]
      is a countable union of a countable union of finite sets, making it countable. 
    \end{enumerate}
  \end{solution}

  \begin{exercise}[Zorich 2.4.4]
    Show that
    \begin{enumerate}
      \item the set of increasing sequences of natural numbers has the same cardinality as the set of fractions of the form $0.\alpha_1 \alpha_2 \ldots$ 
      \item the set of all subsets of countable set has cardinality of the continuum
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Listed. 
    \begin{enumerate}
      \item Given a sequence of increasing naturals $S = (n_1, n_2, \ldots)$, we can define a binary expansion $0.\alpha_1 \alpha_2 \ldots$ where $\alpha_i = 1$ if and only if $i \in \mathbb{N}$ is in $S$ and $\alpha_i = 0$ if not. This is clearly a bijection. 
      \item The set of all segments of increasing natural is equipotent with $2^\mathbb{N}$, since the elements of each sequence define a subset of $\mathbb{N}$. Cantor's diagonalization argument proves that the set of infinite binary expansions is uncountable, and by (a), this proves that $2^\mathbb{N}$ is uncountable. 
    \end{enumerate}
    This is very interesting since $\mathbb{N} \simeq \mathbb{R}$, but $2^{\mathbb{N}} \simeq \mathbb{R}$, and the set of all infinite $q$-ary expansions is equipotent to $\mathbb{R}$ too. 
  \end{solution}

  \begin{exercise}[Zorich 2.4.5]
    Show that 
    \begin{enumerate}
      \item the set $\mathcal{P}(X)$ of subsets of a set $X$ has the same cardinality as the set of all functions $f: X \longrightarrow \{0, 1\}$. 
      \item for a finite set $X$ of $n$ elements, $\card{\mathcal{P}(X)} = 2^n$ 
      \item one can write $\card{\mathcal{P}(X)} = 2^{\text{card}{X}}$, which implies $\text{card}\, \mathcal{P}(\mathbb{N}) = 2^{\card \mathbb{N}} = \card \mathbb{R}$ 
      \item for any set $X$, $\card{X} < 2^{\card{X}}$
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Listed. 
    \begin{enumerate}
      \item An element $Y \in \mathcal{P}(X)$ is a subset of $X$ by definition. Letting 
      \[f_Y (x) = \begin{cases} 0 & \text{ if } x \not\in Y \\
      1 & \text{ if } x \in Y \end{cases}\]
      we can construct the bijective map $Y \mapsto f_Y$. 
      \item We can prove this using the identity (which can be proved using induction) 
        \[\sum_{k=0}^n \binom{n}{k} = 2^n \]
      \item Let $F(X; \, \{0, 1\})$ be the set of all binary valued functions from $X$ to $\{0, 1\}$. From (a), $\card{\mathcal{P}(X)} \simeq F(X; \, \{0, 1\})$. Each binary-valued function $f$ is determined by the assignment $f(x)$ for each $x \in X$. Since $f(x)$ has two possible values, the assignment of $f(x)$ for all $x \in X$ has $\{0, 1\}^{\card{X}}$ possible choices. This gives another bijection $F(X; \, \{0, 1\}) \simeq \{0, 1\}^{\card{X}}$, so 
      \[\mathcal{P}(X) \simeq \{0, 1\}^{\card{X}} \implies \card{\mathcal{P}(x)} = \card(\{0, 1\}^{\card{X}}) = 2^{\card{X}} \]
      \item If $X$ is finite, then letting $n = \card{X}$, we can simply prove $n < 2^n$ by induction (which we will not do here). If $X$ is countable, then $\mathcal{P}(X)$ is uncountable (from 2.4.4) and so using (c), 
      \[\card{X} = \card{\mathbb{N}} < \card{\mathbb{R}} = \card{\mathcal{P}(X)} = 2^{\card{X}}\]
      For uncountable sets (and for the two cases mentioned above), we can use Cantor's theorem, which states that $\card{X} < \card{\mathcal{P}(X)}$, and so using (c), we have $\card{X} < \card{\mathcal{P}(X)} = 2^{\card{X}}$. 
    \end{enumerate}
  \end{solution}

  \begin{exercise}[Zorich 2.4.6]
    Let $X_1, \ldots, X_m$ be a finite system of finite sets. Show that 
    \begin{align*}
      \card \bigg( \bigcup_{i=1}^m X_i \bigg) & = \sum_{i_1} \card{X_{i_1}} - \sum_{i_1 < i_2} \card(X_{i_1} \cap X_{i_2}) + \ldots \\
      & \sum_{i_1 < i_2 < i_3} \card(X_{i_1} \cap X_{i_2} \cap X_{i_3}) - \ldots + (-1)^{m-1} \card (X_1 \cap \ldots \cap X_m) \\
      & = \sum_{k=1}^{m} \sum_{1 \leq i_1 \ldots i_k \leq m} (-1)^{k-1} \, \card\bigg( \bigcap_{j=1}^k X_{i_j} \bigg)
    \end{align*}
  \end{exercise}
  \begin{solution}
    Ignoring Russell's paradox (defining the universe set of all sets), we can use the commutative, associative, and distributive properties of $\cup, \cap$ on the algebra of sets. We prove using induction on $m$. For $m=1$, we trivially have $\card{X_1} = \card{X_1}$, and for $m = 2$, we claim 
    \[\card( X_1 \cup X_2) = \card(X_1) + \card(X_2) - \card(X_1 \cap X_2)\]
    $X_1$ and $X_2 \setminus X_1$ are clearly exclusive sets by definition, with $X_1 \cup X_2 = X_1 \cup (X_2 \setminus X_1)$, so 
    \[\card(X_1 \cup X_2) = \card\big( X_1 \cup (X_2 \setminus X_1) \big) = \card(X_1) + \card(X_2 \setminus X_1) \tag{2} \label{CardOne}\]
    By definition, the set $X_2 \setminus X_1$ and $X_1 \cap X_2$ are disjoint and satisfies $X_2 = (X_2 \setminus X_1) \cup (X_1 \cap X_2)$ (also by definition), so 
    \[\card(X_2) = \card(X_2 \setminus X_1) + \card(X_1 \cap X_2) \tag{3} \label{CardTwo}\]
    and substituting \eqref{CardTwo} into \eqref{CardOne} gives the claim for $m=2$. Assuming that the claim is satisfied for some $m$, we have 
    \begin{align*}
      \card \bigg( \bigcup_{i=1}^{m+1} X_i \bigg) & = \card \bigg( \bigg[ \bigcup_{i=1}^m X_i \bigg] \cup X_{m+1} \bigg) & \\
      & = \card \bigg( \bigcup_{i=1}^k X_i \bigg) + \card(X_{m+1}) - \card\bigg( \bigg[ \bigcup_{i=1}^m X_i\bigg] \cap X_{m+1} \bigg) & (\text{claim for } m=2) \\
      & = \card \bigg( \bigcup_{i=1}^k X_i \bigg) + \card(X_{m+1}) - \card\bigg(\bigcup_{i=1}^m (X_i \cap X_{m+1}) \bigg) & (\text{distributive prop.})  \\
      & = \sum_{k=1}^m \sum_{1 \leq i_1 \ldots i_k \leq m} (-1)^{k-1} \card\bigg( \bigcap_{j=1}^k X_{i_j} \bigg) + \card(X_{m+1}) \\ 
      & \;\;\;\;\;\; - \sum_{k=1}^m \sum_{1 \leq i_1 \ldots i_k \leq m} (-1)^{k-1} \card\bigg( \bigcap_{j=1}^k (X_{i_j} \cap X_{m+1}) \bigg) 
    \end{align*}
    With a bit of thought, we can see that the $k$th term of the second summation contributes to adding another term to the $k+1$th summation term of the first. Therefore, we must try to shift the summation over by $1$ index. Let us simplify this by taking the summations and extracting the first and last term, respectively. We have 
    \begin{align*}
    \sum_{k=1}^m \sum_{1 \leq i_1 \ldots i_k \leq m} (-1)^{k-1} \card\bigg( \bigcap_{j=1}^k X_{i_j} \bigg) &= \sum_{1 \leq i_1 \ldots i_k \leq m} \card(X_{i_1}) \\
    &+ \sum_{k=2}^m \sum_{1 \leq i_1 \ldots i_k \leq m} (-1)^{k-1} \card\bigg( \bigcap_{j=1}^k X_{i_j} \bigg)
    \end{align*}
    and 
    \begin{align*}
      \sum_{k=1}^m \sum_{1 \leq i_1 \ldots i_k \leq m} & (-1)^{k-1} \card\bigg( \bigcap_{j=1}^k (X_{i_j} \cap X_{m+1}) \bigg) \\
      & = \sum_{k=1}^m \sum_{1 \leq i_1 \ldots i_k \leq m} (-1)^{k-1} \card\bigg( \bigg[ \bigcap_{j=1}^k X_{i_j} \bigg] \cap X_{m+1} \bigg) \\
      & =  \sum_{k=1}^{m-1} \sum_{1 \leq i_1 \ldots i_k \leq m} (-1)^{k-1} \card\bigg( \bigcap_{j=1}^k (X_{i_j} \cap X_{m+1}) \bigg) + (-1)^{m-1} \card\bigg( \bigcap_{j=1}^{m+1} X_j \bigg) \\
      & = \sum_{k=2}^{m} \sum_{1 \leq i_1 \ldots i_k \leq m} (-1)^{k-2} \card\bigg( \bigcap_{j=1}^{k-1} (X_{i_j} \cap X_{m+1}) \bigg) + (-1)^{m-1} \card\bigg( \bigcap_{j=1}^{m+1} X_j \bigg)
    \end{align*}
    So subtracting the summations gives 
    \begin{align*}
      \sum_{k=1}^m & \sum_{1 \leq i_1 \ldots i_k \leq m} (-1)^{k-1} \card\bigg( \bigcap_{j=1}^k X_{i_j} \bigg) - \sum_{k=1}^m \sum_{1 \leq i_1 \ldots i_k \leq m} (-1)^{k-1} \card\bigg( \bigcap_{j=1}^k (X_{i_j} \cap X_{m+1}) \bigg) + |X_{m+1}| \\
      & = \sum_{1 \leq i_1 \ldots i_k \leq m} \card(x_i) + \sum_{k=2}^m \sum_{1 \leq i_1 \ldots i_k \leq m} (-1)^{k-1} \card\bigg( \bigcap_{j=1}^k X_{i_j} \bigg) + \card(X_{m+1})\\
      & \;\;\;\;\; + \sum_{k=2}^{m} \sum_{1 \leq i_1 \ldots i_k \leq m} (-1)^{k-1} \card\bigg( \bigcap_{j=1}^{k-1} (X_{i_j} \cap X_{m+1}) \bigg) + (-1)^{m} \card\bigg( \bigcap_{j=1}^{m+1} X_j \bigg) \\
      & = \sum_{1 \leq i_1 \ldots i_k \leq m+1} \card(X_i) + \sum_{k=2}^m \sum_{1 \leq i_1 \ldots i_k \leq m} (-1)^{k-1} \Bigg[ \card\bigg( \bigcap_{j=1}^k X_{i_j} \bigg) \\
      & +  \card\bigg( \bigg[ \bigcap_{j=1}^{k-1} X_{i_j} \bigg] \cap X_{m+1} \bigg) \Bigg] + (-1)^{m} \card\bigg( \bigcap_{j=1}^{m+1} X_j \bigg) 
    \end{align*}
    and since the set of sequences of $k$ terms bounded by $m+1$ (of form $1 \leq i_1 \ldots i_k \leq m+1$) is the set of sequences of $k$ terms bounded by $m$ (of form $1 \leq i_1 \ldots i_k \leq m$) unioned with the set of sequences of $k$ terms with max element $m+1$ (of form $1 \leq i_1 \ldots i_k = m+1$), we have 
    \[\sum_{1 \leq i_1 \ldots i_k \leq m} (-1)^{k-1} \Bigg[ \card\bigg( \bigcap_{j=1}^k X_{i_j} \bigg) +  \card\bigg( \bigg[ \bigcap_{j=1}^{k-1} X_{i_j} \bigg] \cap X_{m+1} \bigg) \Bigg] = \sum_{1 \leq i_1 \ldots i_k \leq m+1} \card\bigg( \bigcap_{j=1}^k X_{i_j} \bigg)\]
    and therefore, substituting the above and observing that the independent terms are the first and last terms of the summation gives 
    \begin{align*}
      \card \bigg( \bigcup_{i=1}^{m+1} X_i \bigg) & = \sum_{1 \leq i_1 \ldots i_k \leq m+1} \card(X_i) + \sum_{k=2}^m \sum_{1 \leq i_1 \ldots i_k \leq m+1} (-1)^{k-1} \card\bigg( \bigcap_{j=1}^k X_{i_j} \bigg) \\ 
      & \;\;\;\;\;\;\;\;\;\; + \ldots + (-1)^{m} \card\bigg( \bigcap_{j=1}^{m+1} X_j \bigg) \\
      & = \sum_{k=1}^{m+1} \sum_{1 \leq i_1 \ldots i_k \leq m+1} (-1)^{k-1} \card\bigg( \bigcap_{j=1}^k X_{i_j} \bigg) 
    \end{align*}
  \end{solution}

  \begin{exercise}[Zorich 2.4.7]
    On the closed interval $[0, 1] \subset \mathbb{R}$, describe the sets of numbers $x \in [0, 1]$ whose ternary representation $x = 0.\alpha_1 \alpha_2 \ldots $, $\alpha_i \in \{0, 1, 2\}$ has the property. 
    \begin{enumerate}
      \item $\alpha_1 \neq 1$
      \item $\alpha_1 \neq 1$ and $\alpha_2 \neq 1$ 
      \item For all $i \in \mathbb{N}$, $\alpha_i \neq 1$ (the Cantor set)
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Listed. 
    \begin{enumerate}
      \item $[0, \frac{1}{3}) \cup [\frac{2}{3}, 1)$ 
      \item $[0, \frac{1}{9}) \cup [\frac{2}{9}, \frac{3}{9}) \cup [\frac{6}{9}, \frac{7}{9}) \cup [\frac{8}{9}, 1)$
      \item Made by recursively removing the middle third of every partitioned intervals. 
    \end{enumerate}
  \end{solution}

  \begin{exercise}[Zorich 2.4.8]
    Show that 
    \begin{enumerate}
      \item the set of numbers $x \in [0, 1]$ whose ternary representation does not contain $1$ has the same cardinality as the set of all numbers whose binary representation has the form $0.\beta_1 \beta_2 \ldots$ 
      \item the Cantor set has the same cardinality as the closed interval $[0, 1]$ 
    \end{enumerate}
  \end{exercise}
  \begin{solution}
    Listed. 
    \begin{enumerate}
      \item We can define a bijection $0.\alpha_1 \alpha_2 \ldots \mapsto 0.\beta_1 \beta_2 \ldots$ as $\alpha_i = 0 \iff \beta_i = 0$ and $\alpha_i = 1 \iff \beta_i = 2$. 
      \item The map above defines a bijection between the Cantor set and the set of all infinite binary expansions in $[0, 1]$, which is uncountable by Cantor's diagonalization theorem. 
    \end{enumerate}
  \end{solution}

