Cooking your own meals and planning out your diet is hard. You're probably lacking in motivation, experience, money, or time to plan your own diet. You want to develop a system that hits all of these \textit{essential} points. 
\begin{enumerate}
  \item \textbf{Health}. Your food should be healthy. The value of cooking yourself comes from being able to decide what to put in your food. If you cook crap food, then even if it is cheap, you can still get instant processed food for probably cheaper, which negates the effort you put into making meals.  

  \item \textbf{Taste}. Your diet should be tasty. You're stressed from the rest of your life already, so you shouldn't endure more from having to eat. 


  \item \textbf{Cost}. It should cost you much less than eating out. If your meals comprise of \$10 worth of ingredients each, then you might as well eat out (and you can still get relatively healthy meals with this budget) without having to go through the effort of eating food. Ideally, I would spend \$1 to \$5 a meal. 

  \item \textbf{Time}. It should be take little time to prepare meals, relative to how may servings you make. I personally prepare a lot of meals at the beginning of the semester, preferably 2-4 months ahead. Therefore, before classes start, you can move in, do a ton of grocery shopping, and make these meals all in one go. 

\end{enumerate}
Here are some nonessential---yet perhaps important---points. I find the first point essential for me but the second to be not. 
\begin{enumerate}
  \item \textbf{Sustainable}. The meal can be stored for long periods of time and can be reheated. 
  \item \textbf{Variety}. You should have a variety of meals to avoid eating the same thing every day. 
\end{enumerate}

Here is a guideline I developed over the years. It is divided up into 3 categories. 
\begin{enumerate}
  \item \textbf{Nutrition}. What macros/micros are important to you? You might want a heavy protein diet, or a low carb diet. You also want to be aware of which vitamins and minerals you are taking.  
  \item \textbf{Groceries}. Once you have a nutrition plan in mind, you should choose the appropriate food.  
  \item \textbf{Cooking}. You should know how to prepare the food. 
\end{enumerate}

