\section{Compactness} 

  Now we move onto another pillar of topology: compactness. I personally think compactness is the hardest conceptually, so I will provide a lot of explanation on it here. The thing to keep in mind when talking about compactness is that it describes a space being finite or not being able to escape to infinity.  
  
  \begin{definition}[Covers]
    A collection $\C$ of subsets of a space $X$ is said to \textbf{cover} $X$, or to be a \textbf{covering} of $X$, if the union of the elements of $\C$ is equal to $X$. It is called an \textbf{open covering} of $X$ if its elements are open subsets of $X$. 
  \end{definition}

  \begin{definition}[Compact Space]
    A topological space $X$ is said to be \textbf{compact} if every open covering of $X$ contains a finite subcovering (i.e. a finite collection of subcovers) of $X$. If $K \subset X$ is a compact topological space in the subspace topology, then $K$ is said to be a \textbf{compact subspace} of $X$. 
  \end{definition}

  Note that compactness is a property of a topological space in its entirety. This is opposed to considering open and closed sets, which exist as a subset of a topological space. While openness behaves differently depending on its embedding space, compactness stays constant. Therefore, we don't have to worry about talking about which space a compact set is embedded in. 
  However, we will provide a definition for which it makes sense to talk about compact \textit{subsets} of a space. 

  \begin{definition}[Compact Subspaces]
    Given topological space $X$ and $A \subset Y$, we call $A$ a \textbf{compact subspace} of $X$ if either of the two equivalent conditions are met. 
    \begin{enumerate}
      \item $A$ is a compact space with respect to the subspace topology endowed from $X$. 
      \item If every $X$-open cover of $A$ has a finite subcover of $A$. 
    \end{enumerate}
  \end{definition}
  \begin{proof}
    We can prove bidirectionally. 
    \begin{enumerate}
      \item Suppose that $K$ is compact in $X$. Then given any open cover $\{U_\alpha\}_\alpha$ of $K$, there exists a finite subcover $\{U_i\}_{i}$. Now let there exist an open cover $\{V_\alpha\}$ in $Y$, but every $V_\alpha = U_\alpha \cap Y$ for some $U_\alpha$ open in $X$. Therefore, we can take the finite subcover $\{V_i = V_i \cap Y\}_i$. 

      \item Suppose that $K$ is compact in $Y$. Then given any open cover $\{V_\alpha\}$ of $K$, there exists a finite subcover $\{V_i\}_i$. Now let there exist an open cover $\{U_\alpha\}$ in $X$. Then we set $\{V_\alpha = U_\alpha \cap Y\}_\alpha$, which has a finite subcover $\{V_i = U_i \cap Y\}$, and therefore we can take $\{U_i\}$ as our finite subcover in $X$. 
    \end{enumerate}
  \end{proof}
   
  The concept of compactness does not seem intuitive at first glance. The reason why compactness is such an important property for a space to have is because $X$ being compact tells us that we can \textit{always} analyze the entire $X$ using a finite union of open sets, which can simplify the space greatly. That is, it a measure of finiteness of a space. You should realize that it is much easier to prove that a space is \textit{not} compact, since all we have to do is find a \textit{single} covering that doesn't contain a finite subcovering. 

  \begin{example}[Non-Compact Spaces]
    We show some examples of non-compact spaces. 
    \begin{enumerate}
      \item $\mathbb{R}$ is not compact since given the open covering $\mathcal{F} = \{ (x, x + 1) \subset \mathbb{R} \mid x \in \mathbb{R} \}$, then if we assume that $\mathcal{F}$ has a finite subcovering, it is the case that $\max\{x_1, x _2, \ldots, x_n \} + 1 \not\in \mathcal{F}$, so this is not a covering. 
      \item For $\mathbb{R}$, you can also get a nested covering $\{(-r, +r) \mid r > 0\}$, which also has no finite subcovering. 

      \item Take $X = (1, 1)$ and take $\mathcal{F} = \{ (\frac{1}{n}, 1] \mid n \in \mathbb{N} \}$, which is a cover but has no finite subcover. 
    \end{enumerate}
  \end{example}

  \begin{example}[Open Square is Not Compact]
    The subset $Y \equiv (0,1) \times (0,1) \subset \mathbb{R}^2$ is not compact. That is, we can choose to cover the subspace by the finite union of open sets. 
    \begin{equation}
      [0,1]^2 \subset \bigcup_{k=0}^\infty \Big( \frac{2^k - 1}{2^k}, \frac{2^{k+1} - 1}{2^{k+1}} \Big) \times (0,1)
    \end{equation}

    \begin{figure}[H]
      \centering 
      \begin{tikzpicture}[scale=2]
        \draw[dashed] (0,0) rectangle (1,1);
        \draw[dashed, fill=lightgray] (0,0) rectangle (0.5,1);
        \draw[dashed, fill=lightgray] (0.5,0) rectangle (0.75,1);
        \draw[dashed, fill=lightgray] (0.75,0) rectangle (0.875,1);
      \end{tikzpicture}
      \caption{We show the first three elements of the infinite union that covers the open square. }
      \label{fig:closed_square_compact}
    \end{figure}
  \end{example}

  According to Terry Tao, a compact set is ``small,'' in the sense that it is easy to deal with. While this may sound counterintuitive at first, since $[0,1]$ is considered compact while $(0,1)$, a subset of $[0,1]$, is considered noncompact. More generally, a set that is compact may be large in area and complicated, but the fact that it is compact means we can interact with it in a finite way using open sets, the building blocks of topology. That finite collection of open sets makes it possible to account for all the points in a set in a finite way. This is easily noticed, since functions defined over compact sets have more controlled behavior than those defined over noncompact sets. Similarly, classifying noncompact spaces are more difficult and less satisfying. 

  In general, it's pretty easy to prove that a set is not compact. We just need to find one example of an open cover that does not have a finite subcover. To prove that set \textit{is} compact, we must show that for \textit{every} open cover, we can get a finite subcover. 

  \begin{lemma}[Finite Sets]
    A finite set $S$ is compact with respect to any topology. 
  \end{lemma}
  \begin{proof}
    This is because there are only a finite number of open sets to begin with, so every open cover is a finite cover. 
  \end{proof}

  \begin{lemma}
    A finite union of compact sets is compact. 
  \end{lemma}
  \begin{proof}
    It suffices to prove for two sets $A, B$ by induction. Take an arbitrary cover $\mathscr{L}$ of $A \cup B$. Then $\mathscr{L}$ is a cover of $A$, so it has a finite subcover $\mathscr{F} \subset \mathscr{L}$. It is also a cover of $B$, so it has a finite subcover $\mathscr{G} \subset \mathscr{L}$. Therefore, $\mathscr{F} \cup \mathscr{G} \subset \mathscr{L}$ is a cover of $A \cup B$, and since it is the union of finite covers, it is finite. 
  \end{proof}

  The first property we will state---similar to that for connected spaces---is that continuous images of compact spaces is compact, which establishes compactness as a topological property. 

  \begin{theorem}[Continuous Images of Compact Sets are Compact]
    The image of a compact space under a continuous map is compact.
  \end{theorem}
  \begin{proof}
    Let $f: X \rightarrow Y$ be continuous, and let $X$ be compact. Let $\mathcal{C}$ be a covering of the set $f(X)$ by sets open in $Y$. Then, the preimage of these sets is the collection
    \begin{equation}
      \{f^{-1}(\mathcal{A}) \; | \; \mathcal{A} \in \mathcal{C}\}
    \end{equation}
    which clearly covers $X$. But since $X$ is compact, a finite number of them, say
    \begin{equation}
      f^{-1} (\mathcal{A}_1), f^{-1} (\mathcal{A}_2), \ldots , f^{-1} (\mathcal{A}_n)
    \end{equation}
    covers $X \implies \mathcal{A}_1, \mathcal{A}_2, \ldots, \mathcal{A}_n$ covers $f(X)$. 
  \end{proof}

  \begin{corollary}[Compactness is a Topological Property]
    Let $X \cong Y$. Then $X$ compact $\iff$ $Y$ compact. 
  \end{corollary}

\subsection{Properties on Compact, Closed, and Hausdorff Spaces}

  The three most general statements about compact sets are that closed subsets of compact sets are compact, continuous images of compact sets are compact, and a nested sequence of closed nonempty sets in a compact space is nonempty. 

  \begin{theorem}[Closed Subsets of Compact Sets are Compact]
    Every closed subset of a compact space is compact. 
  \end{theorem}
  \begin{proof}
    This proof is quite trivial. Let $Y$ be a closed subset of compact space $X$. Given a covering $\mathcal{C}$ of $Y$ by sets open in $X$, let us form an open covering $\B$ of $X$ by adjoining to $\mathcal{C}$ the single open set $X \setminus Y$. Then, we an see that both $\B$ and $\mathcal{C} \cup (X \setminus Y)$ covers $X$. 
    \begin{equation}
      \B = \mathcal{C} \cup (X \setminus Y)
    \end{equation}
    Since $\B$ is finite, the right hand side must also be expressible as a finite union. Looking through $\B$, we can throw away all the open sets that are entirely in $X \setminus Y$. What remains is a finite covering of $Y$. 
  \end{proof}

  \begin{corollary}
    If $F$ is closed and $K$ is compact, then $F \cap K$ is compact. 
  \end{corollary}

  A natural question to ask is: what about the converse? It turns out that Hausdorff provides the sufficient properties. 

  \begin{theorem}[Compact Subsets of Hausdorff Spaces are Closed]
    If $X$ is Hausdorff and $A \subset X$ is compact. Then $A$ is closed in $X$. 
  \end{theorem}
  \begin{proof}
    We wish to show that $X \setminus A$ is open. Let $x \in X \setminus A$. Then, for each point $a \in A$, we can choose disjoint neighborhoods $U_a \ni x$ and $V_a \ni a$ (using the Hausdorff condition). The collection 
    \begin{equation}
      \{V_a \mid a \in A\}
    \end{equation}
    is an open covering of $A$. Since $A$ is compact, there must exist a finite subcover $V_1, V_2, \ldots , V_n$. Therefore, $\cup_{i=1}^{n} V_i$ contains $Y$ and is disjoint from the intersection of open neighborhoods of $x$
    \begin{equation}
      U \coloneqq \bigcap_{i=1}^n U_i
    \end{equation}
    Therefore, $U$ is an open neighborhood of $x_0$, disjoint from $A \implies X \setminus A$ is open $\implies A$ is closed.
  \end{proof}

  But is it minimal? It turns out that if we don't have Hausdorff, then a compact subset is not necessarily closed. 

  \begin{example}[Line with 2 Origins]
    Consider the line with 2 origins, $L = (-\infty, 0) \cup (0, +\infty) \cup \{r, s\}$. Then $A = \{r\} \cup (0, 1] = [0, 1]$ is not compact, but it is not closed since $r$ is also a limit point. 
  \end{example}

  \begin{lemma}[Compact Subsets and Points Can Be Separated in Hausdorff Space]
    If $Y$ is a compact subset of a Hausdorff space $X$ and $x$ is not in $Y$, then there exist disjoint open sets $U$ and $V$ of $X$ containing $x$ and $Y$, respectively. 

    \begin{figure}[H]
      \centering 
      \begin{tikzpicture}[scale=0.7]
        \draw (0,0) ellipse (5 and 2);
        \draw[dashed] (-3,0) circle (1);
        \draw[fill] (-3,0) circle (0.05);
        \node[left] at (-3,0) {$x$};
        \draw (3,0) circle (1); 
        \node[below] at (3,0.8){$Y$};
        \node[below] at (3, 0.4){compact};
        \node[above] at (-3,0.3) {$U$};
        \draw[dashed] (3,0) circle (1.3); 
        \node at (2,1) {$V$};
      \end{tikzpicture}
      \label{fig:hausdorff_compact}
    \end{figure}
  \end{lemma}

  \begin{theorem}[Continuous Mappings from Compact to Hausdorff Spaces]
    Let $X$ be compact, $Y$ Hausdorff, and $f: X \to Y$ be a continuous map. Then, $f$ is a closed map, and moreover, 
    \begin{enumerate}
      \item $f$ injective $\implies$ $f$ is an embedding. 
      \item $f$ surjective $\implies$ $f$ is a quotient map. 
      \item $f$ bijective $\implies$ $f$ is a homeomorphism. 
    \end{enumerate}
  \end{theorem}
  \begin{proof} 
    We prove the four statements. 
    \begin{enumerate}
      \item To see that $f$ is a closed map, let $C \subset X$ be closed. Then $C$ is compact, and so $f(C)$ is also compact. But since $Y$ is Hausdorff, $f(C)$ is closed. 
      \item 
    \end{enumerate}
  \end{proof}

  The previous theorem is nice in that we can prove a lot of the nice properties of topological spaces without doing much work. Note that compactness and Hausdorff goes very well together in that you can get a lot more out of a space when they are together. 

\subsection{Nested Sequence Lemma}

  Now lets prove something more set-theoretic about compact sets, namely its cardinality. 

  \begin{lemma}[Nested Sequence Lemma]
    If $X$ is compact, then for any sequence of closed, nonempty sets $C_1 \supset C_2 \supset \ldots$, we must have 
    \begin{equation}
      \bigcup_{n=0}^{\infty} C_n \neq \emptyset
    \end{equation}
  \end{lemma}

  \begin{definition}[Isolated Point]
    A point $x$ is called an \textbf{isolated point} if $\{x\}$ is an open set. 
  \end{definition}

  \begin{theorem}[Uncountability of Compact Hausdorff Spaces]
    If $X$ is a nonempty, compact, Hausdorff space with no isolated points, then $X$ is uncountable. 
  \end{theorem}
  \begin{proof}
    
  \end{proof}

  Thus, any closed interval in $\mathbb{R}$ is uncountable. We didn't need to use decimal expansions to prove it. 

\subsection{Finite Products and Intersections of of Compact Spaces}

  We now introduce a useful lemma that will come around in many future cases. 

  \begin{lemma}[Tube Lemma]
    Consider the product space $X \times Y$, where $Y$ is compact. If $N$ is an open set $X \times Y$ containing the slice $x_0 \times Y$ of $X \times Y$, then $N$ contains some tube $W \times Y$ about $x_0 \times Y$, where $W$ is a neighborhood of $x_0$ in $X$. 

    \begin{figure}[H]
      \centering 
      \begin{tikzpicture}
        \draw[dashed, fill=lightgray] (3.5,0) rectangle (4.3,5);
        \draw (0,0)--(0,5);
        \draw[<->] (-1,0)--(8,0);
        \node[right] at (8,0) {$X$};
        \node[left] at (0,2.5) {$Y$};
        \draw[<->] (-1,5)--(8,5);
        \draw[thick] (4,0)--(4,5);
        \draw[dashed] (3,0) rectangle (4.8,1);
        \draw[dashed] (3.5,0.6) rectangle (4.5,2);
        \draw[dashed] (3.1,1.8) rectangle (6,3.5);
        \draw[dashed] (2.5,3.4) rectangle (4.3,4.7);
        \draw[dashed] (2.8,4) rectangle (5, 5);
        \draw[thick] (3.5,0)--(4.3,0);
        \node[above] at (4,5) {$W$};
        \node[below] at (4,0) {$x_0$};
        \draw[fill] (4,0) circle (0.05);
        \draw[dashed] plot [smooth] coordinates {(2.5,0) (2.3, 1) (2,3) (2.2,5)};
        \draw[dashed] plot [smooth] coordinates {(5.4,0) (6.5,2) (6.5,4) (5.7,5)};
        \node[left] at (6,4.5) {$N$};
      \end{tikzpicture}
      \label{fig:tube_lemma}
    \end{figure}
  \end{lemma}
  \begin{proof}
    Let us cover $x_0 \times Y$ by basis elements $U \times V$ (for the topology of $X \times Y$) lying in $N$. The space $x_0$ is compact since it is homeomorphic to $Y \implies$ we can cover $x_0 \times Y$ by finitely such basis elements
    \begin{equation}
      U_1 \times V_1, U_2 \times V_2, \ldots , U_n \times V_n
    \end{equation}
    Without loss of generality, we can assume that each $U_i \times V_i$ has a nontrivial intersection with $x_0 \times Y$, since otherwise, it would be superfluous. Now, we define the intersection of all the open neighborhoods of $x_0$ in $X$ of the basis elements $U_i \times V_i$. That is, let
    \begin{equation}
      W \equiv \bigcup_{i=1}^n U_i
    \end{equation}
    As an intersection of open sets, $W$ is also open containing $x_0$. With this well-defined tube $W \times Y$, we claim that it is entirely contained within $N$. That is, given a point $x \times y \in W \times Y$, consider the corresponding point $x_0 \times y$ that is the image of the projection of $x\times y$ onto $x_0 \times Y$. Clearly, $x_0 \times y$ belongs to some $U_k \times V_k$ (for some $k$) $\implies y \in V_k$. Since $x \in W$, $x$ is clearly in $U_k$, meaning that $x \times y \in U_k \times V_k \subset N$, as desired. 
  \end{proof}

  \begin{theorem}[Finite Products]
    The product of finitely many compact spaces is compact. 
  \end{theorem}
  \begin{proof}
    Using induction, it suffices to prove that the product of 2 compact spaces is compact. Let $X$ and $Y$ be compact spaces. By the tube lemma, for each $x \in X$, there exists a neighborhood $W_x$ of $x$ such that the tube $W_x \times Y$ can be covered with finitely (by compactness of $Y$) many open sets in $X \times Y$. The collection of all neighborhoods $W_x$ is an open covering of $X$. By compactness of $X$, there exists a finite subcollection
    \begin{equation}
      W_1, W_2, \ldots , W_k
    \end{equation}
    covering $X$. The finite union of the tubes 
    \begin{equation}
      \bigcup_{i=1}^k W_i \times Y
    \end{equation}
    clearly covers $X \times Y$, meaning that $X \times Y$ is compact. 
  \end{proof}

  \begin{definition}[Finite Intersection Condition]
    A collection $\mathcal{C}$ of subsets of $X$ is said to satisfy the \textbf{finite intersection condition} if for every finite subcollection 
    \begin{equation}
      \{\mathcal{C}_1, \mathcal{C}_2, \ldots , \mathcal{C}_n\}
    \end{equation}
    of $\mathcal{C}$, the intersection
    \begin{equation}
      \bigcap_{i=1}^n \mathcal{C}_i
    \end{equation}
    is nonempty. 
  \end{definition}

  Clearly, the empty sets cannot below to any collection with the finite intersection property. Additionally, the condition is trivially satisfied if the intersection over the entire collection is non-empty or if the collection is nested. However, here is one example that does satisfy the finite intersection condition. 

  \begin{example}
    Let $X = (0,1)$ and for each positive integer $i$, $X_i$ is the set of elements of $X$ having a decimal expansion with digit $0$ in the $i$th decimal place. Then, any finite intersection of $X_i$'s is nonempty, but the intersection of all $X_i$ for $i \in \mathbb{N}$ is empty, since no element of $(0,1)$ has all zero digits. 
  \end{example}

  Here is an analogous example to the previous one. 

  \begin{example}
    In the space $\mathbb{R}$, let us define $C_i \equiv \mathbb{R} \setminus \{i\}$. That is, $C_i$ is $\mathbb{R}$ missing a point at $i$. Then, the collection of all $C_i$'s does satisfy the finite intersection condition. We show below the finite intersection of the five subsets $C_0, C_1, C_2, C_3, C_4$. 

    \begin{figure}[H]
      \centering 
      \begin{tikzpicture}[scale=0.8]
        \draw[<->] (-1,0)--(6,0);
        \draw[<->] (-1,1)--(6,1);
        \draw[<->] (-1,2)--(6,2);
        \draw[<->] (-1,3)--(6,3);
        \draw[<->] (-1,4)--(6,4);
        \draw[<->] (-1,-1)--(6,-1);
        \draw[fill=white] (0,0) circle (0.07); 
        \draw[fill=white] (1,1) circle (0.07); 
        \draw[fill=white] (2,2) circle (0.07); 
        \draw[fill=white] (3,3) circle (0.07); 
        \draw[fill=white] (4,4) circle (0.07); 
        \draw[fill=white] (0,-1) circle (0.07); 
        \draw[fill=white] (1,-1) circle (0.07); 
        \draw[fill=white] (2,-1) circle (0.07); 
        \draw[fill=white] (3,-1) circle (0.07); 
        \draw[fill=white] (4,-1) circle (0.07); 
        \node[right] at (6,-1) {$\bigcap_{i=0}^4 C_i$};
        \node[right] at (6,0) {$C_0$};
        \node[right] at (6,1) {$C_1$};
        \node[right] at (6,2) {$C_2$};
        \node[right] at (6,3) {$C_3$};
        \node[right] at (6,4) {$C_4$};
      \end{tikzpicture}
      \label{fig:ex}
    \end{figure}
  \end{example}

  \begin{theorem}
    Let $X$ be a topological space. Then $x$ is compact if and only if for any collection $\mathcal{C}$ of closed sets in $X$ satisfying the finite intersection condition, the intersection 
    \begin{equation}
      \bigcap_{C \in \mathcal{C}} C
    \end{equation}
    of all the elements of $\mathcal{C}$ is nonempty. 
  \end{theorem}
  \begin{proof}
    Given a collection $S$ fo subets of $X$, let 
    \begin{equation}
      \mathcal{C} \equiv \{X \setminus A \; | \; A \in S\}
    \end{equation}
    be the collection of their complements. Then, the following statements hold 
    \begin{enumerate}
      \item $S$ is a collection of open sets if and only if $\mathcal{C}$ is a collection of closed sets. 
      \item The collection $S$ covers $X$ if and only if the intersection 
      \begin{equation}
        \bigcap_{C \in \mathcal{C}} C
      \end{equation}
      of all the elements of $\mathcal{C}$ is empty. 

      \item The finite subcollection $\{A_1, A_2, \ldots , A_n\}$ of $S$ covers $X$ if and only if the intersection of the corresponding elements $C_i \equiv X \setminus A$ of $\mathcal{C}$ is empty. 
    \end{enumerate}
    Clearly, (1) is trivial, and (2) and (3) follows from DeMorgan's Law. 
    \begin{equation}
      X \setminus \bigcup_{\alpha \in J} A_\alpha = \bigcap_{\alpha \in J} (X \setminus A_\alpha)
    \end{equation}
    Using statement 3, the existence of a finite collection of closed sets $C$ in $X$ satisfying the finite intersection condition is equivalent to its complements (which are open sets) covering $X$, which is precisely the definition of compactness. 
  \end{proof}

  Clearly, the previous example in the real line $\mathbb{R}$ shows that $\mathbb{R}$ is indeed not compact. 

  \begin{corollary}
    The space $X$ is compact if and only if every collection $\C$ of subsets of $X$ satisfying the finite intersection condition, the intersection 
    \begin{equation}
      \bigcap_{A \in \C} \bar{A}
    \end{equation}
    of their closures is nonempty. 
  \end{corollary}

\subsection{Compact Subsets of Euclidean Space}

  In order to construct new compact spaces from old ones, we must prove compactness for a number of fundamental spaces. The real number line is a good starting point, but before we can fully appreciate the Heine-Borel theorem, we focus on the properties of closedness and boundedness. 

  \begin{theorem}[Compact Subsets of Metric Spaces are Bounded]
    Let $(X, d)$ be a metric space and $K \subset X$ be compact. Then, $K$ must be bounded. 
  \end{theorem} 
  \begin{proof}
    Given nonempty $K$, choose a point $x \in K$. Then, construct the open covering. 
    \begin{equation}
      \mathcal{F} = \{ B_n (x) \mid n \in \mathbb{N} \}
    \end{equation}
    which is indeed a covering since the naturals are unbounded in $\mathbb{R}$. Then by compactness there is a finite subcovering, and so there must be some maximum $n$. 
  \end{proof}

  \begin{lemma}[Closed Intervals are Compact]
    Any closed interval $[a, b] \subset \mathbb{R}$ is compact. 
  \end{lemma} 
  \begin{proof}
    Let $\mathcal{F}$ be a covering, and let 
    \begin{equation}
      S = \{x \in [a, b] \mid [a, x] \text{ is contained in finite subcollection of } \mathcal{F} \}
    \end{equation}
    We wish to show that $b \in S$. Since $S$ is bounded by $b$, let $s = \sup\{s\}$. We claim that $s \in S$. 
    \begin{enumerate}
      \item We know that $a \in S$ since $a$ is covered by at least one given set in $S$, so if $s = a$ we are done. Else if $s > a$, it must also be the case that $s \leq b$, so $s \in (a, b)$. We choose some open $U \in \mathcal{F}$ containing $S$, i.e. there exists $\epsilon > 0$ s.t. $(s - \epsilon, s] \subset U \implies s - \epsilon$ is not an upper bound for $S$. So $\exists x$ with $s - \epsilon < x \leq s$ with $s \in S$. So $[a, s]$ is contained in finitely many sets of $\mathcal{F}$, so $s \in S$. 

      \item Now we will claim $s = b$. Suppose otherwise, i.e. $s < b$. Then 
      \begin{equation}
        [a, s] \subset U_1 \cup \ldots \cup U_n \text{ for } U_i \in \mathcal{F}
      \end{equation}
      But $s \in U_i$ for some $i$, and for some $\epsilon$, $[s, s + \epsilon) \subset U$, and so $[a, s + \frac{\epsilon}{2}] \subset \cup_{i=1}^{n} U_i$, which implies $s + \frac{\epsilon}{2} \in S$, which contradicts that $s \in S$. So $s = b$. 
    \end{enumerate}
  \end{proof}

  Now we can extend this to $\mathbb{R}^n$. 

  \begin{theorem}[Heine-Borel Theorem]
    A subset $A$ of $\mathbb{R}^n$ is compact if and only if it is closed and bounded in the Euclidean metric $d$ or the square metric $\rho$. 
  \end{theorem}
  \begin{proof}
    The forwards direction is actually very easy. Since $A$ is compact in Hausdorff $\mathbb{R}^n$, it must be closed, and since $\mathbb{R}^n$ is a metric space, it must be bounded. So we are done. For the other way, we assume that a subset $A$ is closed and bounded. If $A$ is bounded, then $\exists R > 0$ s.t. $d(0, x) < R$ for all $x \in A$. So $A \subset [-R, R]^n = B$, where $B$ is compact and $A$ is a closed subset. Hence $A$ is compact as well. 

    Note that bounded w.r.t. metric $d$ is equivalent to being bounded w.r.t metric $\rho$. 
  \end{proof}

  The Heine-Borel theorem is like a reverse implication of the fact that compact sets must be closed and---if in a metric space---bounded. However, the double implication is not achieved when working in general metric spaces!

  \begin{example}[Closed and Bounded in Metric Space But Not Compact]
    Consider the space $X = \mathbb{Q}$ and $A = [1, \sqrt{2}] \subset X$. $A$ is closed and bounded, but not compact. 
  \end{example}

  \begin{example}
    The unit sphere $S^{n-1}$ and the closed ball $B^n$ in $\mathbb{R}^n$ are compact since they are closed and bounded. The set
    \begin{equation}
      A \equiv \{(x, \frac{1}{x}) \; | \; 0 < x \leq 1\}
    \end{equation}
    is closed in $\mathbb{R}^2$, but is not compact since it is not bounded. The set 
    \begin{equation}
      S \equiv \{(x, \sin{\frac{1}{x}}) \; | \; 0<x\leq 1\}
    \end{equation}
    is bounded in $\mathbb{R}^2$, but it is not compact since it is not closed. 
  \end{example}

  Now we present another theorem in analysis that is actually a topological property. 

  \begin{theorem}[Extreme Value Theorem]
    Let $X$ be compact and $f: X \rightarrow \mathbb{R}$ be continuous. Then $f$ attains both its maximum and minimum in $X$.  
  \end{theorem}
  \begin{proof}
    $f$ continuous must imply that $f(X)$ is compact in $\mathbb{R}$, which means that it is closed and bounded. Since it is bounded, it must have a least upper bound. Since it's closed, it must contain its least upper bound, and so attains its maximum. Similarly for minimum.
  \end{proof}

  The theorem says that $f: [a, b] \to \mathbb{R}$ is bounded, and that bound is \textit{realized} by something. Note that for function on noncompact $(-1, 1)$---say $f(x) = x$---it is bounded but the extrema are not achieved. It doesn't even need to be bounded, e.g. $f(x) = \frac{x}{1 - x^2}$. 

  \begin{theorem}[Uniform Continuity Theorem]
    Let $f: X \rightarrow Y$ be a continuous map of the compact metric space $(X,d_X)$ to the metric space $(Y, d_Y$. Then, $f$ is uniformly continuous. That is, given $\epsilon > 0$, there exists a $\delta > 0$ such that for any two points $x_1, x_2 \in X$, 
    \begin{equation}
      d_X (x_1, x_2) < \delta \implies d_Y \big( f(x_1), f(x_2)\big) < \epsilon
    \end{equation}
  \end{theorem}

  \begin{theorem}[Compact Metric Spaces are Complete]
    A metric space is compact iff it is complete and totally bounded. 
  \end{theorem}
  \begin{proof}
    
  \end{proof}

\subsection{Limit Point Compactness}

  We now state different, weaker types of compactness. 

  \begin{definition}[Countably Compact]
    A space $X$ is said to be \textbf{countably compact} if every countably open cover has a finite subcover. 
  \end{definition}

  \begin{definition}[Limit Point Compactness]
    A space $X$ is said to be \textbf{limit point compact} if every infinite subset of $X$ has a limit point. 
  \end{definition}

  \begin{theorem}
    Compactness $\implies$ limit point compactness.  
  \end{theorem}

  \begin{lemma}[Lebesgue Number Lemma]
    Let $\C$ be an open covering of the metric space $(X, d)$. If $X$ is compact, then there is a $\delta > 0$ such that for each subset of $X$ having diameter than $\delta$, there exists an element of $\C$ containing it. This number $\delta$ is called a \textbf{Lebesgue number} for the covering $\C$. 
  \end{lemma}

  Another theorem of calculus, suitably generalized to topological spaces, is stated. 

\subsection{Sequential Compactness}

  \begin{definition}[Sequentially Compact]
    A space $X$ is said to be \textbf{sequentially compact} if every sequence of points in $X$ has a subsequence that converges to a point $x \in X$. 
  \end{definition}

  \begin{theorem}[Equivalence of Compactness in Metrizable Spaces]
    Let $(X, \T)$ be a metrizable space. Then the following are equivalent: 
    \begin{enumerate}
      \item $X$ is compact. 
      \item $X$ is limit point compact. 
      \item $X$ is sequentially compact. 
      \item $X$ is countably compact. 
    \end{enumerate}
  \end{theorem}

\subsection{Local Compactness}

  \begin{definition}[Locally Compact]
    A space $X$ is said to be \textbf{locally compact} at $x$ if there is some compact subset $C$ of $X$ that contains a neighborhood of $x$. If $X$ is locally compact at each of its points, $X$ is simply to be \textbf{locally compact}. 
  \end{definition}

  \begin{example}
    The real line $\mathbb{R}$ is locally compact since any point $x \in \mathbb{R}$ lies within a certain closed interval $[a,b]$, which is compact. The subspace $\mathbb{Q}$ is not locally compact. 
  \end{example}

  Two of the most well-behaved classes of spcaes to deal with are metrizable spaces and compact Hausdorff spaces. If a given space is not one of these types, the next best thing one can hope for is that it is a subspace of one of these spaces. Clearly, a subspace of a metrizable space is itself metrizable, so one does not get any new spaces this way. However, a subspace of a compact Hausdorff space need not be compact. This leads to the question: Under what conditions is a space homeomorphic to a subspace of a compact Hausdorff space? 

  \begin{definition}[Compactification]
    Let $X$ be a locally compact Hausdorff space. Take some object outside $X$, denoted by the symbol $\infty$, and adjoin it to $X$, forming the set
    \begin{equation}
      Y = X \cup \{\infty\}
    \end{equation}
    Topologize $Y$ by defining the collection of open sets in $Y$ to be the sets of the following types:
    \begin{enumerate}
      \item $U$, where $U$ is an open subset of $X$. 
      \item $Y \setminus C$, where $C$ is a compact subset of $X$.
    \end{enumerate}
    Then, this space $Y$ is called the \textbf{one-point compactification of $X$}. This is in some sense the minimal compactification of $X$. 
  \end{definition}

  We briefly show that this set of open sets on $Y$ is indeed a topology. First, $\emptyset$ is of type 1 and $Y$ itself is of type 2. Given $U_i$ of type 1 and $Y \setminus C_i$ of type 2, we have the intersections of two sets
  \begin{align*}
    &U_1 \cap U_2 & \text{ is type 1} \\
    &(Y \setminus C_1) \cap (Y \setminus C_2) = Y \setminus (C_1 \cup C_2) & \text{ is type 2} \\
    &U_1 \cap (Y \setminus C_1) = U_1 \cap (X \setminus C_1) & \text{ is type 1} 
  \end{align*}
  along with the arbitrary union of sets
  \begin{align*}
    &\bigcup U_\alpha = U & \text{ is type 1} \\
    &\bigcup (Y \setminus C_\beta) = Y \setminus (\bigcap C_\beta) = Y \setminus C & \text{ is type 2} \\
    &(\bigcup U_\alpha) \cup ( \bigcup (Y \setminus C_\beta)) = U \cup (Y \setminus C) = Y \setminus (C \setminus U) & \text{ is type 2}
  \end{align*}
  We now present some properties of one-point compactifications. 

  \begin{theorem}
    Let $X$ be a locally compact Hausdorff space which is not compact, and let $Y$ be a one-point compactification of $X$. Then $Y$ is a compact Hausdorff space. Additionally, since $X \subset Y$ with $Y \setminus X$ consisting of a single point, $\bar{X} = Y$. 
  \end{theorem}

  \begin{example}[Extended Real Number Line]
    The one-point compactification of the real line $\mathbb{R}$ is homeomorphic to the circle $S^1$. That is, 
    \begin{equation}
      \mathbb{R} \cup \{\infty\} \cong S^1
    \end{equation}
    $\mathbb{R} \cup \{\infty\}$ is called the \textbf{extended real number line}. 
    
    \begin{figure}[H]
      \centering 
      \begin{tikzpicture}[scale=0.7]
        \draw (0,0) circle (3);
        \node[above] at (0,3) {$s$};
        \draw[<->] (-7,0)--(7,0);
        \node[above] at (6.7,0) {$\mathbb{R}$};
        \draw[->, dashed] (0,3)--(3.2,-0.2);
        \draw[->, dashed] (0,3)--(2.7,-2.5);
        \draw[->, dashed] (0,3)--(6.8,-1);
        \draw[fill=white] (0,3) circle (0.1);
        \draw[->, dashed] (0,3)--(-2.8, -2.8);
        \draw[fill] (3,0) circle (0.05);
        \draw[fill] (-2.349,-1.866) circle (0.05);
        \draw[fill] (5.1,0) circle (0.05);
        \draw[fill] (2.622,1.458) circle (0.05);
        \draw[fill] (-1.448,0) circle (0.05);
        \draw[fill] (1.471,0) circle (0.05); 
        \draw[fill] (2.371, -1.838) circle (0.05);
        \node[left] at (2.371, -1.838) {$a$};
        \node[below left] at (1.471,0) {$p(a)$};
        \node[above left] at (-1.448,0) {$p(b)$};
        \node[right] at (-2.349,-1.866) {$b$};
        \node[above right] at (5.1,0) {$p(d)$};
        \node[above right] at (2.622,1.458) {$d$};
        \node[below right] at (3,0) {$c = p(c)$};
      \end{tikzpicture}
      \caption{We can visualize this homeomorphism by visualizing the stereographic projection $p: S^1 \setminus \{s\} \rightarrow \mathbb{R}$. } 
      \label{fig:extended_r}
    \end{figure}
  \end{example}

  \begin{example}[2-Sphere]
    The one point-compactification of the real plane $\mathbb{R}^2$ is homeomorphic to the 2-sphere $S^2$. That is, 
    \begin{equation}
      \mathbb{R}^2 \cup \{\infty\} \cong S^2
    \end{equation}
  \end{example}

  \begin{lemma}
    Let $X$ be a Hausdorff space. Then $X$ is locally compact at $x$ if and only if for every neighborhood $U$ of $x$, there is a neighborhood $V$ of $x$ such that $\bar{V}$ is compact and $\bar{V} \subset U$. 
  \end{lemma}

  \begin{corollary}
    Let $X$ be a locally compact Hausdorff space with $Y$ a subspace of $X$. If $Y$ is closed in $X$ or open in $X$, then $Y$ is locally compact. 
  \end{corollary}

  \begin{corollary}
    A space $X$ is homeomorphic to an open subset of a compact Hausdorff space if and only if $X$ is locally compact and Hausdorff. 
  \end{corollary}

\subsection{Countable Compactness}

\subsection{Tychonoff Theorem}

  \begin{theorem}[Tychonoff Theorem]
    An arbitrary product of compact spaces is compact under the product topology. 
  \end{theorem}

  \begin{definition}
    A space $X$ is \textbf{completely regular} if one-point sets are closed in $X$ and if for each point $x_0$ and each closed set $A$ not containing $x_0$, there is a continuous function $f: X \longrightarrow [0,1]$ such that $f(x_0) = 1$ and $f(A) = \{0\}$. 
  \end{definition}

  \begin{theorem}
    A subspace of a completely regular space is completely regular. A product of completely regular spaces is completely regular. 
  \end{theorem}

  \begin{theorem}
    If $X$ is completely regular, then $X$ can be imbedded in $[0,1]^J$ for some $J$. 
  \end{theorem}

  \begin{corollary}
    Let $X$ be a space. The following are equivalent: 
    \begin{enumerate}
      \item $X$ is completely regular. 
      \item $X$ is homeomorphic to a subspace of a compact Hausdorff space. 
      \item $X$ is homeomorphic to a subspace of a normal space. 
    \end{enumerate}
  \end{corollary}

  \begin{definition}
    A \textbf{compactification} of a space $X$ is a compact Hausdorff space $Y$ containing $X$ such that $X$ is dense in $Y$ (that is $\bar{X} = Y$). Two compactifications $Y_1$ and $Y_2$ of $X$ are said to be \textbf{equivalent} if there is a homeomorphism $h: Y_1 \longrightarrow Y_2$ such that $h(x) = x$ for every $x \in X$. 
  \end{definition}

  \begin{theorem}
    Let $X$ be completely regular, and let $\beta(X)$ be its Stone-Cech compatification. Then every bounded continuous real-valued function on $X$ can be uniquely extended to a continuous real-valued function on $\beta(X)$. 
  \end{theorem}

  \begin{lemma}
    Let $A \subset X$, and let $f: A \longrightarrow Z$ be a continuous map of $A$ into the Hausdorff space $Z$. There is at most one extension of $f$ to a continuous function $g: \bar{A} \longrightarrow Z$. 
  \end{lemma}

  \begin{theorem}
    Let $X$ be completely regular. Let $Y_1, Y_2$ be two compactifications of $X$ having the extension property. Then there is a homeomorphism $\phi$ of $Y_1$ onto $Y_2$ such that $\phi(x) = x$ for each $x \in X$. 
  \end{theorem}

