\section{Open and Closed Sets} 

  The first thing to define is a topology. 

  \begin{definition}[Topology]
    Let $X$ be a set and $\T$ be a family of subsets of $X$. Then $\T$ is a \textbf{topology} on $X$\footnote{I will use script letters to denote topologies and capital letters to denote sets.} if it satisfies the following properties. 
    \begin{enumerate}
      \item \textit{Contains Empty and Whole Set}: 
      \begin{equation}
        \emptyset, X \in \T
      \end{equation}

      \item \textit{Closure Under Union}. If $\{U_\alpha\}_{\alpha \in A}$ is a class of sets in $\T$, then 
      \begin{equation}
        \bigcup_{\alpha \in A} U_\alpha \in \T
      \end{equation}

      \item \textit{Closure Under Finite Intersection}: If $U_1, \ldots, U_n$ is a finite class\footnote{Note that we restrict property 3 to be a \textit{finite} intersection because it turns out that the finiteness of intersection allows us to prove many nice properties about topologies, which we will mention later. Another reason is that if we remove this finite restriction, the open ball topology on $\mathbb{R}$ would imply that $\cap_{i = 1}^{\infty} ( - 1/i, +1/i ) = 0$ is an open set $\implies$ all points are open sets too, which is generally not what we want in analysis. 
      } of sets in $\T$, then 
      \begin{equation}
       \bigcap_{i=1}^{n} U_i \in \T
      \end{equation}
    \end{enumerate}
    A \textbf{topological space} is denoted $(X, \T)$. 
  \end{definition}

  For the sake of giving at least one nontrivial example, here is an example of a finite topology. 

  \begin{example}[Topologies of a Set of Cardinality 3]
    There are a total of 29 topologies that we can construct on $\{1, 2, 3\}$. Two such examples are 
    \begin{enumerate}
      \item $\{\emptyset, \{1, 2\}, \{1, 2, 3\}\}$ 
      \item $\{\emptyset, \{3\}, \{2, 3\}, \{1, 2, 3\}\}$
    \end{enumerate}
  \end{example} 

  When we define a new topology, we must first prove that they are topologies, and so these definitions are really theorems. However, I will introduce them as definitions and reserve the theorem environment for actual theorems. 

  \begin{definition}[Discrete, Indiscrete Topologies]
    Given a set $X$, 
    \begin{enumerate}
      \item $2^X$ is a topology, called the \textbf{discrete topology}. 
      \item $\{\emptyset, X \}$ is a topology, called the \textbf{indiscrete topology}. 
    \end{enumerate}
  \end{definition}
  \begin{proof}
    Listed. 
    \begin{enumerate}
      \item The first property is trivially proven. From the theorems of set theory, $U_\alpha \subset X \implies \cup U_\alpha \subset X \implies \cup U_\alpha \in 2^X$. Finally the same logic holds for intersection as well. 
      \item The first property is trivially proven. We can check for the 4 combinations of unions and intersections and see that they all result in either $\emptyset$ or $X$. 
    \end{enumerate}
  \end{proof}

  \begin{definition}[Finer, Coarser Topologies]
    Suppose that $\T$ and $\T^\prime$ are two topologies on a given set $X$. If $\T \subset \T^\prime$, we say that $\T^\prime$ is \textbf{finer} than $\T$, or equivalently, we say that $\T$ is \textbf{coarser} than $\T^\prime$. 
  \end{definition}

  We can think of the topology of a set $X$ as a truck full of gravel as the open sets. If the gravel is smashed into smaller, finer pieces, then the amount of stuff that we can make from the finer gravel increases, which corresponds to a bigger topology. Clearly, the indiscrete topology is the coarsest topology and the discrete topology is the finest. 

  \begin{theorem}[Intersection of Topologies]
    Given a family of topologies $\{\T\}_{\alpha \in A}$, the set 
    \begin{equation}
      \T = \bigcap_{\alpha \in A} \T_\alpha
    \end{equation}
    is a topology. 
  \end{theorem}

  \begin{corollary}[Unique Coarsest and Finest Topology]
    Given a family of topologies $\{\T\}_{\alpha \in A}$, there exists 
    \begin{enumerate}
      \item a unique smallest topology on $X$ containing all the collections $\T_\alpha$. 
      \item a unique largest topology on $X$ contained in each $\T_\alpha$. 
    \end{enumerate}
  \end{corollary} 

  \begin{example}
    Let $X = \{a, b, c\}$, and let 
    \begin{align}
      \T_1 & = \{\emptyset, X, \{a\}, \{a, b\}\} \\
      \T_2 & = \{\emptyset, X, \{a\}, \{b, c\}\}
    \end{align}
    We claim that the 
    \begin{enumerate}
      \item smallest topology containing $\T_1, \T_2$ is 
      \begin{equation}
        \T_{1 \cup 2} = \{\emptyset, X \{a\}, \{b\}, \{a, b\}, \{b, c\}\}
      \end{equation} 
      Note that this is not simply the union of topologies. The union wouldn't have $\{b\}$, making it not a topology. 

      \item largest topology contained in $\T_1, \T_2$ is 
      \begin{equation}
        \T_{1 \cap 2} = \{\emptyset, X, \{a\}\}
      \end{equation}
      Note that this is simply the intersection of the two topologies. 
    \end{enumerate}
  \end{example}

\subsection{Open Sets}

  This leads to the most general definition of an open set. Note that an open set doesn't really mean anything without talking about with respect to its topology. 

  \begin{definition}[Open Set]
    The elements of $\T$ are called \textbf{open sets} in $X$.\footnote{As implied from the definition of a topology, the arbitrary union and finite intersection of any number of open sets is an open set.} 
    \begin{enumerate}
      \item An open set $U$ which contains a point $x$ is called an \textbf{open neighborhood} of $x$, denoted $U_x$. 
      \item Given an open neighborhood $U_x$ of $x$, the set $U_x \setminus \{x\}$ is called the \textbf{punctured open neighborhood} of $x$. 
    \end{enumerate}
  \end{definition}

\subsection{Basis} 

  So far so good. We have introduced a topology as a collection satisfying some properties, but given a topology on $X$, it is often a bit difficult to see whether an arbitrary set $S \subset X$ is open. For this, we have the following. 

  \begin{lemma}[A Set Full of Open Sets is Open]
    Given $X$ with a topology $\T$, let $S \subset X$. Then $S$ is open if for every $x \in S$, there exists an open neighborhood $U_x$ satisfying $x \in U_x \subset S$. 
  \end{lemma}
  \begin{proof}
    Since we can set 
    \begin{equation}
      S = \bigcup_{x \in S} U_x
    \end{equation}
    it is an arbitrary union of open sets and therefore must be open. 
  \end{proof}

  We want to continue analyzing the properties of a topology, but sometimes working with the entire topology is a bit thorny. There is a tamer representation of a topology, which can also give us the starting point to \textit{construct} topologies. 

  \begin{definition}[Basis] 
    \label{def:basis}
    If $X$ is a set, a \textbf{basis} on $X$ is a collection $\B$ of subsets of $X$ (called \textbf{basis elements}) such that
    \begin{enumerate}
      \item For each $x \in X$, there is at least one basis element $B \in \B$ containing $x$. That is, the elements of $\B$ covers $X$. 
      \item If $x$ belongs to the intersection of two basis elements $B_1$ and $B_2$, then there is a basis element $B_3$ containing $x$ such that $B_3 \subset (B_1 \cap B_2)$. 
    \end{enumerate}
  \end{definition} 

  The name gives away the clue that a topology may be created from this basis.  

  \begin{definition}[Basis to Topology]
    \label{def:basis-to-topology}
    Given a basis $\B$ on a set $X$, we can define a topology $\T$, called the \textbf{topology generated by $\B$}, in the following equivalent ways. 
    \begin{enumerate}
      \item $\T$ consists of subsets $U$ of $X$ satisfying the property that for each $x \in U$, there exists a basis element $B \in \B$ such that $x \in B \subset U$.\footnote{Note that since we can always set $U = \emptyset$, the basis doesn't need to contain $\emptyset$. }
      \begin{center}
        \begin{tikzpicture}
        \draw[dashed] plot [smooth cycle] coordinates {(0,0) (1,1) (3,1) (3,0) (1,-1)};
        \node [right] at (0.5,0) {U};
        \draw[fill=lightgray,dashed] (2,0.7) circle [radius=0.35];
        \draw[fill] (2, 0.7) circle [radius=0.05];
        \node [right] at (1.95,0.7) {x};
        \node [right] at (4,-0.5) {B};
        \draw (1.95,0.6)--(4,-0.5);
        \end{tikzpicture}
      \end{center}

      \item $\T$ consists of all possible unions of elements in $\B$. 
      \begin{equation}
        \T \equiv \Big\{ \bigcup_i B_i \mid B_i \in \B\Big\}
      \end{equation}
    \end{enumerate}
  \end{definition} 
  \begin{proof}
    We prove that the 2 methods generate a topology, and then finally prove that it they are the same topology. 
    \begin{enumerate}
      \item Clearly, $\emptyset$ and $X$ itself are in $\T$. To prove property 2, given a certain indexed family of subsets $\{U_\alpha\}_{\alpha \in I}$ of $\T$, we must show that 
      \begin{equation}
        U = \bigcup_{\alpha \in I} U_\alpha \in \T
      \end{equation}
      Given $x \in U$, there exists at least one index $\alpha$ such that $x \in U_\alpha$. Since $U_\alpha \in \T$ already, there exists a basis element $b \in \B$ such that $x \in b \subset U_\alpha$. But 
      \begin{equation}
        U_\alpha \subseteq U \implies b \subset U
      \end{equation}
      So, by definition, any arbitrary union of $U$ of these subsets is also in $\T$. 

      To prove property 3, we must show that 
      \begin{equation}
        W = \bigcap_{\alpha \in I} U_\alpha \in \T
      \end{equation}
      Given $x \in W$, by definition of a basis element, there exists a $b \in \B$ such that 
      \begin{equation}
        x \in b \subset (U_\beta \cap U_\gamma) \forall \beta, \gamma \in I \implies \text{ there exists } \Tilde{b} \in \B \text{ s.t. } x \in \Tilde{b} \subset \bigcap_{\alpha \in I} U_\alpha
      \end{equation}
      By definition, $W$ is also open. Since this arbitrary set of subsets $\T$ suffices the 3 properties, it is a topology of $X$ by definition. 

      \item $(\rightarrow)$ Given a collection of elements in $\B$, they are also elements of $\T$. Since $\T$ is a topology, their union in also in $\T$. 

      $(\leftarrow)$ Given an open set $U \in \T$, for every point $x \in U$, by definition we can choose a basis element $b \in \B$ such that $x \in b \subset U$. Then, the union of all these basis elements is by definition $b$. 
        
    \end{enumerate}
  \end{proof}

  We have learned how to go from a basis to a topology. The following lemma tells us how to identify a basis within a topology. 

  \begin{theorem}[Topology to Basis]
    Let $(X, \T)$ be a topological space, and let $\B$ be a collection of open subsets of $X$ such that for every open set $U$ and each $x \in U$, there exists an element $B \in \B$ such that
    \begin{equation}
      x \in B \subset U
    \end{equation}
    Then, $\B$ is a basis for the topology of $X$. 
  \end{theorem}
  \begin{proof}
    Note that there are two claims here: $\B$ is a basis and the topology that $\B$ generates is equal to $\T$. 
    \begin{enumerate}
      \item To prove that $\B$ is a basis, note that $X$ is an open set, and by assumption, for every $x \in X$, there exists a $B \in \B$ s.t. $x \in B \subset X$. Therefore $\B$ covers $X$. Now take two basis elements $B_1, B_2 \in \B$ with $x \in B_1 \cap B_2$. Since we know that $B_1, B_2$ are open, $B_1 \cap B_2$ is open and so for each $x \in B_1 \cap B_2$, there exists a basis element $B_3$ s.t. $x \in B_3 \in (B_1 \cap B_2)$. Thus $\B$ is a basis. 

      \item Let us call $\T^\prime$ the topology generated by $\B$. Then, given $U \in \T$, by assumption for any $x \in U$, there exists a basis element $B \in \B$ s.t. $x \in B \subset U$, so $x \in \T^\prime$. Conversely, if $U \in \T^\prime$, then $U$ is an arbitrary union of elements $B \in \B$ where each $B$ is open in $\T$, so $U \in \T$. So $\T = \T^\prime$. 
    \end{enumerate}
  \end{proof} 

  Characterizing topologies in terms of basis is quite effective since we can work with more manageable sets. 

  \begin{lemma}[Fineness w.r.t. Basis]
    Given two topologies $\T$ and $\T^\prime$ with their bases $\B$ and $\B^\prime$, respectively, the following are equivalent. 
    \begin{enumerate}
      \item $\T^\prime$ is finer than $\T$. 
      \item For each $x \in X$ and basis element $B \in \B$ containing $x$, there exists a basis element $B^\prime \in \B^\prime$ such that $x \in B^\prime \subset B$. 
    \end{enumerate}
  \end{lemma}

  So we have seen how we can take a collection of sets satisfying the basis properties and construct a topology as the union of the sets in this collection. What happens if we can relax some of these conditions? Note that the first condition was that the basis elements must cover $X$. This is non-negotiable. However, if we remove the second requirement that a basis element must be contained in an intersection of basis elements, we can get a \textit{subbasis}. 

  \begin{definition}[Subbasis]
    A \textbf{subbasis} $\mathscr{S}$ for a topology on $X$ is a collection of subsets of $X$ whose union is equal to $X$. 
  \end{definition}

  \begin{theorem}[Subbasis to Topology]
    Given a subbasis $\mathscr{S}$ on a set $X$, the \textbf{topology generated by $\mathscr{S}$} is defined to be the collection $\T$ of all unions of finite intersections of elements of $\mathscr{S}$. 
  \end{theorem}
  \begin{proof}
    It suffices to show that the collection of finite intersections of elements form a basis. 
  \end{proof}
 
\subsection{Limit Points and Closed Sets} 

  First, we need to learn what it generally means for a point to be infinitesimally close to a set. If we take a point and draw smaller and smaller circles around it, the circle itself should still overlap with $S$, no matter how small it gets.

  \begin{definition}[Limit Point]
    Given a topological space $(X, \T)$, let $x \in X$ be a point and $S \subset X$ a subset. $x$ is a \textbf{limit point of $S$} if every punctured neighborhood of $x$ intersects $S$. The set of all limit points of a set $S$ is denoted $S^\prime$.  
  \end{definition} 

  Note that limit point are generally used to talk about points that are infinitesimally close to a set $S$. A limit point may not necessarily be in $S$, and a point of $S$ may not necessarily be a limit point. This is why we use a \textit{punctured neighborhood}, rather than an open neighborhood. For continuity as we will see later, we just talk about neighborhoods since we also claim that the limit exists and the function value is the limit.

  \begin{example}[Examples of Limit Points]
    What about the limit points that are not in $S$? Generally, there are two instances. 
    \begin{enumerate}
      \item Let $S$ represent the gray area. $B$ is in the ``interior'' of $S$ and therefore is a limit point. $A$ and $C$ are on the ``boundary'' of $S$ yet not in $S$, and we can show that they are limit points as well. 
        
      \begin{figure}[H]
        \centering 
        \begin{tikzpicture}
          \draw[fill=lightgray, dashed] plot [smooth cycle] coordinates {(0,0) (1,1) (3,1) (3,0) (1,-1)};
          \draw [fill] (3, 1) circle [radius=0.05];
          \node [above right] at (3,1) {A};
          \draw [fill] (2,0) circle [radius=0.05];
          \node [above] at (2,0) {B};
          \draw [fill] (0,0) circle [radius=0.05];
          \node [left] at (0,0) {C};
          \draw [fill] (5,0) circle [radius=0.05];
          \node [right] at (5, 0) {D};
        \end{tikzpicture}
        \caption{Points $A, B, C$ are limit points of the open set. }
        \label{fig:limit_boundary}
      \end{figure}

      \item A point can be at the ``convergence point'' of a sequence. 
      \begin{figure}[H]
        \centering 
        \begin{tikzpicture}
          \draw [fill] (0, 2.2) circle [radius=0.05];
          \draw [fill] (1, 2.4) circle [radius=0.05];
          \draw [fill] (2, 2) circle [radius=0.05];
          \draw [fill] (2.5, 1.8) circle [radius=0.05];
          \draw [fill] (2.6, 1.6) circle [radius=0.05];
          \draw [fill] (2.65, 1.67) circle [radius=0.05];
          \draw [fill] (2.654, 1.64) circle [radius=0.05];
          \draw [fill] (2.6543, 1.63) circle [radius=0.05];
          \node [right] at (2.6543, 1.63) {p};
        \end{tikzpicture}
        \caption{Note that if $S$ is a sequence of points in $\mathbb{R}^{2}$ that converges to $p$ without ever hitting it, we can say that $p \not\in S$ is a limit point of $S$.}
        \label{fig:limit_sequence}
      \end{figure}
    \end{enumerate}
  \end{example}

  \begin{example}[Examples of Non-Limit Points]
    There are generally two instances of non-limit points. Let $X = \mathbb{R}$ and $S = (0, 1) \cup \{2\}$. 
    \begin{enumerate}
      \item $5$ is clearly not a limit point. 
      \item $2$, although in $S$, is not a limit point since we are talking about the punctured neighborhood. A point in $S$ that is not a limit point is called an \textbf{isolated point}. 
    \end{enumerate}
  \end{example}

  \begin{example}[Counterintuitive Limit Points in Lower Limit Topology]
    Note that given an interval $(a, b) \subset \mathbb{R}$ in the lower limit topology, $a$ is a limit point but $b$ is \textit{not} a limit point!
  \end{example}

  \begin{theorem}[Union of Limit Points is Limit Points of Union]
    Let $A_1, \ldots, A_n$ be a finite collection of sets. Then 
    \begin{equation}
      \bigcup_{i=1}^n A_i^\prime = \bigg( \bigcup_{i=1}^n A_i \bigg)^\prime
    \end{equation}
  \end{theorem}
  \begin{proof}
    Let the LHS be $W$ and the RHS be $V$. If $x \in W$, $x \in A_i^\prime$ for some $i$, and so for all $\epsilon > 0$, there exists a $B_\epsilon^\circ (x)$ s.t. 
    \begin{equation}
      B_\epsilon^\circ (x) \cap A_i \neq \emptyset \implies B_\epsilon^\circ (x) \cap \bigg( \bigcup_{i=1}^n A_i \bigg) \neq \emptyset
    \end{equation}
    which means that $x \in V$. Now assume that $x \in V$. Then for all $\epsilon > 0$, there exists a $B_\epsilon^\circ (x)$ s.t. 
    \begin{equation}
      B_\epsilon^\circ (x) \cap \bigg( \bigcup_{i=1}^n A_i \bigg) \neq \emptyset
    \end{equation}
    which implies that $B_\epsilon^\circ (x) \cap A_i \neq \emptyset$ for some $i$, which means that $x \in A_i^\prime \subset W$. 
  \end{proof}

  Note that this is clearly not true for infinite unions. Look at the countable set $\mathbb{Q} \subset \mathbb{R}$. Each $\{q\}^\prime = \emptyset$, but $\mathbb{Q}^\prime = \mathbb{R}$. Look at the uncountable set $\mathbb{R}$. Each $\{x \in \mathbb{R}\}^\prime = \emptyset$, but $\mathbb{R}^\prime = \mathbb{R}$. 

  \begin{definition}[Closed Set]
    A set $S \subset X$ is \textbf{closed}\footnote{Note that open and closed sets are not mutually exclusive. A set might be open, closed, both, or neither. A set that is both open and closed is called \textbf{clopen}.} if it satisfies either of the equivalent properties. 
    \begin{enumerate}
      \item Its complement $X \setminus S$ is open in $\T$. 
      \item It contains all of its limit points. 
    \end{enumerate}
  \end{definition}
  \begin{proof}
    We prove bidirectionally. 
    \begin{enumerate}
      \item ($\rightarrow$) Given that $S$ contains all its limit points, then let $x \in S^c$. $x$ is not a limit point of $S$ since if it were, then it would be in $S$, and so there exists a punctured open neighborhood $B_\epsilon^\circ (x)$ of $x$ s.t. $S \cap B_\epsilon^\circ (x) = \emptyset$. Since $x \not\in S$, we also have $S \cap B_\epsilon (x) = \emptyset$, which implies that $B_\epsilon (x) \subset S^c$. Since for every $x \in S^c$, there exists a $B_\epsilon (x) \subset S^c$, $S^c$ is open. 

      \item ($\leftarrow$) For simplicity, it suffices to prove if $S$ open, then $S^c$ is closed. Given that $S$ is open, we have for every $x \in S$, there exists $B_\epsilon (x) \subset S$, which implies that $B_\epsilon (x) \cap S^c = \emptyset$. Since there exists an $B_\epsilon (x)$ that does not contain points in $S^c$, $x$ cannot be a limit point of $S^c$, and so there exists no limit points of $S^c$ in $S$. Therefore, all limit points of $S^c$ are in $S^c$, proving that $S^c$ is closed.  
    \end{enumerate}
  \end{proof}
  
  \begin{theorem}[Topological Space wrt Closed Sets]
    Let $X$ be a topological space. Then, the following conditions hold
    \begin{enumerate}
      \item $\emptyset$ and $X$ are clopen.
      \item Arbitrary intersections of closed sets are closed. 
      \item Finite unions of closed sets are closed. 
    \end{enumerate}
  \end{theorem}
  \begin{proof}
    Listed. 
    \begin{enumerate}
      \item Let $x$ be a limit point of $\cap_\alpha F_\alpha$, and we want to show that $x \in \cap_\alpha F_\alpha$. By definition of limit points, for every $\epsilon > 0$, we have 
      \begin{equation}
        B_\epsilon (x) \cap \bigg( \bigcap_\alpha F_\alpha \bigg)
      \end{equation}
      which means that $B_\epsilon (x) \cap F_\alpha \neq \emptyset$ for all $\alpha$. This means that $x$ is a limit point for every $F_\alpha$, and since they are all closed, $x \in F_\alpha$ for all $\alpha$, which implies that $x \in \cap_\alpha F_\alpha$. 
    \end{enumerate}
  \end{proof}

\subsection{Dense Subsets}

  \begin{definition}[Dense Subsets]
    Let $S \subset (X, \mathscr{T})$. $S$ is \textbf{dense} in $X$ if every point $p \in X$ is a limit point of $S$. In other words, for any point $p \in X$ and any open neighborhood $U_p$ of $p$, $U_p \cap S$ is nontrivial. Otherwise, $p$ is a point of $S$. 
  \end{definition}

  The following example is a crucial fact for proving further properties of topological spaces. 

  \begin{example}
    $\mathbb{Q}^{n}$ is a dense set of $\mathbb{R}^{n}$ with the open ball topology. If we have the discrete topology of $\mathbb{R}^{2}$, an open neighborhood of a point is the point itself, so no limit points would exist beyond the points in $S$ itself. So $\mathbb{Q}^{n}$ is not dense in $\mathbb{R}^{n}$ with this topology. 
  \end{example}

\subsection{Interiors and Closures}

  Now that we've determined limit points, we would like to extend sets into their limit points. The process of doing this is called the \textit{closure} of a set. 

  \begin{definition}[Closure]
    The \textbf{closure} of set $S$ is $\overline{S}$ is defined in the following equivalent ways. 
    \begin{enumerate}
      \item $\overline{S} = S \cup S^\prime$, i.e. the union of itself and its limit points. 
      \item $\overline{S}$ is the intersection of all closed sets $C$ containing $S$. 
    \end{enumerate}
  \end{definition}
  \begin{proof}
    
  \end{proof}

  \begin{theorem}
    If $X$ is a metric space and $E \subset X$, then 
    \begin{enumerate}
      \item $\overline{E}$ is closed. 
      \item $E = \overline{E}$ if and only if $E$ is closed. 
      \item $\overline{E} \subset F$ for every closed set $F \subset X$ such that $E \subset F$. That is, if $E \subset F$ closed, then ``increasing" the size of $E$ to its closure will not make it greater than $F$. 
    \end{enumerate}
  \end{theorem}
  \begin{proof}
    Listed. 
    \begin{enumerate}
      \item Let $x$ be a limit point of $\overline{E}$. Then, for every $\epsilon > 0$, we have $B_\epsilon (x) \cap \overline{E} \neq \emptyset$, which means that either $B_\epsilon (x) \cap E \neq \emptyset$ (in which case $x \in E^\prime \implies x \in \overline{E}$ and we are done) or $B_\epsilon (x) \cap E^\prime \neq \emptyset$. We wish to prove that in the latter case, $x$ being a limit point of $E^\prime$ still implies that $x$ is a limit point of $E$. Since $B_\epsilon (x) \cap E^\prime \neq \emptyset$, there must exist a $y \in B_\epsilon (x) \cap E^\prime$. Since $y \in E^\prime$, we can construct an open ball $B_\delta (y)$ containing elements of $E$, and since $B_\epsilon (x)$ is open, we can contain $B_\delta (y)$ entirely within $B_\epsilon (x)$. Therefore, 
      \[B_\delta (y) \cap E \neq \emptyset \implies B_\epsilon (x) \cap E \neq \emptyset\]
      therefore, $x \in E^\prime \implies x \in \overline{E}$. 

      \item If $E$ is closed, then $E^\prime \subset E \implies \overline{E} = E \cup E^\prime = E$. If $E = \overline{E} = E \cup E^\prime$, then $E^\prime \subset E \implies E$ is closed. 

      \item Since $E \subset F$, it suffices to prove that $E^\prime \subset F$. Consider a limit point $x$ of $E$. Then every punctured open neighborhood of $x$ satisfies $B_\epsilon^\circ (x) \cap E \neq \emptyset$. But since $E \subset F$, we have 
      \[B_\epsilon^\circ (x) \cap F \neq \emptyset\]
      and so $x$ is also a limit point of $F$. But since $F$ is closed, $x \in F$. Therefore, $\overline{E} = E \cup E^\prime \subset F$. 
    \end{enumerate}
  \end{proof}

  The first two statements (1) and (2) imply the following. 

  \begin{corollary}
    The closure of the closure of $E$ is equal to the closure of $E$. 
  \end{corollary}
  \begin{proof}
    We know that $\overline{\overline{E}} \supset \overline{E}$, so we must prove that $\overline{\overline{E}} \subset \overline{E}$, which is equivalent to proving that $\overline{E}^\prime \subset \overline{E}$. Let $x \in \overline{E}^\prime$, i.e. is a limit point of $\overline{E}$. Then, for every $\epsilon > 0$, we have $B_\epsilon (x) \cap \overline{E} \neq \emptyset$. Pick a point $y$ from this intersection, and since $B_\epsilon (x)$ is open, we can construct an open ball $B_\delta (y)$ fully contained in $B_\epsilon (x)$. Since $y \in \overline{E}$, $y$ is a limit point of $E$, which implies 
    \begin{equation}
      B_\delta (y) \cap E \neq \emptyset \implies B_\epsilon (x) \cap E \neq \emptyset
    \end{equation}
    and therefore $x$ is a limit point of $E$, $x \in \overline{E}$. 
  \end{proof}
  

  \begin{example}
    If $S$ is an open ball, $\Bar{S}$ is the closed ball. 
  \end{example}

  From semantics, it may seem like the interior and exterior (defined later) are related, but from a mathematical point of view, the interior and closure are dual notions. 

  \begin{definition}[Interior]
    Let $S \subset X$. Then, the following definitions of the \textbf{interior} of $S$, denoted $S^\circ$, are equivalent. 
    \begin{enumerate}
      \item $x \in S^\circ$ if $\exists U_x \ni x$ s.t. $U_x \subset S$. 
      \item $S^\circ$ is the union of all open sets contained in $S$. 
      \item $S^{o}$ is the complement of the closure of the complement of S. 
      \begin{equation}
        S^{o} = \big(\overline{S^{c}}\big)^{c}
      \end{equation}
    \end{enumerate}
  \end{definition}
  \begin{proof}
    
  \end{proof}

  An interior point means that we can always contain the point in $S$ with some ``breathing room." By definition an open set is a set where all of its points are interior points. A set is then said to be open if every point has this breathing room. This can be useful when defining differentiation at a point within an open set, since we can always find a neighborhood to take limits on. 

  \begin{lemma}[Open and Closed in Terms of Interiors and Closures]
    Let $S$ be a subset of some topological space $X$. 
    \begin{enumerate}
      \item $S$ is open iff $S = S^{o}$. $S^{o}$ is always open.
      \item $S$ is closed iff $S = \overline{S}$. $\overline{S}$ is always closed. 
    \end{enumerate}
  \end{lemma}

  \begin{theorem}[Clopen sets in Reals]
    There are no proper clopen sets in $\mathbb{R}$. 
  \end{theorem}

\subsection{Exteriors and Boundaries}

  \begin{definition}[Exteriors]
    Let $S \subset X$. The \textbf{exterior} of $S$, denoted $S^e$, is defined in the following equivalent ways.\footnote{We can informally think of the exterior being strictly outside of $S$ and its boundary.}
    \begin{enumerate}
      \item $S^e$ is the complement of the closure of $S$. 
      \item $S^e$ is the interior of the complement of $S$. 
    \end{enumerate}
  \end{definition}
  \begin{proof}
    
  \end{proof}

  \begin{definition}[Boundary]
    Let $S \subset X$. The \textbf{boundary} of $S$, denoted $\partial S$, is defined in the following equivalent ways. 
    \begin{enumerate}
      \item $\partial S$ is the closure minus the interior of $S$ in $X$. 
      \item $\partial S$ is the intersection of the closure of $S$ with the closure of its complement, i.e the set of all points $x$ such that every neighborhood $U_x$ intersects both the interior and exterior. 
      \item $\partial S$ is the set of points that are neither in the exterior nor the interior. 
      \item $x \in \partial S$ if every neighborhood of $x$ intersects both the interior and exterior of $S$. 
    \end{enumerate}
  \end{definition}
  \begin{proof}
    
  \end{proof}

  From the above, we get the intuitive notion that these three parts divide up the whole space. 

  \begin{theorem}[Partitioning of Space]
    Given $S \subset X$, $X$ is partitioned into the interior, boundary, and exterior of $S$. 
    \begin{equation}
      X = S^\circ \sqcup \partial S \sqcup S^e
    \end{equation}
  \end{theorem}
  \begin{proof}
    The fact that 
  \end{proof}

  One counterintuitive result is the \href{https://en.wikipedia.org/wiki/Lakes_of_Wada}{Lakes of Wada}, which are three disjoint connected open sets of the open unit square $(0, 1)^2$ with the property that they \textit{all} have the same boundary. In other words, for any point selected on the boundary of one of the lakes, the other two lakes' boundaries also contain that point. 

