\section{Fields}  

  Our final structure is field, which seems to add only a few more conditions to a ring, but again unlocks more structure. Field theory is usually pretty tame compared to groups and rings. We will present some general results on fields, and then move quickly to construct the field of fractions, where we can formally define the rational and the real numbers. 

  \begin{definition}[Field]
    A \textbf{field} $(F, +, \times)$ is a commutative, associative ring where every nonzero element is a unit. 
  \end{definition}

  \begin{lemma}[Properties of Addition]
    The properties of addition hold in a field. 
    \begin{enumerate}
      \item If $x + y = x + z$, then $y = z$. 
      \item If $x + y = x$, then $y = 0$. 
      \item If $x + y = 0$, then $y = -x$. 
      \item $(-(-x)) = x$. 
    \end{enumerate}
  \end{lemma}
  \begin{proof}
    For the first, we have 
    \begin{align}
      x + y = x + z & \implies -x + (x + y) = -x + (x + z) && \tag{addition is a function} \\
                    & \implies (-x + x) + y = (-x + x) + z && \tag{$+$ is associative} \\
                    & \implies 0 + y = 0 + z && \tag{definition of additive inverse} \\
                    & \implies y = z && \tag{definition of identity}
    \end{align} 
    For the second, we can set $z = 0$ and apply the first property. For the third, we have 
    \begin{align}
      x + y = 0 & \implies -x + (x + y) = -x + 0 && \tag{addition is a function} \\
                & \implies (-x + x) + y = -x + 0 && \tag{$+$ is associative} \\
                & \implies 0 + y = -x + 0 && \tag{definition of additive inverse} \\
                & \implies y = -x && \tag{definition of identity}
    \end{align}
    For the fourth, we simply follow that if $y$ is an inverse of $z$, then $z$ is an inverse of $y$. Therefore, $-x$ being an inverse of $x$ implies that $x$ is an inverse of $-x$. $-(-x)$ must also be an inverse of $-x$. Since inverses are unique\footnote{This is proved in algebra.}, $x = -(-x)$. 
  \end{proof}

  \begin{lemma}[Properties of Multiplication]
    The properties of multiplication hold in a field. 
    \begin{enumerate}
      \item If $x \neq 0$ and $xy = xz$, then $y = z$. 
      \item If $x \neq 0$ and $xy = x$, then $y = 1$. 
      \item If $x \neq 0$ and $xy = 1$, then $y = x^{-1}$. 
      \item If $x \neq 0$, then $(x^{-1})^{-1} = x$. 
    \end{enumerate}
  \end{lemma}
  \begin{proof}
    The proof is almost identical to the first. Since $x \neq 0$, we can always assume that $x^{-1}$ exists. For the first, we have
    \begin{align}
      x y = x z & \implies x^{-1} (x y) = x^{-1} (x z) && \tag{multiplication is a function} \\
                & \implies (x^{-1} x) y = (x^{-1} x) z && \tag{$\times$ is associative} \\
                & \implies 1 y = 1 z && \tag{definition of multiplicative inverse} \\  
                & \implies y = z && \tag{definition of identity}
    \end{align}
    For the second, we can set $z = 1$ and apply the first property. For the third, we have 
    \begin{align}
      xy = 1 & \implies x^{-1} (x y) = x^{-1} 1 && \tag{multiplication is a function} \\
             & \implies (x^{-1} x) y = x^{-1} 1 && \tag{$\times$ is associative} \\
             & \implies 1 y = x^{-1} 1 && \tag{definition of multiplicative inverse} \\
             & \implies y = x^{-1} && \tag{definition of identity}
    \end{align}
    For the fourth, we simply see that $x^{-1}$ is a multiplicative inverse of both $x$ and $(x^{-1})^{-1}$ in the group $(\mathbb{F} \setminus \{0\}, \times)$, and since inverses are unique, they must be equal. 
  \end{proof}

  \begin{lemma}[Properties of Distribution]
    For any $x, y, z \in \mathbb{F}$, the field axioms satisfy 
    \begin{enumerate}
      \item $0 \cdot x = 0$.
      \item If $x \neq 0$ and $y \neq 0$, then $x y \neq 0$.
      \item $-1 \cdot x = -x$. 
      \item $(-x) y = - (xy) = x (-y)$. 
      \item $(-x) (-y) = xy$. 
    \end{enumerate}
  \end{lemma} 
  \begin{proof}
    For the first, note that 
    \begin{align}
      0 x & = (0 + 0) \cdot x = 0 x + 0x 
    \end{align}
    and subtracting $0x$ from both sides gives $0 = 0x$. For the second, we can claim that $xy \neq 0$ equivalently claiming that it will have an identity. Since $x, y \neq 0$, their inverses exists, and we claim that $(xy)^{-1} = y^{-1} x^{-1}$ is an inverse. We can see that by associativity, 
    \begin{equation}
      (y^{-1} x^{-1}) (xy) = y^{-1} (x^{-1} x) y = y^{-1} y = 1
    \end{equation} 
    For the third, we see that 
    \begin{equation}
      0 = 0 \cdot x = (1 + (-1)) \cdot x = 1 \cdot x + (-1) \cdot x = x + (-1) \cdot x 
    \end{equation}
    which implies that $-1 \cdot x$ is the additive inverse. The fourth follows immediately from the third by the associative property. For the fifth we can see that 
    \begin{align}
      (-x) (-y) & = (-1) x (-1) y && \tag{property 3} \\
                & = (-1) (-1) x y && \tag{$\times$ is commutative} \\
                & = -1 \cdot (-xy) && \tag{property 3} \\
                & = -(-xy) && \tag{property 3} \\
                & = xy && \tag{addition property 4}
    \end{align}
  \end{proof}

  \begin{theorem}[Fields are Euclidean Domains]
    Every field is a Euclidean domain. 
  \end{theorem}
  \begin{proof}
    Given $x, y \in \mathbb{F}$, assume $x y = 0$ with $x \neq 0$. Since $x$ is invertible,
    \begin{equation}
      0 = x^{-1} 0 = x^{-1} (x y) = y
    \end{equation}
    Now assuming that $y \neq 0$, since $y$ is invertible, 
    \begin{equation}
      0 = 0 y^{-1} = (x y) y^{-1} = x
    \end{equation}
  \end{proof}

  With this theorem, we have established the hierarchy in the beginning of this section. So as soon as we see a field, we can immediately apply everything we know, such as Euclidean division, unique factorization, GCDs, etc. The converse is not generally true, but with extra assumptions it is. 

  \begin{theorem}[Integral Domains are Embedded in Fields]
    An integral domain is a ring that is isomorphic to a subring of a field. 
  \end{theorem}

  \begin{theorem}[Ideals of Fields]
    The only ideals that exist in a field $\mathbb{F}$ is $\{0\}$ and $\mathbb{F}$ itself. 
  \end{theorem}
  \begin{proof}
    Given a nonzero element $x \in \mathbb{F}$, every element of $\mathbb{F}$ can be expressed in the form of $a x$ or $x a$ for some $a \in \mathbb{F}$. 
  \end{proof}

  The ring $\mathbb{Z}_n$ has all the properties of a field except the property of having inverses for all of its nonzero elements. This leads to the following theorem. 

  \begin{theorem}[Integer Quotient Rings as Finite Fields]
    The ring $(\mathbb{Z}_{n}, +, \times)$ is a field if and only if $n$ is a prime number. 
  \end{theorem}
  \begin{proof}
    $(\rightarrow)$ Assume that $n$ is composite $\implies n = k l$ for $k, n \in \mathbb{N} \implies k, n \neq 0$, but 
    \begin{equation}
      [k]_n [l]_n = [k l]_n = [n]_n = 0
    \end{equation}
    meaning that $\mathbb{Z}_n$ contains $0$ divisors and is not a field. The contrapositive of this states $(\rightarrow)$. \\
    $(\leftarrow)$ Given that $n$ is prime, let $[a]_n \neq 0$, i.e. $[a]_n \neq [0]_n, [1]_n$. The set of $n$ elements 
    \begin{equation}
      [0]_n, [a]_n, [2a]_n, ..., [(n-1)a]_n
    \end{equation}
    are all distinct. Indeed, if $[k a]_n = [l a]_n$, then $[(k-l) a]_n = 0 \implies n = (k-l) a \iff n$ is not prime. Since the elements are distinct, exactly one of them must be $[1]_n$, say $[p a]_n \implies$ the inverse $[p]_n$ exists. 
  \end{proof}

  \begin{corollary}[Invertibility in $\mathbb{Z}_n$]
    For any $n$, $[k]_n$ is invertible in the ring $\mathbb{Z}_n$ if and only if $n$ and $k$ are relatively prime. 
  \end{corollary} 

  We will talk about finite fields again, which are extremely important in Galois theory and in practical applications in e.g. cryptography. 

\subsection{Field of Fractions}

  Given an integral domain, there is a common way to construct a field from it. We simply just ``add'' all the multiplicative inverses. Doing so with the integers and polynomials creates the field of rational numbers and rational functions. 

\subsubsection{The Rational Numbers}

  Now that we've reviewed some fields, let's construct $\mathbb{Q}$ from $\mathbb{Z}$ and verify it's a field. 

  \begin{definition}[Rationals]
    Given the ordered ring of integers $(\mathbb{Z}, +_{\mathbb{Z}}, \times_{\mathbb{Z}}, \leq_{\mathbb{Z}})$ the \textbf{rational numbers} $(\mathbb{Q}, +_{\mathbb{Q}}, \times_{\mathbb{Q}})$ is the quotient space on $\mathbb{Z} \times \mathbb{Z} \setminus \{0\}$ with the equivalence relation $\sim$ 
    \begin{equation}
      (a, b) \sim (c, d) \iff a \times_{\mathbb{Z}} d = b \times_{\mathbb{Z}} c
    \end{equation} 
    and the operation defined 
    \begin{enumerate}
      \item The additive and multiplicative identities are 
      \begin{equation}
        0_{\mathbb{Q}} \coloneqq (0_{\mathbb{Z}}, a), \;\;\; 1_{\mathbb{Q}} \coloneqq (a, a)
      \end{equation}

      \item Addition on $\mathbb{Q}$ is defined 
      \begin{equation}
        (a, b) +_{\mathbb{Q}} (c, d) \coloneqq \big( (a \times_{\mathbb{Z}} d) +_{\mathbb{Z}} (b \times_{\mathbb{Z}} c), b \times_{\mathbb{Z}} d \big) 
      \end{equation}

      \item The additive inverse is defined 
      \begin{equation}
        -(a, b) \coloneqq (-a, b)
      \end{equation}

      \item Multiplication on $\mathbb{Q}$ is defined 
      \begin{equation}
        (a, b) \times_{\mathbb{Q}} (c, d) \coloneqq \big( a \times_{\mathbb{Z}} c, b \times_{\mathbb{Z}} d \big)
      \end{equation} 

      \item The multiplicative inverse is defined 
      \begin{equation}
        (a, b)^{-1} \coloneqq (b, a)
      \end{equation}
    \end{enumerate}
  \end{definition}

  \begin{theorem}[Field Structure]
    $\mathbb{Q}$ is a field. 
  \end{theorem}
  \begin{proof}
    We do a few things. 
    \begin{enumerate}
      \item Verify the additive identity. 
      \begin{equation}
        (a, b) + (0, c) = (ac + 0b, bc) = (ac, bc) \sim (a, b)
      \end{equation}
      \item Verify the multiplicative identity. 
      \begin{equation}
        (a, b) \times (c, c) = (ac, bc) \sim (a, b)
      \end{equation}
      \item Additive inverse is actually an inverse. 
      \begin{equation}
        (a, b) + (-a, b) = (ab + (-ba), bb) = (0, bb) \sim (0, 1)
      \end{equation}
      \item Multiplicative inverse is actually an inverse. 
      \begin{equation}
        (a, b) \times (b, a) = (ab, ba) = (ab, ab) \sim (1, 1)
      \end{equation}
      \item Addition is commutative. 
      \begin{equation}
        (a, b) + (c, d) = (ad + bc, bd) = (cb + ad, bd) = (c, d) + (a, b)
      \end{equation}
      \item Addition is associative. 
      \begin{align}
        (a, b) + ((c, d) + (e, f)) & = (a, b) + (cf + de, df) \\
                                   & = (adf + bcf + bde, bdf) \\
                                   & = (ad + bc, bd) + (e, f) \\
                                   & = ((a, b) + (c, d)) + (e, f)
      \end{align}
      \item Multiplication is commutative. 
      \begin{equation}
        (a, b) \times (c, d) = (ac, bd) = (ca, db) = (c, d) \times (a, b)
      \end{equation}
      \item Multiplication is associative. 
      \begin{align}
        (a, b) \times ((c, d) \times (e, f)) & = (a, b) \times (ce, df) \\ 
                                             & = (ace, bdf) \\
                                             & = (ac, bd) \times (e, f) \\
                                             & = ((a, b) \times (c, d)) \times (e, f)
      \end{align}
      \item Multiplication distributes over addition. 
        \begin{align}
          (a, b) \times ((c, d) + (e, f)) & = (a, b) \times (c, d) + (a, b) \times (e, f) \\
                                          & = (ac, bd) + (ae, bf) \\
                                          & = (abcf + abde, b^2 df) \\
                                          & = (acf + ade, bdf)  
                                          & = (a, b) \times (cf + de, df)
        \end{align}
    \end{enumerate}
  \end{proof}

  Great, so we have established that $\mathbb{Q}$ is a field. The next property we want to formalize is order. There are countless ways to do it, but I just take the difference and claim that it is greater than $0$. Note that given a set, we can really put whatever order we want on it. However, consider the field with the following order. 
  \begin{equation}
    \mathbb{F} = \{0, 1\}, \; 0 < 1
  \end{equation} 
  This does not behave well with respect to its operations because for example if we have $0 < 1$, then adding the same element to both sides should preserve the ordering. But this is not the case since $0 + 1 = 1 > 1 + 1 = 0$. While it may be easy to define an order, we would like it to be an ordered field. 

  \begin{definition}[Ordered Field]
    An \textbf{ordered field} is a field that has an order satisfying 
    \begin{enumerate}
      \item $y < z \implies x + y < x + z$ for all $x \in \mathbb{F}$. 
      \item $x > 0, y > 0 \implies xy > 0$. 
    \end{enumerate}
  \end{definition}

  \begin{theorem}[Properties]
    In an totally ordered field, 
    \begin{enumerate}
      \item $x > 0 \implies -x < 0$. 
      \item $x \neq 0 \implies x^2 > 0$. 
      \item If $x > 0$, then $y < z \implies xy < xz$. 
    \end{enumerate}
  \end{theorem} 
  \begin{proof}
    The first property is a single-liner 
    \begin{equation}
      0 < x \implies 0 + -x < x + -x \implies -x < 0 
    \end{equation}
    For the second property, it must be the case that $x > 0$ or $x < 0$. If $x > 0$, then by definition $x^2 > 0$. If $x < 0$, then 
    \begin{equation}
      x^2 = 1 \cdot x^2 = (-1)^2 \cdot x^2 = (-1 \cdot x)^2 = (-x)^2
    \end{equation}
    and since $-x > 0$ from the first property, we have $x^2 = (-x)^2 > 0$. For the third, we use the distributive property. 
    \begin{align}
      y < z & \implies 0 < z - y \\ 
            & \implies 0 = x 0 < x(z - y) = xz - xy \\
            & \implies xy < xz
    \end{align}
  \end{proof}

  \begin{theorem}[Ordered Field Structure]
    Second, $\mathbb{Q}$ is an ordered field. The order $\leq_{\mathbb{Q}}$ defined on the rationals as 
    \begin{equation}
      (a, b) \leq_{\mathbb{Q}} (c, d) \iff ad \leq_{\mathbb{Z}} bc
    \end{equation}
    is a total order. Remember that we have defined $b, d > 0$.  
  \end{theorem}
  \begin{proof}
    For the order property, we have 
    \begin{enumerate}
      \item Reflexive. 
      \begin{equation}
        (a, b) \leq_{\mathbb{Q}} (a, b) \iff ab \leq_{\mathbb{Z}} ab
      \end{equation} 

      \item Antisymmetric. 
      \begin{align}
        (a, b) \leq_{\mathbb{Q}} (c, d) & \implies ad \leq_{\mathbb{Z}} bc
        (c, d) \leq_{\mathbb{Q}} (a, b) & \implies bc \leq_{\mathbb{Z}} ad
      \end{align} 
      This implies that both $ad = bc$, which by definition means that they are in the same equivalence class. 

      \item Transitivity. Assume that $(a, b) \leq (c, d)$ and $(c, d) \leq (e, f)$. Then, we notice that $b, d, f > 0$ and therefore by the ordered ring property\footnote{If $a \leq b$ and $0 \leq c$, then $ac \leq bc$.} of $\mathbb{Z}$, we have 
      \begin{align}
        (a, b) \leq_{\mathbb{Q}} (c, d) & \implies ad \leq_{\mathbb{Z}} bc \implies adf \leq_{\mathbb{Z}} bcf \\ 
        (c, d) \leq_{\mathbb{Q}} (e, f) & \implies cf \leq_{\mathbb{Z}} de \implies bcf \leq_{\mathbb{Z}} bde
      \end{align}
      Therefore from transitivity of the ordering on $\mathbb{Z}$ we have $adf \leq bde$. By the ordered ring property\footnote{If $a \leq b$, then $a + c \leq b + c$.}  we have $0 \leq bde - adf = d(be - af)$. But notice that $d > 0$ from our definition of rationals, and therefore it must be the case that $0 \leq be - af \implies af \leq_{\mathbb{Z}} be$, which by definition means $(a, b) \leq_{\mathbb{Q}} (e, f)$. 
    \end{enumerate}
    For the ordered field property, we have 
    \begin{enumerate}
      \item Assume that $y = (a, b) \leq (c, d) = z$. Let $x = (e, f)$. Then $x + y = (af + be, bf)$, $x + z = (cf + de, df)$. Therefore 
      \begin{align}
        (af + be) df & = adf^2 + bedf \\ 
                     & \leq bcf^2 + bedf \\
                     & = (cf + de) bf
      \end{align} 
      But $(af + be) df = (cf + de) bf$ is equivalent to saying $(af + be, bf) \leq_{\mathbb{Q}} (cf + de, df)$, i.e. $x + y \leq x + z$!  

      \item Let $x = (a, b), y = (c, d)$. Since $0 < x, 0 < y$, by construction this means that $0 < a, 0 < c$ (since $b, d > 0$ in the canonical rational form). By the ordered ring property of the integers, $0 < ac$. So 
      \begin{equation}
        0 < ac \iff 0 \cdot bd < ac \cdot 1 \iff (0, 1) < (ac, bd)  \iff 0_{\mathbb{Q}} < (a, c) \times_{\mathbb{Q}} (b, d) = x y
      \end{equation}
    \end{enumerate}
  \end{proof} 

  \begin{theorem}[Vector Space Structure]
    Third, it is an inner product space over $\mathbb{Q}$. 
    \begin{enumerate}
      \item Addition and scalar multiplication is defined according to the field rules above. 
      \item The inner product, along with the induced norm and metric, are defined
      \begin{equation}
        \langle x, y \rangle \coloneqq xy, \quad |x| \coloneqq \begin{cases} x & \text{ if } x \geq 0 \\ -x & \text{ if } x < 0 \end{cases}, \quad d(x, y) \coloneqq |x - y|
      \end{equation}
      \item The topology induced by the metric is the set of open intervals $(x, y)$, which also coincides with the order topology. 
    \end{enumerate}
  \end{theorem}

  We have successfully defined the rationals, but now these are almost completely separate elements. We know that all integers are rational numbers, and so to show that the rationals are an extension of $\mathbb{Z}$ we want to identify a \textit{canonical injection} $\iota: \mathbb{Z} \rightarrow \mathbb{Q}$. This can't just be any canonical injection; it must preserve the algebraic structure between the two sets and must therefore be a \textit{ring homomorphism}. 

  \begin{theorem}[Canonical Injection of $\mathbb{Z}$ to $\mathbb{Q}$ is an Ordered Ring Homomorphism]
    Let us define the canonical injection $\iota: \mathbb{Z} \rightarrow \mathbb{Q}$ to be $\iota(a) = (a, 1)$. This is a ring homomorphism. Additionally, it preserves order: for $a, b \in \mathbb{Z}$, 
    \begin{equation}
      a \leq_{\mathbb{Z}} b \iff \iota(a) \leq_{\mathbb{Q}} \iota(b)
    \end{equation}
  \end{theorem}
  \begin{proof} 
    We show a few things. 
    \begin{enumerate}
      \item Preservation of addition. 
        \begin{align}
          \iota(a) +_{\mathbb{Q}} \iota(b) & = (a, 1) +_{\mathbb{Q}} (b, 1) \\
                                           & = (1a +_{\mathbb{Z}} 1b, 1^2) \\
                                           & = (a +_{\mathbb{Z}} b, 1) \\
                                           & = \iota(a +_{\mathbb{Z}} b) 
        \end{align}
      \item Preservation of multiplication. 
        \begin{align}
          \iota(a) \times_{\mathbb{Q}} \iota(b) & = (a, 1) \times_{\mathbb{Q}} (b, 1) \\
                                                & = (a \times_{\mathbb{Z}} b, 1^2) \\
                                                & = (a \times_{\mathbb{Z}} b, 1) \\
                                                & = \iota(a \times_{\mathbb{Z}} b, 1)
        \end{align}
      \item Preservation of multiplicative identity. 
        \begin{equation}
          \iota(1_{\mathbb{Z}}) = (1, 1) = 1_{\mathbb{Q}}
        \end{equation}
    \end{enumerate}
    To show that it preserves the order, we have 
    \begin{align}
      a \leq_{\mathbb{Z}} b & \iff a \cdot 1 \leq_{\mathbb{Z}} b \cdot 1 \\
                            & \iff (a, 1) \leq_{\mathbb{Q}} (b, 1) \\
                            & \iff \iota(a) \leq_{\mathbb{Q}} \iota(b)
    \end{align}
  \end{proof} 

  \begin{example}[Numbers]
    The rationals, reals, and complex numbers are all fields.\footnote{Quaternions are not!}
  \end{example}

  Note the subfield structure $\mathbb{Q} \subset \mathbb{R} \subset \mathbb{C}$. However, we will find that there are tons of other fields lurking in between $\mathbb{Q}$ and $\mathbb{C}$ other than $\mathbb{R}$. We can actually say that there are no subfields of $\mathbb{Q}$. 

  \begin{lemma}[Rationals are a Minimal Field]
    Every subfield of $\mathbb{C}$ contains $\mathbb{Q}$. 
  \end{lemma}
  \begin{proof}
    Must contain $0$ and $1$. Keep adding $1$ and inverting it to get $\mathbb{Z}$. Now $\mathbb{Z}$ must contain units so $1/n$ also contained. Then multiply the elements to get $\mathbb{Q}$. 
  \end{proof} 

  \begin{theorem}[Finite Fields]
    There are no finite ordered fields. 
  \end{theorem} 
  \begin{proof}
    Assume $\mathbb{F}$ is such an ordered field. It must be the case that $0, 1 \in \mathbb{F}$, with $0 < 1$. Therefore, we also have $0 + 1 < 1 + 1 \implies 1 < 1 + 1$. Repeating this we get 
    \begin{equation}
      0 < 1 < 1 + 1 < 1 + 1 + 1 < \ldots
    \end{equation}
    where these elements must be distinct (since only one of $>, <, =$ must be true for a totally ordered set). Since this can be done for a countably infinite number of times, $\mathbb{F}$ cannot be finite. 
  \end{proof}

\subsubsection{Rational Functions}

  Given a field $F$, we have constructed the Euclidean domain $F[x]$. However, this is one step away from being a field. We mimick the construction of the rational numbers $\mathbb{Q}$ as a quotient space over $\mathbb{Z} \times (\mathbb{Z} \setminus \{0\})$ by taking $F[x] \times (F[x] \setminus \{0\})$ and putting a quotient on it. 
  
  \begin{definition}[Rational Functions]
    The \textbf{rational functions} are defined to be the field of quotients (really just 2-tuples) of the form 
    \begin{equation}
      F(x) \coloneqq \bigg\{ \frac{f(x)}{g(x)} \; \bigg| \; f(x), g(x) \in F[x], g(x) \neq 0 \bigg\}
    \end{equation}
    where addition and multiplication is defined in the usual sense.
  \end{definition}

  \begin{theorem}[Partial Fractions Decomposition]
    Let $f(x), g(x) \in F[x]$ where $\deg(f(x)) < \deg(g(x))$. If $g(x) = u(x) v(x)$ where $u, v$ are relatively prime, then there are polynomials $a(x), b(x)$ with $\deg(a) < \deg(u), \deg(b) < \deg(v)$ s.t. 
    \begin{equation}
      \frac{f(x)}{g(x)} = \frac{a(x)}{u(x)} + \frac{b(x)}{v(x)}
    \end{equation}
    By induction, we can prove this for any finite set of irreducible polynomials. 
  \end{theorem}
  \begin{proof}
    We describe an algorithm to get this decomposition. There are polynomials $s(x), t(x)$ s.t. $1 = s(x) u(x) + t(x) v(x)$. Therefore, 
    \begin{equation}
      \frac{f(x)}{ u(x) v(x)} = \frac{f(x) t(x)}{u(x)} + \frac{f(x) s(x)}{v(x)}
    \end{equation}
    and we can use the Euclidean algorithm to write 
    \begin{align}
      \frac{f(x) t(x)}{u(x)} & = q(x) + \frac{a(x)}{u(x)}, \qquad \deg(a) < \deg(u) \\
      \frac{f(x) s(x)}{v(x)} & = q(x) + \frac{a(x)}{u(x)}, \qquad \deg(b) < \deg(v)
    \end{align}
    which implies 
    \begin{equation}
      \frac{f(x)}{u(x) v(x)} = \frac{a(x)}{u(x)} + \frac{b(x)}{v(x)}
    \end{equation}
  \end{proof}

  \begin{example}
    Consider the rational function $\frac{x + 3}{x^3 (x - 1)^2}$. Applying the Euclidean algorithm, we find that 
    \begin{equation}
      1 = (3x^2 + 2x + 1) (x - 1)^2 - (3x - 4) x^3
    \end{equation}
    and so 
    \begin{align}
      \frac{x + 3}{x^3 (x - 1)^2} & = \frac{(x + 3)(3x^2 + 2x + 1)}{x^3} - \frac{(x + 3)(3x - 4)}{(x - 1)^2} \\
                                  & = \frac{11x^2 + 7x + 3}{x^3} + \frac{-11x + 15}{(x - 1)^2}
    \end{align}
  \end{example}

\subsection{Ordered Fields}

\subsection{Field Extensions as Vector Spaces}

  Now that we have established the theory of general rings and for polynomials, we will delve deeper into the theory of fields by talking about \textit{field extensions}, which allows us to model fields as vector spaces. This will help us in classifying fields along with providing the foundations of Galois theory. This method in which we have taken higher powers of $\alpha$ to reveal elements in $\mathbb{Q}$ reveals a deeper structure of a finite-dimensional vector space, which will be useful for analyzing certain fields in the examples below. Note that a vector space is only well-defined over a field $F$, so we will consider \textit{field extensions} from now on. First, recall that a field is trivially a vector space. 

  \begin{lemma}[Fields are a Vector Space]
    A field $F$ is a 1-dimensional vector space over itself. 
  \end{lemma} 

  It turns out that we can generalize this a bit more. 

  \begin{theorem}[Fields are a Vector Space over Subfields]
    \label{thm:fields_vector_space}
    Let $F$ be a subfield of $K$. Then $K$ is a $F$-vector space. 
  \end{theorem}
  \begin{proof}
    A $F$-vector space has $0$, addition, and multiplication by $F$. $K$ indeed has $0$, addition, and we can multiply any element of $K$ by an element of $F$. The extra axioms follow but are too verbose to write a full proof. 
  \end{proof}

  \begin{corollary} 
    $\mathbb{R}$ is an infinite-dimensional vector space over $\mathbb{Q}$. 
  \end{corollary}
  \begin{proof}
    The fact that it is a vector space immediately follows from $\mathbb{Q} \hookrightarrow \mathbb{R}$. For dimensionality, the outline is to show that $\{\sqrt{p} \mid p \text{ prime }\}$ are linearly independent. This takes work to prove and won't do it. 
  \end{proof}

  Therefore, by constructing a subfield, we can model the original field as a vector space. This additional structure warrants a name. 

  \begin{definition}[Field Extension]
    If $F \subset K$ are fields, then this is called a \textbf{field extension}. Its \textbf{degree} is the $F$-dimension of $K$, denoted
    \begin{equation}
      [K:F] \coloneqq \dim_F (K) 
    \end{equation}
  \end{definition} 

  So when we are given a field, we can automatically treat it as a vector space. Furthermore, if we are given a field extension, we can treat the larger field as a vector space over the smaller field (though it may be finite or infinite-dimensional). As we create concatenated field extensions, sometimes called a \textit{tower}, the dimensions behave nicely as well. 

  \begin{theorem}[Tower Rule]
    $E \hookrightarrow F \hookrightarrow K$ are field extensions. Then $E \hookrightarrow K$ is a field extension with degree 
    \begin{equation}
      [K:E] = [K:E] [E:F]
    \end{equation}
  \end{theorem}
  \begin{proof}
    Let $\alpha_1, \ldots \alpha_m$ be a basis for $E$ over $F$ and $\beta_1, \ldots, \beta_n$ be a basis for $K$ over $E$. We claim that $\{\alpha_i \beta_j\}$ is a basis for $K$ over $F$, with multiplication done in the field $K$. We check linear indepdence. Let $\beta \in K$ be arbitrary. Then by the $E$-basis, we have 
    \begin{equation}
      \beta = \sum_{j=1}^n x_j \beta_j
    \end{equation} 
    But since $x_j \in E$, there are elements 
    \begin{equation}
      x_j = \sum_{i=1}^m y_{ij} \alpha_i
    \end{equation}
    and so combining we get 
    \begin{equation}
      \beta = \sum_{j=1}^n \sum_{i=1}^m y_{ij} (\alpha_i \beta_j) 
    \end{equation}
    To prove linear independence, suppose $\beta = 0$. Then we have 
    \begin{equation}
      0 = \sum_{j=1}^n \sum_{i=1}^m y_{ij} (\alpha_i \beta_j) = \sum_{j=1}^n \bigg( \sum_{i=1}^m y_{ij} \alpha_i \bigg) \beta_j 
    \end{equation}
    Since $\beta_1, \ldots, \beta_n$ are linearly indpendent, we must have $\sum_{i=1}^m y_{ij} \alpha_i = 0$ for all $j$. But since $\alpha_i$'s are linear independent, this means $y_{ij} = 0$ for all $i, j$. 
  \end{proof} 

\subsection{Finite Fields} 

  \begin{theorem}[Wedderburn's little theorem]
    Every finite integral domain is a field. 
  \end{theorem} 
  \begin{proof}
    Let $R$ be a finite integral domain and $a \in R$ be nonzero. Since every element is regular, the map $x \mapsto ax$ is an injective function. Since $R$ is finite this map is also surjective. In particular, there is some $b \in R$ s.t. $ab = 1$, i.e. $a$ is a unit. 
  \end{proof}

  We know that a field---as an integral domain---has characteristic $0$ or prime $p$. We also know that a field is a vector space, at least over itself. But now that we have shown that a field can be modeled as a vector space, we can use this to say something more specific about finite fields, which will lead to a complete classification of finite fields. 

  \begin{theorem}[Characteristic Determines Base Field of Vector Space]
    \label{thm:char_field}
    Given a field $F$, 
    \begin{enumerate}
      \item If $\Char(F) = p$, then $F$ is a vector space over $\mathbb{Z}_p$. 
      \item If $\Char(F) = 0$, then $F$ is a vector space over $\mathbb{Q}$. 
    \end{enumerate}
  \end{theorem}
  \begin{proof}
    
  \end{proof} 

  Therefore, just from the characteristic we can classify all fields as vector spaces over either $\mathbb{Q}$ or $\mathbb{Z}_p$. Now if we focus on finite fields, we can do a reverse classification. 

  \begin{theorem}[Finite Fields Have Cardinality $p^d$]
    Let $F$ be a finite field. Then $|F| = p^n$ for some $n \in \mathbb{N}$. 
  \end{theorem}
  \begin{proof}
    $F$ is a vector space over $\mathbb{Z}_p$ from \ref{thm:char_field}. Since $F$ has finitely many elements, $F$ has a finite spanning set, which implies $\dim_{\mathbb{Z}_p} F \leq + \infty$. Let $d$ be the dimension and $\{b_1, \ldots, b_d\}$ be the basis. The elements of $F$ are 
    \begin{equation}
      a_1 b_1 + \ldots + a_d b_d
    \end{equation}
    with $a_1, \ldots a_d \in \mathbb{Z}_p$. Thus there are $p^d$ elements of $F$, so $F \simeq \mathbb{Z}_p^d$. 
  \end{proof}

  In fact, for \textit{every} prime power there exists a unique field. Therefore we can create a bijection by proving the converse. 

  \begin{theorem}[Field for Every $p^d$]
    For every prime $p$ and $n \in \mathbb{N}$, there exists a field with $q = p^d$ elements, unique up to isomorphism.  
  \end{theorem} 
  \begin{proof} 
    Let $f(x) = x^q - x \in \mathbb{Z}_p [x]$. Then this polynomial has a splitting field $K \supset \mathbb{Z}$. Now we claim the roots of $f(x)$ in $K$ are distinct and form a subfield $F_q \subset K$. This will complete the proof since $F_q \subset K$ and $K \subset F_q \implies K = F_q$. Assume $\alpha, \beta \in K$ are roots of $f(x)$, and so $\alpha^p = \alpha$ and $\beta^p = \beta$
    \begin{enumerate}
      \item $\alpha + \beta \in K$ since by a modification of Freshman's dream, $(\alpha + \beta)^p = \alpha^p + \beta^p = \alpha + \beta$.\footnote{We induct on $n$ for $q = p^n$. For $n=1$, this is trivial by Freshmans dream. Now assume it holds for some $n \in \mathbb{N}$. Then $(x + y)^{p^{n+1}} = ( (x + y)^{p^n} )^p = (x^{p^n} + y^{p^n})^p = (x^{p^n})^p + (y^{p^n})^p = x^{p^{n+1}} + y^{p^{n+1}}$. }
      \item $(-\alpha)^q = (-1)^q \alpha^q = (-1)^q \alpha = -\alpha$ since $-1 = 1$ or $q$ is odd. 
      \item $\alpha \beta \in K$ since $\mathbb{Z}_p$ is a field and so $(\alpha \beta)^p = \alpha^p \beta^p = \alpha \beta$. 
      \item For multiplicative inverses, let $\alpha \neq 0$. Then $(\alpha^{-1})^p = (\alpha^{p})^{-1} = \alpha^{-1}$. 
      \item For all $p$, $0$ and $1$ are roots so $0, 1 \in K$. 
    \end{enumerate}
    Now we show that $K$ consists of distinct roots. Certainly $0 \in K$ with multiplicity $1$ since $f(x) = x (x^{q-1} - x)$. Now suppose nonzero $r \in K$ is a root with multiplicity $m$. The multiplicity of $r$ is the multiplicity of $0$ of 
    \begin{equation}
      f(x + r) = (x + r)^q - (x + r) = x^q + r^q - x - r = x^q - x
    \end{equation}
    where the final step follows from $0 = r^q - r$ since $r \in K$. Therefore $r$ has multiplicity $1$. Since $K[x]$ has unique factorization property, it follows that $m=1$ and every $r$ is a simple root. 

    To show that every field with $p^n$ elements is unique, let $F$ be such a field. We claim that $\Char(F) = p \implies \mathbb{Z}_p \subset F$. We claim that every element of $F$ is a root of $f(x) = x^q - x \in \mathbb{Z}_p [x]$, where $F$ is the splitting field. Let $G = F^\ast$ be the multiplicative group of units. Since $F$ is a field, then $|F^\ast| = |F| - 1 = p^d - 1$, and by constructing the cyclic group $\langle g \rangle \subset G$ for any $g \in G$, we know by Lagrange's theorem that $g^{|G|} = 1_G$, which implies that for all $x \in F$, 
    \begin{enumerate}
      \item If $x \neq 0$ then $x^{p^d - 1} = x \implies x^{p^d} = x$ and so $x \in K$. 
      \item If $x = 0$ then $x^{p^d} - x = 0$ and so $x \in K$. 
    \end{enumerate}
    Therefore $F \subset K$ with $|F| = |K|$ both finite, and so $F = K$. 
  \end{proof} 

  From this, we can write for every prime $p$ and natural $n$ the finite field of order $p^n$ as $\mathbb{F}_{p^n}$. It is clear that if $n = 1$ then $\mathbb{F}_p \simeq \mathbb{Z}_p$. The final result we will show is a hierarchy of subfields. 

  \begin{theorem}[Hierarchy of Fields]
    For a given prime $p$, if $p^m < p^n$, then 
    \begin{equation}
      F_{q^m} \subset F_{q^n} \iff m \mid n
    \end{equation}
  \end{theorem}

