\section{Galois Theory}

  In general, given a polynomial $f(x)$, we would like to find effective ways to factor it or find its roots. This depends on which polynomial ring $R[x]$ that we view $f(x)$ as an element of, and finding such a unique factorization requires us to work in UFDs at least. But solving even one of problems is quite hard with ring theory alone, and we need to combine our ideas of polynomial rings, groups, and vector spaces to develop effective means of solving polynomials. This is known as \textit{Galois theory}. 

  One nice property where finding the roots and the factorization coincide is when $f(x)$ splits in $R[x]$, and we usually construct these rings that allow $f(x)$ to split by \textit{extending} existing rings $R$ into a bigger ring $S$. 

  \begin{definition}[Ring Extension]
    A pair of rings $R, S$ where $R$ is a subring of $S$ is called a \textbf{ring extension}, denoted $R \hookrightarrow S$, $S/R$ or $R \subset S$. If we have multiple extensions $R \subset S \subset T$, this is called a \textbf{tower}. 
  \end{definition}

  We have already come across several examples of extensions. 

  \begin{example}[Ring Extensions]
    Listed.
    \begin{enumerate}
      \item $\mathbb{Z} \subset \mathbb{Q}$ is a ring extension.
      \item $\mathbb{Z} \subset \mathbb{Z}[i]$ is a ring extension. 
      \item $\mathbb{Q} \subset \mathbb{R} \subset \mathbb{C}$ is a ring extension tower. 
    \end{enumerate}
  \end{example}

  Now we present the motivation for working with extensions of fields that rings in general. Recall that a field is trivially a vector space, followed by the more general result. 

  \begin{theorem}[Fields are a Vector Space over Subfields]
    \label{thm:fields_vector_space}
    Let $F$ be a subfield of $K$. Then $K$ is a $F$-vector space. 
  \end{theorem}
  \begin{proof}
    A $F$-vector space has $0$, addition, and multiplication by $F$. $K$ indeed has $0$, addition, and we can multiply any element of $K$ by an element of $F$. The extra axioms follow but are too verbose to write a full proof. 
  \end{proof}

  \begin{corollary} 
    $\mathbb{R}$ is an infinite-dimensional vector space over $\mathbb{Q}$. 
  \end{corollary}
  \begin{proof}
    The fact that it is a vector space immediately follows from $\mathbb{Q} \hookrightarrow \mathbb{R}$. For dimensionality, the outline is to show that $\{\sqrt{p} \mid p \text{ prime }\}$ are linearly independent. This takes work to prove and won't do it. 
  \end{proof}

  Therefore, by constructing a subfield $F$ of a field $K$, we can model $K$ as a $F$-vector space (though it may be finite or infinite-dimensional). This additional structure warrants a name. 

  \begin{definition}[Field Extension]
    If $F \subset K$ are fields, then this is called a \textbf{field extension}. Its \textbf{degree} is the $F$-dimension of $K$, denoted
    \begin{equation}
      [K:F] \coloneqq \dim_F (K) 
    \end{equation}
  \end{definition} 

  As we concatenate field extensions, sometimes called a \textit{tower}, the dimensions behave nicely as well. 

  \begin{theorem}[Tower Rule]
    If $E \hookrightarrow F \hookrightarrow K$ are two field extensions, then $E \hookrightarrow K$ is a field extension with degree 
    \begin{equation}
      [K:E] = [K:E] [E:F]
    \end{equation}
  \end{theorem}
  \begin{proof}
    Let $\alpha_1, \ldots \alpha_m$ be a basis for $E$ over $F$ and $\beta_1, \ldots, \beta_n$ be a basis for $K$ over $E$. We claim that $\{\alpha_i \beta_j\}$ is a basis for $K$ over $F$, with multiplication done in the field $K$. We check linear indepdence. Let $\beta \in K$ be arbitrary. Then by the $E$-basis, we have 
    \begin{equation}
      \beta = \sum_{j=1}^n x_j \beta_j
    \end{equation} 
    But since $x_j \in E$, there are elements 
    \begin{equation}
      x_j = \sum_{i=1}^m y_{ij} \alpha_i
    \end{equation}
    and so combining we get 
    \begin{equation}
      \beta = \sum_{j=1}^n \sum_{i=1}^m y_{ij} (\alpha_i \beta_j) 
    \end{equation}
    To prove linear independence, suppose $\beta = 0$. Then we have 
    \begin{equation}
      0 = \sum_{j=1}^n \sum_{i=1}^m y_{ij} (\alpha_i \beta_j) = \sum_{j=1}^n \bigg( \sum_{i=1}^m y_{ij} \alpha_i \bigg) \beta_j 
    \end{equation}
    Since $\beta_1, \ldots, \beta_n$ are linearly indpendent, we must have $\sum_{i=1}^m y_{ij} \alpha_i = 0$ for all $j$. But since $\alpha_i$'s are linear independent, this means $y_{ij} = 0$ for all $i, j$. 
  \end{proof} 

  In fact, it is a natural question to find the ``minimal field'' $F$ such that $f(x)$ splits in $F[x]$, which is called the \textit{splitting field} of $f(x)$. To do this, we must start with $F$ and extend it to $K$ such that $K$ is a field. This is done through \textit{adjoining} an element $\alpha$ to $F$, getting a larger set $F[\alpha]$. This guarantees both a ring and vector space structure but requires another condition (that $\alpha$ be a root of an irreducible polynomial in $F[x]$) to be a field. Once this is done, we have successfully created a field extension, and finding the splitting field is not too hard from here. 

  What's next? Now that we can construct the splitting field $K/F$, we want to try and find some \textit{symmetries} between its roots. That is, if $\alpha, \beta$ are the roots of some $f(x) \in F[x]$, can we find a transformation group $G$ such that knowing one root allows us to uncover the other roots? There is in fact such a structure out there, known as the \textit{Galois group}---an automorphism group on the finite set of roots. This will be our roadmap in this section. 

\subsection{Ring Extensions Through Adjoining} 

  We deal with the first problem of constructing ring extensions. We have already seen the idea of taking ring $R$ and adjoining an element $x$ to it to get a minimal ring containing both $R$ (as a subring) and $x$. This was the polynomial ring. Now we can replace the indeterminate $x$ with an element $\alpha$ from a larger ring $S/R$ and do the exact same thing to get a ring extension of $R$. 

  \begin{definition}[Adjoining Ring] 
    Given ring extension $R \hookrightarrow S$ with finite $\alpha = \{\alpha_1, \ldots, \alpha_n\} \subset S$, the \textbf{ring $R$ adjoined by $\alpha$} is defined in the two equivalent ways. 
    \begin{enumerate}
      \item $R[\alpha]$ is the minimal ring containing $R$ and $\alpha$. 

      \item We take the $\alpha_i$'s and map it through all polynomials in $R[x_1, \ldots, x_n]$. 
      \begin{equation}
        R[\alpha] = R[\alpha_1, \ldots, \alpha_n] \coloneqq \{ f(\alpha_1, \ldots, \alpha_n) \in S \mid f \in R[x_1, \ldots, x_n]\} \subset S
      \end{equation} 

      \item We can iteratively construct $R[\alpha]$.\footnote{Note that if $R \hookrightarrow S$ is a ring extension, then $R[\alpha]$ is a ring, and so it makes sense to write $R[\alpha][\beta] \subset S$. It had better be the case that this ring is consistent with equivalent constructions.} 
      \begin{equation}
        R \subset R[\alpha_1] \subset R[\alpha_1, \alpha_2] \subset \ldots \subset R[\alpha_1, \ldots, \alpha_n] 
      \end{equation}
    \end{enumerate}
  \end{definition} 
  \begin{proof}
    We show a few things. 
    \begin{enumerate}
      \item \textit{$R[\alpha]$ in (2) is a ring}. Given two elements $\phi, \gamma \in F[\alpha]$, there exists polynomials $f, g \in F[x]$ s.t. $\phi = f(\alpha), \gamma = g(\alpha)$. Since $F[x]$ is a ring, we see that 
      \begin{align}
        \phi + \gamma & = f(\alpha) + g(\alpha) = (f + g)(\alpha) \\
        \phi \cdot \gamma & = f(\alpha) \cdot g(\alpha) = (fg)(\alpha)
      \end{align} 
      Furthermore, it is easy to check that $0$ and $1$ are the images of $\alpha$ through the $0$ and $1$ polynomials. 
      \item \textit{(1) $\implies$ (2)}. This is pretty obvious since $R$ is a subring of (2) and $\alpha$ is just the image of $\alpha$ under $f(x) = x$. 
      \item \textit{(2) $\iff$ (3)}. By induction, it suffices to show that $R[\alpha, \beta] = R[\alpha][\beta] = R[\beta][\alpha]$. 
      \item \textit{(3) $\implies$ (1)}. By induction, it suffices to show that $R[\alpha_1]$ is the minimal field a
    \end{enumerate}
  \end{proof} 

  What allows us to make this inclusion proper is that the $\alpha \in K$, which does not necessarily have to be in $F$, \textit{extends} this field a bit further, but since we can only map the one element $\alpha$, it may not cover all of $K$. Most of the times, we will work with adjoining rings of univariate polynomial elements. 
  \begin{equation}
    F[\alpha] \coloneqq \{ f(\alpha) \in F \mid f \in F[x]\} \subset K
  \end{equation} 
  Let's go through some examples.

  \begin{example}[Radical Extensions of $\sqrt{2}$]
    Let $F = \mathbb{Q}$ and $K = \mathbb{C}$. We claim $\mathbb{Q}[\sqrt{2}] = \{a + b \sqrt{2} \mid a, b \in \mathbb{Q} \}$.
    \begin{enumerate}
      \item $\mathbb{Q}[\sqrt{2}] \subset \{a + b \sqrt{2} \mid a, b \in \mathbb{Q} \}$. $\mathbb{Q}[\sqrt{2}]$ are elements of the form
      \begin{equation}
        f(\sqrt{2}) = a_n (\sqrt{2})^n + a_{n-1} (\sqrt{2})^{n-1} + \ldots + a_2 (\sqrt{2})^2 + a_1 \sqrt{2} + a_0
      \end{equation} 
      This can be written by collecting terms, of the form $a + b \sqrt{2}$. 

      \item $\mathbb{Q}[\sqrt{2}] \supset \{a + b \sqrt{2} \mid a, b \in \mathbb{Q} \}$. Given an element $a + b \sqrt{2}$, this is clearly in $\mathbb{Q}[\sqrt{2}]$ since it is the image of $\sqrt{2}$ under the polynomial $f(x) = a + bx$. 
    \end{enumerate}
  \end{example} 

  Given this, we may extrapolate this pattern and claim that $\mathbb{Q}[\sqrt{2} + \sqrt{3}]$ consists of all numbers of form $a + (\sqrt{2} + \sqrt{3}) b$. However, this is \textit{not} the case. 

  \begin{example}[Radical Extensions of $\sqrt{2} + \sqrt{3}$]
    Given any element $\beta \in \mathbb{Q}[\sqrt{2} + \sqrt{3}]$, it is by definition of the form 
    \begin{equation}
      \beta = \sum_{k=0}^n a_k (\sqrt{2} + \sqrt{3})^k 
    \end{equation} 
    Clearly $1, \sqrt{2} + \sqrt{3} \in \mathbb{Q}[\sqrt{2} + \sqrt{3}]$ by mapping $\sqrt{2} + \sqrt{3}$ through the polynomials $f(x) = 1$ and $f(x) = $. However, we can see that $(\sqrt{2} + \sqrt{3})^2 = 5 + \sqrt{6}$,\footnote{where we use $\sqrt{6}$ as notation for $\sqrt{2} \cdot \sqrt{3}$} and so $\sqrt{6} \in \mathbb{Q}[\sqrt{2} + \sqrt{3}]$. Furthermore, we have $(\sqrt{2} + \sqrt{3})^3 = 11 \sqrt{2} + 9 \sqrt{3}$, and so with the ring properties we can conclude that 
    \begin{align}
      \frac{1}{2} \big[ (11 \sqrt{2} + 9 \sqrt{3}) - 9 (\sqrt{2} + \sqrt{3})\big] = \sqrt{2} & \in \mathbb{Q}[\sqrt{2} + \sqrt{3}] \\
      -\frac{1}{2} \big[ (11 \sqrt{2} + 9 \sqrt{3}) - 11 (\sqrt{2} + \sqrt{3})\big] = \sqrt{3} & \in \mathbb{Q}[\sqrt{2} + \sqrt{3}] \\
    \end{align} 
    If we go a bit further, we can show that 
    \begin{equation}
      \mathbb{Q}[\sqrt{2} + \sqrt{3}] = \{a + b \sqrt{2} + c \sqrt{3} + d\sqrt{6} \mid a, b, c, d \in \mathbb{Q} \}
    \end{equation}
  \end{example}

  \begin{example}[Showing Two Extensions are Equal]
    We claim that $\mathbb{Q}[\sqrt{6}, \sqrt{5}] = \mathbb{Q}[\sqrt{6} + \sqrt{5}]$. 
    \begin{enumerate}
      \item $\mathbb{Q}[\sqrt{6}, \sqrt{5}] \subset \mathbb{Q}[\sqrt{6} + \sqrt{5}]$ because the sum of $\sqrt{6}$ and $\sqrt{5}$ lies in $\mathbb{Q}[\sqrt{6}, \sqrt{5}]$. 
      \item $\mathbb{Q}[\sqrt{6} + \sqrt{5}] \subset \mathbb{Q}[\sqrt{6}, \sqrt{5}]$. We have 
        \begin{equation}
          \frac{1}{\sqrt{5} + \sqrt{6}} = \frac{1}{\sqrt{5} + \sqrt{6}} \frac{\sqrt{5} - \sqrt{6}}{\sqrt{5} - \sqrt{6}} = \sqrt{6} - \sqrt{5} \in \mathbb{Q}[\sqrt{5} + \sqrt{6}] 
        \end{equation}
        which implies that 
        \begin{equation}
          \sqrt{6} = \frac{1}{2} ((\sqrt{6} + \sqrt{5}) + (\sqrt{6} - \sqrt{5})) \in \mathbb{Q}[\sqrt{5} + \sqrt{6}] 
        \end{equation}
        and so $\sqrt{5} = (\sqrt{6} + \sqrt{5}) - \sqrt{6} \in \mathbb{Q}[\sqrt{5} + \sqrt{6}]$. 
    \end{enumerate}
  \end{example}

  So far, we've gotten used to modeling fields as vector spaces. In general, an adjoining ring $R[\alpha]$ is another ring containing $R$ as a subring. Now if $F$ is a field, then $F \subset F[\alpha]$ is a ring extension, and since we are working with the base field $F$, we can model $F[\alpha]$ as a vector space. 

  \begin{theorem}[Adjoining Field is Finite-Dimensional Vector Space]
    If $F \subset K$ is a field and $\alpha \in K$, then 
    \begin{enumerate}
      \item $F[\alpha]$ is a \textit{finite-dimensional} vector space over $F$. 
      \item If $f(x) = a_n x^n + \ldots + a_0$ is any polynomial with root $\alpha$, then $\{1, \alpha, \ldots, \alpha^{n-1}\}$ spans $F[\alpha]$.\footnote{Note that this does not mean that it is a basis.} 
    \end{enumerate}
  \end{theorem}
  \begin{proof}
    An element of $F[\alpha]$ is of the form 
    \begin{equation}
      f(\alpha) = \sum_{k=0}^n a_k \alpha^k
    \end{equation} 
    for some $f \in F[x]$, and so it is immediate that $\{\alpha^k\}_{k \in \mathbb{N}_0}$ spans $F[\alpha]$. We claim that $\alpha^{n-1+i}$ is in $S$ for all $i > 0$. By induction, if $i = 1$, then 
    \begin{equation}
      \alpha^n = -\frac{1}{a_n} \big( a_{n-1} \alpha^{n-1} + \ldots + a_0 \big)
    \end{equation}
    which proves the claim. Now assume that $\alpha^n, \alpha^{n+1}, \ldots, \alpha^{n-1+i} \in \Span\{1, \ldots, \alpha^{n-1}\}$. Then 
    \begin{equation}
      \alpha^i f(\alpha) = 0 \implies a_n \alpha^{n+i} + \alpha_{n-1} \alpha^{n+i-1} + \ldots + a_0 \alpha^i = 0 
    \end{equation}
    and so 
    \begin{equation}
      \alpha^{n+i} = -\frac{1}{a_n} \big(a_{n-1} \alpha^{n+i-1} + \ldots + a_0 \alpha^i)
    \end{equation}
    which means that $\alpha^{n+i} \in \Span\{1, \ldots, \alpha^{n-1}\}$, completing the proof. 
  \end{proof} 

\subsection{Field Extensions}

  This is nice since we have a vector space structure on $F[\alpha]$ unlike just a ring structure on $R[\alpha]$. What we should think is that $F[\alpha]$ ends up becoming both a ring and a vector space, but not yet a field. We would like to find conditions in which $F[\alpha]$ indeed does become a field, which at this point it is commonly denoted $F(\alpha)$ (rather than square brakcets to emphasize it is a field). It turns out that it will happen if any only if $\alpha$ is \textit{algebraic}.  

  \begin{definition}[Algebraic Numbers]
    Given a field extension $F \subset K$, an element $\alpha \in K$ is \textbf{algebraic over $F$} if $\alpha$ is some root of $f(x) \in F[x]$.\footnote{By default we have $F = \mathbb{Q}$.}
  \end{definition}

  \begin{example}[Algebraic Numbers]
    We list a few examples. 
    \begin{enumerate}
      \item $\sqrt{2} \in \mathbb{R}$ is algebraic over $\mathbb{Q}$ since it is a root of $x^2 - 2 \in \mathbb{Q}[x]$. 
      \item $i \in \mathbb{C}$ is algebraic over $\mathbb{R}$ since it is a root of $x^2 + 1 \in \mathbb{R}[x]$. It is also a root of $x^4 + 2 x^2 + 1 \in \mathbb{R}[x]$. 
    \end{enumerate}
  \end{example}

  Note that given an element $\alpha \in K$ that is algebraic over $F$, there may be multiple polynomials $f(x) \in F[x]$ that has $\alpha$ as a root. In fact, this set forms an ideal. 

  \begin{lemma}[Polynomials Vanishing at $\alpha$ Forms an Ideal]
    Let $K/F$ be a field extension and fix $\alpha \in K$ that is algebraic over $F$. Then the set of polynomials $f(x) \in F[x]$ with root $\alpha$ forms an ideal over $F[x]$. 
  \end{lemma}
  \begin{proof}
    Let us denote the set as $I$. We prove the two properties of ideals.  
    \begin{enumerate}
      \item Consider $f(x), g(x) \in I$. Then $(f + g)(\alpha) = f(\alpha) + g(\alpha) = 0 + 0 = 0 \implies (f + g)(x) \in I$. 
      \item Consider $h(x) \in F[x], f(x) \in I$. Then $(hf)(\alpha) = h(\alpha) f(\alpha) = h(\alpha) 0 = 0 \implies (hf)(x) \in I$. 
    \end{enumerate}
  \end{proof}

  Since $F$ is a field, $F[x]$ is a PID, and so it must be generated by some element. The scaling of the coefficients doesn't really matter (since we are working in a field so we can always divide the leading coefficient), so we can assume that it is monic. This is called the \textit{minimal polynomial}. 

  \begin{definition}[Minimal Polynomial]
    \label{def:min_poly}
    Let $F \subset K$ be a field extension and $\alpha \in K$. The \textbf{minimal polynomial} of $\alpha$ is defined in the equivalent ways. 
    \begin{enumerate}
      \item It is the monic polynomial of least degree among all polynomials in $F[x]$ having $\alpha$ as a root. 
      \item It is the generator of the ideal of all polynomials in $F[x]$ with root $\alpha$. 
    \end{enumerate}
    We claim that it always exists and is unique. 
  \end{definition} 
  \begin{proof}
    Take the evaluation homomorphism $\ev_\alpha: F[x] \to K$ and look at the ideal $\ker{\ev_\alpha}$. Since $F[x]$ is a PID, call the generator $f(x)$ and normalize the leading coefficient to $1$. We claim that this is the minimal polynomial. This proves existence. 

    TBD. 
  \end{proof}

  \begin{example}[Minimal Polynomials]
    Let $\mathbb{Q} \subset \mathbb{R}$ be a field extension. 
    \begin{enumerate}
      \item The minimal polynomial for $\alpha = \sqrt{2} \in \mathbb{R}$ in $\mathbb{Q}[x]$ is 
      \begin{equation}
        x^2 - 2 = (x + \sqrt{2}) (x - \sqrt{2})
      \end{equation}

      \item The minimal polynomial for $\alpha = \sqrt{2} \in \mathbb{R}$ in $\mathbb{R}[x]$ is 
      \begin{equation}
        x - \sqrt{2}
      \end{equation}

      \item The minimal polynomial for $\alpha = \sqrt{2} + \sqrt{3}$ in $\mathbb{Q}[x]$ is 
      \begin{equation}
        x^4 - 10x^2 + 1 = (x - (\sqrt{2} + \sqrt{3}))(x - (\sqrt{2} - \sqrt{3}))(x - (-\sqrt{2} + \sqrt{3}))(x - (-\sqrt{2} - \sqrt{3}))
      \end{equation}
    \end{enumerate}
  \end{example}

  Note that the minimal polynomial is also irreducible. Let's prove this. 

  \begin{lemma}[Minimal Polynomial is Irreducible]
    Let $K/F$ be a field extension. The minimal polynomial $f(x) \in F[x]$ of $\alpha \in K$ is irreducible in $F[x]$. 
  \end{lemma}
  \begin{proof}
    Assume that $f(x)$ is reducible. Then we factor it into $f(x) = g(x) h(x)$, which implies that 
    \begin{equation}
      0 = f(\alpha) = g(\alpha) h(\alpha) \implies g(\alpha) = 0 \text{ or } h(\alpha) = 0
    \end{equation}
    since $F$ is a field and therefore does not contain zero divisors. Choosing both $g(x), h(x)$ to be of degree strictly lower than $f(x)$ would contradict the minimality requirement of $f(x)$, so $f(x)$ must be irreducible. 
  \end{proof}

  Since the minimal polynomial is irreducible, we can see that from Bezout's identity that the minimal polynomial generates a maximal ideal, and so the quotient ring is actually a field. 

  \begin{theorem}[Quotient Ring of Minimal Polynomial is a Field]
    Let $F$ be a field and $f(x) \in F[x]$, then 
    \begin{equation}
      f(x) \text{ is irreducible in } F[x] \iff \frac{F[x]}{\langle f(x) \rangle} \text{ is a field}
    \end{equation} 
  \end{theorem}
  \begin{proof}
    We first claim that $f(x)$ is irreducible if and only if $f(x)$ is a maximal ideal.  
    \begin{enumerate}
      \item If $f(x)$ is irreducible, then consider the ideal $I = \langle f(x) \rangle$ and add some $g(x) \in F[x] \setminus I$. Since $F[x]$ is a PID, $f(x)$ is prime, and so $\gcd(f, g) = 1$. From Bezout's identity, there must exist $a(x), b(x) \in F[x]$ s.t. $a(x) f(x) + b(x) g(x) = 1$, which implies that any larger ideal than $\langle f(x) \rangle$ must be the entire $F[x]$. 

      \item If $f(x)$ is reducible, then let $f(x) = g(x) h(x)$. Then $\langle f(x) \rangle \subsetneq \langle g(x) \rangle \subsetneq F[x]$, and so $\langle f(x) \rangle$ is not maximal. 
    \end{enumerate}
    Then with \ref{thm:quotient_ring_fields}, we know that $I$ is a maximal ideal if and only if $F[x]/I$ is a field, and the proof is complete. 
  \end{proof}

  Now we went through all this trouble of determining sufficient conditions for quotient rings to be fields, when our original goal was on adjoining rings. It seemed like a long detour, but this finally pays off, as now we have both a necessary and sufficient condition to reach our original goal. 

  \begin{corollary}[Conditions For Adjoining Ring to be a Field]
    Let $F \hookrightarrow K$ be a field extension and $\alpha \in K$. 
    \begin{enumerate}
      \item If $\alpha$ is algebraic over $F$---i.e. it is the root of some $g(x) \in F[x]$---then $F(\alpha) \subset K$ is a field. 
      \item The $F$-dimension of $K$ $[F(\alpha):F]$ is the degree of the minimal polynomial of $\alpha$.\footnote{It is clear that if there exists \textit{some} $f(x) \in F[x]$ that has root $\alpha$, then it may not be the \textit{unique} one. The dimension resides specifically in the unique minimal polynomial. } 
    \end{enumerate}
  \end{corollary}
  \begin{proof} 
    Let $f(x)$ be the minimal polynomial of $\alpha$ of degree $n$, and we know that $g(x) \in \langle f(x) \rangle \implies g(x) = h(x) f(x)$ for some $h(x) \in F[x]$. Take the surjective ring homomorphism $\ev_\alpha: F[x] \to F[\alpha]$. We know from \ref{def:min_poly} that $\ker(\ev_\alpha) = \langle f(x) \rangle$, and by the first isomorphism theorem of rings, we have 
    \begin{equation}
      \frac{F[x]}{\langle f(x) \rangle} = \frac{F[x]}{\ker(\ev_\alpha)} \simeq F[\alpha]
    \end{equation} 
    which proves the first claim. For dimension, we know that $\{1, \ldots, x^{n-1}\}$ is a basis. 
  \end{proof}
  \begin{proof}
    This alternative proof is constructive in that it actually shows how to compute multiplicative inverses. 

    It is clear that $F[\alpha]$ is a commutative ring since $F$ is a field. So it remains to show that every nonzero element of $\beta \in F[\alpha]$ is a unit. By definition $\beta = p(\alpha)$ for some polynomial $p \in F[x]$. Factor $f \in F[x]$ as the product of irreducible polynomials. Then $\alpha$ must be a root of one of those irreducible factors, say $g(x)$. Note that $g(x) \nmid p(x)$ since $p(\alpha) \neq 0$. Since $g$ is irreducible, we know that $\gcd(g, p) = 1$ and so $\exists s, t \in F[x]$ s.t. 
    \begin{equation}
      1 = s p + t g \implies 1 = s(\alpha) p(\alpha) + t(\alpha) g(\alpha) = s(\alpha) p(\alpha)
    \end{equation}  
    Therefore we have found a multiplicative inverse $s = p^{-1} \in F[\alpha]$. 
  \end{proof} 
  \begin{proof}
    We can also prove it using the vector space structure, though this is for finite-dimensional vector spaces. Treating $F[\alpha]$as a finite-dimensional vector space over $F$, let us define the $F$-linear function\footnote{linearity is easy to check}
    \begin{equation}
      m_b: F[\alpha] \rightarrow F[\alpha], \qquad m_b (\beta) = b\beta
    \end{equation} 
    Since $F[\alpha] \subset K$, $F[\alpha]$ is an integral domain. Thus $\not\exists \beta \in F[\alpha] \setminus \{0\}$ s.t. $b \beta = 0$. This means that the kernel of $m_b$ is $0$, and so $m_b$ is injective. By the rank-nullity theorem, it is bijective, and so there exists a $\beta \in F[\alpha]$ s.t. $b \beta = 1 \implies b$ is a unit. 
  \end{proof} 

  Since $F[\alpha]$ is the smallest ring containing both $F$ and $\alpha$, it immediately follows that it is the smallest \textit{field} containing $F$ and $\alpha$. Therefore, we have unified the two constructions of adjoining fields and quotient rings. 

  \begin{example}[Some Adjoining Fields]
    Here are some examples. 
    \begin{enumerate}
      \item Given any algebraic number $\alpha \in \mathbb{C}$, by definition there exists a $f(x) \in \mathbb{Q}[x]$ with roots $\alpha$, and so $\mathbb{Q}(\alpha) \subset \mathbb{C}$.\footnote{In fact it was historically so common to work solely in subfields of $\mathbb{C}$ that an \textit{algebraic number} meant by default algebraic with respec to the field extension $\mathbb{C}/F$.}
      \item $\mathbb{Q}(\sqrt{3} i)$ is a field of dimension 2 since $\sqrt{3}i$ is a root of the polynomial $f(x) = x^2 + 3$. 
      \item However, $\mathbb{Q}[\pi]$ is not a field since $\pi$ is a \textit{transcendental number}. However we will not prove it now. 
    \end{enumerate}
  \end{example}

  Note that we now have the field extension tower $F \subset F[\alpha] \subset K$. By the tower property, $[K:F] = [F:F(\alpha)] [F(\alpha):F]$, and so $\deg{f(x)} = [F:F(\alpha)]$ must divide $[K:F]$. 

  Now recall quotient rings, which do not necessarily preserve the properties of the original ring. That is, if $F$ is a field, then $F/I$ may not be a field. Using the fundamental ring homomorphism theorem, we can precisely correlate certain quotient maps with adjoining fields. Recall that given a field extension $F \subset K$, the evaluation function $\ev_\alpha: F[x] \rightarrow K$ defined $f(x) \mapsto f(\alpha)$ is a homomorphism. 

  \begin{example}[Simple Quotient Rings as Field Adjoined with 1 Variable]
    Consider the following. 
    \begin{enumerate}
      \item Since $x^2 + 1 \in \mathbb{Z}_7 [x]$ is irreducible, $\mathbb{Z}_7 [x] / \langle x^2 + 1 \rangle$ is a field. 

      \item Since $x^2 + 1 \in \mathbb{R}[x]$ is irreducible of degree 2, so the quotient ring is a field. Furthermore, $i \in \mathbb{C}$ is a root of the degree 2 minimal polynomial, so we have the isomorphism 
      \begin{equation}
        \frac{\mathbb{R}[x]}{\ker(\ev_i)} = \frac{\mathbb{R}[x]}{\langle x^2 + 1 \rangle} \simeq \mathbb{R}[i] = \mathbb{C}
      \end{equation}
      induced by the evaluation map
      \begin{equation}
        \phi: \frac{\mathbb{R}[x]}{\langle x^2 + 1 \rangle} \rightarrow \mathbb{C}, \qquad \phi\big( f(x) \pmod{\langle x^2 + 1 \rangle} \big) = f(i)
      \end{equation}
      Therefore $\mathbb{R}(i)$ has $\mathbb{R}$-dimension 2, which makes sense since we view $\mathbb{C}$ as being isormophic to $\mathbb{R}^2$. 

      \item The evaluation map 
      \begin{equation}
        \ev_{\sqrt{2}}: \mathbb{Q}[x] \mapsto \mathbb{Q}[\sqrt{2}], \qquad \ev_{\sqrt{2}} (f) = f(\sqrt{2}) 
      \end{equation}
      is a homomorphism. Furthermore, it has a kernel $\langle x^2 - 2 \rangle$ since $(x^2 - 2)$ is an irreducible polynomial in $\mathbb{Q}[x]$ containing the root $\sqrt{2}$. Therefore by the fundamental ring homomorphism theorem we have 
      \begin{equation}
        \frac{\mathbb{Q}[x]}{\langle x^2 - 2 \rangle} \simeq \mathbb{Q}[\sqrt{2}]
      \end{equation}
    \end{enumerate}
  \end{example}

  \begin{example}[Extensions of $\sqrt{2}$ and $i$]
    We claim that 
    \begin{equation}
      \mathbb{Q}[\sqrt{2}, i] = \{ a + b \sqrt{2} + ci + d(\sqrt{2} i) \mid a, b, c, d \in \mathbb{Q}\}
    \end{equation}
    From the previous example, we know that $\mathbb{Q}[\sqrt{2}]$ are all numbers of the form $a + b\sqrt{2}$. Now we take $i \in \mathbb{C}$ and map it through all polynomials with coefficients in $\mathbb{Z}[\sqrt{2}]$, which will be of form 
    \begin{equation}
      f(i) = (a_n + b_n \sqrt{2}) i^n + (a_{n-1} + b_{n-1}\sqrt{2}) i^{n-1} + \ldots + (a_2 + b_2 \sqrt{2}) i^2 + (a_1 + b_1 \sqrt{2}) i + (a_0 + b_0 \sqrt{2})
    \end{equation} 
    However, we can see that since $i^2 = -1$, we only need to consider up to degree 1 polynomials of form 
    \begin{equation}
      (a + b \sqrt{2}) + (c + d \sqrt{2}) i 
    \end{equation}
    which is clearly of the desired form. For the other way around, this is trivial since we can construct a linear polynomial as before. 
  \end{example} 

  \begin{example}
    We claim $\mathbb{Q}[\sqrt{3} + i] = \mathbb{Q}[\sqrt{3}, i]$. 
    \begin{enumerate}
      \item $\mathbb{Q}[\sqrt{3} + i] \subset \mathbb{Q}[\sqrt{3}, i]$
      \item $\mathbb{Q}[\sqrt{3} + i] \supset \mathbb{Q}[\sqrt{3}, i]$. Note that 
        \begin{align}
          (\sqrt{3} + i)^3 = 8i & \implies i \in \mathbb{Q}[\sqrt{3} + i] \\
                                & \implies (\sqrt{3} + i) - i = \sqrt{3} \in \mathbb{Q}[\sqrt{3} + i] 
        \end{align}
        Therefore, $\mathbb{Q}[\sqrt{3} + i]$ contains the elements $1, \sqrt{3}, i$, which form the basis of $\mathbb{Q}[\sqrt{3}, i]$. 
    \end{enumerate}
  \end{example}

  \begin{example}[Extensions of $\sqrt{3}i$ and $\sqrt{3}, i$]
    We claim that $\mathbb{Q}[\sqrt{3} i] \subsetneq \mathbb{Q}[\sqrt{3}, i]$. 
    \begin{enumerate}
      \item We can see that $\{1, \sqrt{3}i \}$ span $\mathbb{Q}[\sqrt{3}i ]$ as a $\mathbb{Q}$-vector space. Therefore, 
      \begin{equation}
        \sqrt{3}, i \in \mathbb{Q}[\sqrt{3}, i] \implies \sqrt{3} i \in \mathbb{Q}[\sqrt{3}, i]
      \end{equation} 
      implies that $\mathbb{Q}[\sqrt{3} i] \subset \mathbb{Q}[\sqrt{3}, i]$. 

      \item To prove proper inclusion, we claim that $i \not\in \mathbb{Q}[\sqrt{3}i]$. Assuming that it can, we represent it in the basis $i = b_0 + b_1 \sqrt{3} i$, and so
      \begin{equation}
        -1 = (b_0 + b_1 \sqrt{3} i)^2 = (b_0^2 - 3b_1^2) + 2b_0 b_1 \sqrt{3} i
      \end{equation}
      Therefore we must have $2b_0 b_1 \sqrt{3} = 0 \implies b_0$ or $b_1$ should be $0$. If $b_0 = 0$, then $b_0^2 - 3b_1^2 = -3 b_1^2 \implies b_1^2 = 1/3$, which is not possible since $b_1^2 \in \mathbb{Q}$. If $b_1 = 0$, then $b_0 - 3 b_1^2 = b_0^2 > 0$, and so it cannot be $-1$. 
    \end{enumerate}
  \end{example}

  The most significant property is that a field contains inverses. So how do we compute such an inverse? 

  \begin{example}[Computing Inverses in Field Extensions]
    We can do two problems. 
    \begin{enumerate}
      \item Given $\beta = p(\alpha) = \alpha^2 + \alpha - 1 \in \mathbb{Q}[\alpha]$, where $\alpha$ is a root of $f(\alpha) = \alpha^3 + \alpha + 1$, we first know that $\beta$ must have a multiplicative inverse since $\mathbb{Q}[\alpha]$ is a field. Applying the Euclidean algorithm, we have 
      \begin{equation}
        1 = \frac{1}{3} \big\{ (x+1) f(x) - (x^2 + 2) p(x)\big\} = -\frac{1}{3} (\alpha^2 + 2) p(\alpha)
      \end{equation}
      and so $\beta^{-1} = (\alpha^2 + \alpha - 1)^{-1} = -\frac{1}{3} (\alpha^2 + 2)$. We can check that 
      \begin{align}
        -\frac{1}{3} (\alpha^2 + 2) (\alpha^2 + \alpha - 1) & = -\frac{1}{3} (\alpha^4 + \alpha^3 + \alpha^2 + 2 \alpha - 2) \\
                                                            & = -\frac{1}{3} (\alpha^3 + \alpha - 2) \\
                                                            & = -\frac{1}{3} (-3) = 1
      \end{align}
      \item Given $f(x) = x^3 + 3 \in \mathbb{Q}[x]$, we know (from Eisenstein) that it is irreducible. Therefore we know that $\mathbb{Q}[x]/\langle f(x) \rangle$ is a field. Furthermore, it contains a root $\alpha$, and so 
      \begin{equation}
        \frac{F[x]}{\langle x^3 + x \rangle} \simeq F[\alpha]
      \end{equation}
      Now take an arbitrary element of the field, say $f(\alpha) = \alpha^2 - 1$. Then it must have an inverse call it $g(\alpha)$ that satisfies 
      \begin{equation}
        1 = f(\alpha) g(\alpha) \text{ in } \frac{\mathbb{Q}[x]}{\langle x^3 + 3 \rangle}
      \end{equation}
      But by working in the original ring $\mathbb{Q}[x]$ and not the quotient, this would look something like. 
      \begin{equation}
        1 = f(x) g(x) + k(x) (x^3 + 1) \text{ in } \mathbb{Q}[x]
      \end{equation} 
      for some $k(\alpha)$. Note that we are guaranteed to get this form since $f(x)$ is irreducible and so $\gcd(f, g) = 1$. So we use Euclidean division. 
      \begin{align}
        x^3 + 3 & = (x) (x^2 - 1) + (x + 3) \label{hi}\\
        x^2 - 1 & = (x - 3) (x + 3) + 8
      \end{align}
      Therefore by substituting we have 
      \begin{equation}
        8 = (x - 3) (x^3 + 3) - (x^2 - 2x - 1) (x^2 - 1) 
      \end{equation} 
      and so by taking $\pmod{x^3 + 3}$, we have 
      \begin{equation}
        1 = -\frac{1}{8} (x^2 - 3x - 1) (x^2 - 1) \implies \frac{1}{\alpha^2 - 1} = -\frac{1}{8}(\alpha^2 - 3 \alpha - 1)
      \end{equation}
    \end{enumerate}
  \end{example}

  \begin{example}[Irreducible Quotient Rings]
    On the other hand, if we consider the ring $\mathbb{Z}_5 [x] / \langle x^2 - 2 \rangle$, then $x^2 - 2$ is irreducible in $\mathbb{Z}_5 [x]$ (just plug in $0, 1, 2, 3, 4$), and so since $\mathbb{Z}_5$ is a field, the quotient ring is a field. Then we have 
    \begin{equation}
      \frac{\mathbb{Z}_5 [x]}{\langle x^2 - 2 \rangle} \simeq \mathbb{Z}_5 [\sqrt{2}] = \{a + b \sqrt{2} \mid a, b \in \mathbb{Z}_5 \}
    \end{equation}
  \end{example} 

  Remember that we should not be careless and just mindlessly assume that every quotient ring is a field. This is only when the ideal is generated by a minimal polynomial. Here is an example where the ideal is reducible, but fortunately using the chinese remainder theorem, we can first reduce it as a direct product of quotient rings, and then apply our theorem to represent them as a field structure. 

  \begin{example}[Quotient Ring as Product Rings with Chinese Remainder Theorem]
    Consider the following ring $\mathbb{Z}_5 [x] / \langle x^2 + 1 \rangle$. Then $x^2 + 1$ is reducible since $x^2 + 1 = (x + 2) (x + 3)$, which implies $\langle x^2 + 1 \rangle = \langle x + 2 \rangle \cap \langle x + 3 \rangle$, and so by the Chinese remainder theorem, we have 
    \begin{equation}
      \frac{\mathbb{Z}_5 [x]}{\langle x^2 + 1 \rangle} = \frac{\mathbb{Z}_5 [x]}{\langle x + 2 \rangle \cap \langle x + 3 \rangle} = \frac{\mathbb{Z}_5 [x]}{\langle x + 2 \rangle} \times \frac{\mathbb{Z}_5 [x]}{\langle x + 3 \rangle} \simeq \mathbb{Z}_5 \times \mathbb{Z}_5
    \end{equation}
  \end{example} 

\subsection{Splitting Fields} 

  Great, so we have established conditions for which a field adjoined with an element becomes another field. All that is left to do is to find a splitting field.

  Remember that by previously establishing that $\mathbb{C}$ is algebraically closed, this gives us a ``safe space'' to work in, in the sense that if we take any subfield $F \subset \mathbb{C}$ and find a polynomial $f(x) \in F[x]$, we are \textit{guaranteed} to find a linear factorization of $f$ in $\mathbb{C}[x]$. Therefore, if $K$ is algebraically closed and $F \subset K$ is a field extension, $f(x) \in F[x]$ is guaranteed to \textit{split} completely into linear factors. This is true for \textit{all} $f(x) \in F[x]$, but now if we \textit{fix} $f(x) \in F[x]$, perhaps we don't need the entire field $K$ to split $f(x)$. Maybe we can work in a slightly larger field $E$---such that $F \subset E \subset K$---where $f(x)$ splits in $E$. This process of finding such a minimal field is important to understand the behavior of roots of such polynomials. 

  \begin{definition}[Splitting Field]
    Given a field extension $F \subset K$ and a polynomial $f \in F[x]$, 
    \begin{enumerate}
      \item $f$ \textbf{splits} in $K$ if $f$ can be written as the product of linear polynomials in $K[x]$. 
      \item If $f$ splits in $K$ and there exists no field $E$ s.t. $F \subsetneq E \subsetneq K$, then $K$ is called a \textbf{splitting field} of $f$.\footnote{i.e. the splitting field is the ``smallest'' field that splits $f$.} 
    \end{enumerate}
    We claim that a splitting field always exists for any $f(x) \in F[x]$ and is unique. 
  \end{definition}
  \begin{proof}
    Let us decompose $f(x)$ into irreducible factors over $F$: $f(x) = f_1(x)f_2(x)\cdots f_k(x)$. If all of them are linear, then $F$ is the splitting field for $f(x)$. Otherwise, we can assume that $f_1(x)$ is not linear. Then we can consider the field $K_1 = F[x]/\langle f_1(x) \rangle$, where $f_1(x)$ has a root $\alpha$. Then $f_1(x) = (x - \alpha)g_1(x)$, and $f(x)$ has at least one (but maybe more) linear factor over $K_1$. If all irreducible factors over $K_1$ are linear, stop, otherwise there is an irreducible factor of degree at least 2, and we can repeat the procedure and add its root. Since a polynomial of degree $n$ has at most $n$ roots, the process will eventually stop and all factors will be linear in some extension of $F$.
  \end{proof} 

  \begin{example}[Don't Need Necessarily Complex Numbers to Split]
    Consider the following subfields of $\mathbb{C}$ and observe that they are enough to split a given polynomial. 
    \begin{enumerate}
      \item Let $f(x) = x^2 - 1$. If $f(x) \in \mathbb{R}[x]$, it does split in $\mathbb{R}$. In fact, even if we consider it as an element of $\mathbb{Z}_2 [x]$, it still splits into $(x + 1)(x - 1)$. 
      \item Let $f(x) = x^2 - 2$. If $f(x) \in \mathbb{Q}[x]$, it doesn't split in $\mathbb{Q}$ since the roots $\pm \sqrt{2} \not\in \mathbb{Q}$, but $\pm \sqrt{2}$ are real numbers, so $f(x)$ does in fact split in $\mathbb{R}$ since it splits into $(x + \sqrt{2}) (x - \sqrt{2})$. However, maybe it is not the (smallest) splitting field. 
      \item Let $f(x) = x^2 + 1$. We can see that if we consider it as an element of $\mathbb{Q}[x]$ or $\mathbb{R}[x]$, neither fields split $f(x)$ since $\pm i$ are its roots and therefore are contained in the coefficients of its linear factors. We know that it definitely splits in $\mathbb{C}$, but can we find a smaller field that splits $f(x)$? Perhaps.  
    \end{enumerate}
  \end{example}

  So how does one find a splitting field? Note that in the example above, we have found that there were some roots $\alpha$ of certain polynomials $f(x) \in F[x]$ are not contained in $F$. Therefore, what we want to do is find the smallest field $F$ containing both $F$ and $\alpha$ (plus any other $\alpha$'s). This is precisely the adjoining field $F[\alpha]$, which guarantees unique factorization since $F[\alpha]$ is a Euclidean domain. 

  \begin{lemma}[Square-Free Extensions as Splitting Field]
    If $a$ is not a perfect square in $F$ then $F(\sqrt{a})$ is the splitting field of $f(x) = x^2 - a$. 
  \end{lemma}
  \begin{proof}
    TBD
  \end{proof}

  \begin{example}[Simple Splitting Fields]
    We provide some simple examples to gain intuition. 
    \begin{enumerate}
      \item Let $f(x) = x^2 + 2x + 2 \in \mathbb{Q}[x]$. Then the roots of $f(x)$ are $-1 \pm i$, so 
      \begin{equation}
        f(x) = (x - (-1 + i)) (x - (-1 - i)) 
      \end{equation}
      and we can show that $\mathbb{Q}[-1 - i, -1+i] = \mathbb{Q}[i]$ is the splitting field of $f$. It has dimension $2$ since $f(x)$ is a 2nd degree polynomial that is minimal. 

      \item Let $f(x) = x^2 - 2x - 1 \in \mathbb{Q}[x]$. The roots are $1 \pm \sqrt{2}$, and so 
      \begin{equation}
        f(x) = (x - (1 + \sqrt{2})) (x - (1 - \sqrt{2}))
      \end{equation}
      and so $\mathbb{Q}[\sqrt{2}]$ is the splitting field of $f$. Since $f(x)$ is a 2nd degree polynomial that is minimal. T

      \item Let $f(x) = x^6 - 1 \in \mathbb{Q}[x]$. We can factor 
      \begin{equation}
        f(x) = (x-1) (x + 1) (x^2 + x + 1) (x^2 - x + 1)
      \end{equation} 
      and the non-rational roots are $\frac{\pm 1 \pm \sqrt{3} i}{2}$. Thus the splitting field of $f$ is $\mathbb{Q}[\sqrt{3} i]$. The dimension is not 6 because $f(x)$ is reducible over $\mathbb{Q}$. It is easier to look at this not as a quotient ring, but use the first theorem to find a minimal polynomial with root $\sqrt{3} i$. Indeed, such a polynomial is $x^2 + 3$, which has degree $2$. 
    \end{enumerate}
  \end{example}

  \begin{example}
    Let $f(x) = x^4 - 2 \in \mathbb{Q}[x]$. It follows that the roots are 
    \begin{equation}
      \{ \sqrt[4]{2}, \sqrt[4]{2}, -\sqrt[4]{2}, - \sqrt[4]{2} i \} = \Big\{ \sqrt[4]{2}, \sqrt[4]{2} e^{\frac{2\pi i}{4}}, \sqrt[4]{2} e^{\frac{4\pi i}{4}}, \sqrt[4]{2} e^{\frac{6\pi i}{4}} \Big\}
    \end{equation}
    thus the splitting field of $f$ is 
    \begin{equation}
      \mathbb{Q} \big( \sqrt[4]{2}, \sqrt[4]{2} e^{\frac{2\pi i}{4}}, \sqrt[4]{2} e^{\frac{4\pi i}{4}}, \sqrt[4]{2} e^{\frac{6\pi i}{4}} \big) \subset \mathbb{Q}(\sqrt[4]{2}, e^{\frac{2\pi i}{4}})
    \end{equation}
    since $\sqrt[4]{2} e^{\frac{m \pi i}{4}} \in \mathbb{Q}(\sqrt[4]{2}, e^{\frac{2\pi i}{4}})$. In fact, the two are equal, and to prove this we can see that since we are working in a field, 
    \begin{equation}
      e^{2 \pi i / 4} = \frac{\sqrt[4]{2} e^{2\pi i/4}}{\sqrt[4]{2}} \in \mathbb{Q} \big( \sqrt[4]{2}, \sqrt[4]{2} e^{\frac{2\pi i}{4}}, \sqrt[4]{2} e^{\frac{4\pi i}{4}}, \sqrt[4]{2} e^{\frac{6\pi i}{4}} \big) 
    \end{equation}
    which implies that $\sqrt[4]{2} \in \mathbb{Q} \big( \sqrt[4]{2}, \sqrt[4]{2} e^{\frac{2\pi i}{4}}, \sqrt[4]{2} e^{\frac{4\pi i}{4}}, \sqrt[4]{2} e^{\frac{6\pi i}{4}} \big)$. Therefore we can conclude that the splitting field is 
    \begin{equation}
      \mathbb{Q} \big( \sqrt[4]{2}, \sqrt[4]{2} e^{\frac{2\pi i}{4}}, \sqrt[4]{2} e^{\frac{4\pi i}{4}}, \sqrt[4]{2} e^{\frac{6\pi i}{4}} \big) = \mathbb{Q}(\sqrt[4]{2}, e^{\frac{2\pi i}{4}})
    \end{equation}
    $f(x)$ is certainly minimal in $\mathbb{Q}[x]$ containing all roots, so the dimension of the field is $4$. It is spanned by the powers of $\sqrt[4]{2}$. 
  \end{example} 

  \begin{example} 
    Let $p$ be prime, consider the polynomial $f(x) = x^p - 1$ over $\mathbb{Q}$. Let $\omega = e^{2\pi i/p}$, then $\omega$ is a root of $f(x)$ and we have
    \begin{equation}
      x^p - 1 = (x - 1)(x - \omega)(x - \omega^2)\cdots(x - \omega^{p-1}),
    \end{equation}
    so all roots of $f(x)$ belong to the extension $\mathbb{Q}(\omega)$. Since the minimal polynomial for $\omega$ equals $x^{p-1} + \ldots + 1$, we have $[\mathbb{Q}(\omega) : \mathbb{Q}] = p - 1$.
  \end{example}

  With this splitting condition, we can get a nice set of formulas often introduced in high-school math competitions. 

  \begin{theorem}[Viete's Formulas]
    Given that a polynomial $f(x)$ splits in $F[x]$, i.e.
    \begin{equation}
      f(x) = a_0 \prod_{i = 1}^{n} (x - \alpha_i)
    \end{equation}
    Then the coefficients of $f$ can be presented with the formulas
    \begin{align}
      \sum_{i=1}^n \alpha_i & = - \frac{a_1}{a_0} \\
      \sum_{i_1 < i_2} \alpha_{i_1} \alpha_{i_2} & = \frac{a_2}{a_0} \\
      \sum_{i_1< ...< i_k} \prod_{j = 1}^{k} \alpha_{i_j} & = (-1)^k \frac{a_k}{a_0} \\
      \alpha_1 \alpha_2 \alpha_3 ... \alpha_n & = (-1)^n \frac{a_n}{a_0}
    \end{align}
  \end{theorem}
  \begin{proof}
    
  \end{proof}

  \begin{example}[Multivariate Splitting Fields]
    $x^3 - 2 \in \mathbb{Q}[x]$ has a splitting field. By Eisenstein, $x^3 - 2$ is irreducible, and so 
    \begin{equation}
      F = \frac{\mathbb{Q}[x]}{\langle x^3 - 2 \rangle} 
    \end{equation}
    is a field. Furthermore, there must be a root $\alpha \in F$. We factor this in the field $F$ to get 
    \begin{equation}
      x^3 - 2 = (x - \alpha) (x^2 + \alpha x + \alpha^2) 
    \end{equation}
    Then either $x^2 + \alpha x + \alpha^2$ has a root in $F$ and $F$ is the splitting field, or it is irreducible. It turns out that $x^2 + \alpha x + \alpha^2$ is irreducible in $F$, and so the splitting field is 
    \begin{equation}
      E = \frac{F[y]}{\langle y^2 + \alpha y + \alpha^2 \rangle} \simeq \frac{\mathbb{Q}[x, y]}{\langle x^3 - 2, y^2 + xy + x^2 \rangle}
    \end{equation}
  \end{example}

\subsection{Finite Fields} 

  Before we talk about Galois groups, we detour here to find a classification of \textit{all} finite fields, which is quite a powerful result. We have all the tools needed for this. We know that a field---as an integral domain---has characteristic $0$ or prime $p$. We also know that a field is a vector space, at least over itself. But now that we have shown that a field can be modeled as a vector space, we can apply this to finite fields. 

  \begin{theorem}[Characteristic Determines Base Field of Vector Space]
    \label{thm:char_field}
    Given a field $F$, 
    \begin{enumerate}
      \item If $\Char(F) = p$, then $F$ is a vector space over $\mathbb{Z}_p$. 
      \item If $\Char(F) = 0$, then $F$ is a vector space over $\mathbb{Q}$. 
    \end{enumerate}
  \end{theorem}
  \begin{proof}
    
  \end{proof} 

  Therefore, just from the characteristic we can classify all fields as vector spaces over either $\mathbb{Q}$ or $\mathbb{Z}_p$. Now if we focus on finite fields, we can do a reverse classification. 

  \begin{theorem}[Finite Fields Have Cardinality $p^d$]
    Let $F$ be a finite field. Then $|F| = p^n$ for some $n \in \mathbb{N}$. 
  \end{theorem}
  \begin{proof}
    $F$ is a vector space over $\mathbb{Z}_p$ from \ref{thm:char_field}. Since $F$ has finitely many elements, $F$ has a finite spanning set, which implies $\dim_{\mathbb{Z}_p} F \leq + \infty$. Let $d$ be the dimension and $\{b_1, \ldots, b_d\}$ be the basis. The elements of $F$ are 
    \begin{equation}
      a_1 b_1 + \ldots + a_d b_d
    \end{equation}
    with $a_1, \ldots a_d \in \mathbb{Z}_p$. Thus there are $p^d$ elements of $F$, so $F \simeq \mathbb{Z}_p^d$. 
  \end{proof}

  In fact, for \textit{every} prime power there exists a unique field. Therefore we can create a bijection by proving the converse. 

  \begin{theorem}[Field for Every $p^d$]
    For every prime $p$ and $n \in \mathbb{N}$, there exists a field with $q = p^d$ elements, unique up to isomorphism.  
  \end{theorem} 
  \begin{proof} 
    Let $f(x) = x^q - x \in \mathbb{Z}_p [x]$. Then this polynomial has a splitting field $K \supset \mathbb{Z}$. Now we claim the roots of $f(x)$ in $K$ are distinct and form a subfield $F_q \subset K$. This will complete the proof since $F_q \subset K$ and $K \subset F_q \implies K = F_q$. Assume $\alpha, \beta \in K$ are roots of $f(x)$, and so $\alpha^p = \alpha$ and $\beta^p = \beta$
    \begin{enumerate}
      \item $\alpha + \beta \in K$ since by a modification of Freshman's dream, $(\alpha + \beta)^p = \alpha^p + \beta^p = \alpha + \beta$.\footnote{We induct on $n$ for $q = p^n$. For $n=1$, this is trivial by Freshmans dream. Now assume it holds for some $n \in \mathbb{N}$. Then $(x + y)^{p^{n+1}} = ( (x + y)^{p^n} )^p = (x^{p^n} + y^{p^n})^p = (x^{p^n})^p + (y^{p^n})^p = x^{p^{n+1}} + y^{p^{n+1}}$. }
      \item $(-\alpha)^q = (-1)^q \alpha^q = (-1)^q \alpha = -\alpha$ since $-1 = 1$ or $q$ is odd. 
      \item $\alpha \beta \in K$ since $\mathbb{Z}_p$ is a field and so $(\alpha \beta)^p = \alpha^p \beta^p = \alpha \beta$. 
      \item For multiplicative inverses, let $\alpha \neq 0$. Then $(\alpha^{-1})^p = (\alpha^{p})^{-1} = \alpha^{-1}$. 
      \item For all $p$, $0$ and $1$ are roots so $0, 1 \in K$. 
    \end{enumerate}
    Now we show that $K$ consists of distinct roots. Certainly $0 \in K$ with multiplicity $1$ since $f(x) = x (x^{q-1} - x)$. Now suppose nonzero $r \in K$ is a root with multiplicity $m$. The multiplicity of $r$ is the multiplicity of $0$ of 
    \begin{equation}
      f(x + r) = (x + r)^q - (x + r) = x^q + r^q - x - r = x^q - x
    \end{equation}
    where the final step follows from $0 = r^q - r$ since $r \in K$. Therefore $r$ has multiplicity $1$. Since $K[x]$ has unique factorization property, it follows that $m=1$ and every $r$ is a simple root. 

    To show that every field with $p^n$ elements is unique, let $F$ be such a field. We claim that $\Char(F) = p \implies \mathbb{Z}_p \subset F$. We claim that every element of $F$ is a root of $f(x) = x^q - x \in \mathbb{Z}_p [x]$, where $F$ is the splitting field. Let $G = F^\ast$ be the multiplicative group of units. Since $F$ is a field, then $|F^\ast| = |F| - 1 = p^d - 1$, and by constructing the cyclic group $\langle g \rangle \subset G$ for any $g \in G$, we know by Lagrange's theorem that $g^{|G|} = 1_G$, which implies that for all $x \in F$, 
    \begin{enumerate}
      \item If $x \neq 0$ then $x^{p^d - 1} = x \implies x^{p^d} = x$ and so $x \in K$. 
      \item If $x = 0$ then $x^{p^d} - x = 0$ and so $x \in K$. 
    \end{enumerate}
    Therefore $F \subset K$ with $|F| = |K|$ both finite, and so $F = K$. 
  \end{proof} 

  From this, we can write for every prime $p$ and natural $n$ the finite field of order $p^n$ as $\mathbb{F}_{p^n}$. It is clear that if $n = 1$ then $\mathbb{F}_p \simeq \mathbb{Z}_p$. The final result we will show is a hierarchy of subfields. 

  \begin{theorem}[Hierarchy of Fields]
    For a given prime $p$, if $p^m < p^n$, then 
    \begin{equation}
      F_{q^m} \subset F_{q^n} \iff m \mid n
    \end{equation}
  \end{theorem}

\subsection{Automorphism Groups} 

  Now we delve into the heart of Galois theory. Once we know that field extensions are vector spaces, what consistutes linear maps? A first guess would be a ring homomorphism, but this may not be true. Therefore, we need some additional constraint. 

  \begin{definition}[Field Embedding]
    Let $F \subset K_1, K_2$ be two field extensions. An \textbf{$F$-embedding} is a ring homomorphism $\phi: K_1 \to K_2$ such that $\phi(a) = a$ for all $a \in F$.An \textbf{$F$-automorphism of $K$} is an $F$-embedding $\phi: K \to K$. 
  \end{definition}

  \begin{lemma}[$F$-Embeddings as Linear Maps]
    Given $F \subset K_1, K_2$, an $F$-embedding $\phi: K_1 \to K_2$ is a linear map of $F$-vector spaces. 
  \end{lemma}
  \begin{proof}
    
  \end{proof}

  \begin{lemma}[Bijection Between Embeddings and Roots]
    Let $F \subset K_1, K_2$ be two field extensions and $K_1 \simeq F[x] / \langle f(x) \rangle$ for some irreducible polynomial $f(x) \in F[x]$. Then there exists a bijection between 
    \begin{enumerate}
      \item The $F$-embeddings $\phi: K_1 \to K_2$. 
      \item The roots $r \in K_2$ s.t. $f(r) = 0$. 
    \end{enumerate}
    In other words, for every root $r$ of $f(x)$ in $K_2$, there exists a unique $F$-embedding. 
  \end{lemma}

  \begin{definition}[$F$-Automorphism Group of $K$]
    Given field extension $K/F$, the group of all $F$-automorphisms of $K$ with composition is called the \textbf{$F$-automorphism group of $K$}, denoted $\Aut(K/F)$.\footnote{Often, this is introduced as the Galois group, but technically we need some extra conditions. To minimize confusion, I will refer to this as the automorphism group.}
  \end{definition}
  \begin{proof}
    This is indeed a group under composition. The identity map $\iota \in \Aut(K/F)$. The composition is clearly closed. Now given that $\phi \in \Aut(K/F)$, $\phi^{-1}$ is also an automorphism that is constant on $F$, so it is also in $\Aut(K/F)$. 
  \end{proof} 

  Essentially, the $F$-automorphism group is a subgroup of the ring automorphism group of $K$ that doesn't vary $F \subset K$. It is essentially a transformation group. Let's provide a few examples to derive some of the automorphism groups. 

  \begin{example}[Computing $\Aut(\mathbb{Q}(\sqrt{2})/\mathbb{Q})$]
    Given the field extension $\mathbb{Q} \subset \mathbb{Q}(\sqrt{2})$, we have for any $a, b \in \mathbb{Q}$ and $\phi \in \Aut(\mathbb{Q}(\sqrt{2})/\mathbb{Q})$, we have 
    \begin{equation}
      \phi(a + b \sqrt{2}) = \phi(a) + \phi(b \sqrt{2}) = a + b \phi(\sqrt{2}) 
    \end{equation}
    So $\phi$ is completely determined by the value of $\phi(\sqrt{2})$. Now let $\phi(\sqrt{2}) = \alpha + \beta \sqrt{2}$ for some $\alpha, \beta \in \mathbb{Q}$. We have 
    \begin{align}
      2 = \phi(2) = \phi(\sqrt{2} \sqrt{2}) = \phi(\sqrt{2})^2 = (\alpha + \beta\sqrt{2})^2 = (\alpha^2 + 2 \beta^2) + 2 \alpha \beta \sqrt{2} 
    \end{align}
    Therefore $\alpha \beta = 0$ and $\alpha^2 + 2 \beta^2 = 2$. With further casework, we must have $\alpha = 0, \beta = \pm 1$. In conclusion, there are exactly two $\mathbb{Q}$-automorphisms of $\mathbb{Q}(\sqrt{2})$. 
    \begin{enumerate}
      \item The identity map $\iota(a + b \sqrt{2}) = a + b \sqrt{2}$, and 
      \item The conjugation map $\phi(a + b \sqrt{2}) = a - b \sqrt{2}$. 
    \end{enumerate}
  \end{example}

  \begin{example}[Computing $\Aut(\mathbb{Q}(2^{1/3})/\mathbb{Q})$]
    Given the field extension $\mathbb{Q} \subset \mathbb{Q}(\sqrt[3]{2})$, let us write $\xi = \sqrt[3]{2}$. Then since $\mathbb{Q}(\xi)$ is a vector space with basis $1, \xi, \xi^2$, we can write any element as $a + b \xi + c \xi^2$, and so by definition an element $\phi \in \Aut(\mathbb{Q}(\xi)/\mathbb{Q})$ must satisfy 
    \begin{equation}
      \phi(a + b \xi + c \xi^2) = a + b \phi(\xi) + c \phi(\xi)^2
    \end{equation}
    So the action is completely determined by the value of $\phi(\xi)$. Suppose $\phi(\xi) = \alpha + \beta \xi + \gamma \xi^2$ for some $\alpha, \beta, \gamma \in \mathbb{Q}$. Through some derivation we have 
    \begin{align}
      2 & = \phi(2) = \phi(\xi^3) = (\phi(\xi))^3 = (\alpha + \beta \xi + \gamma \xi^2)^3 \\ 
        & = (\alpha^3 + 2\beta^3 + 4\gamma^3) + 3(\alpha^2\beta + \beta^2\gamma + \gamma^2\alpha)\xi + 3(\alpha\beta^2 + \beta\gamma^2 + \gamma\alpha^2)\xi^2
    \end{align} 
    From the linear independence of $1, \xi, \xi^2$ we can see that 
    \begin{align}
      \alpha^3 + 2\beta^3 + 4\gamma^3 &= 2 \\
      \alpha^2\beta + \beta^2\gamma + \gamma^2\alpha &= 0 \\
      \alpha\beta^2 + \beta\gamma^2 + \gamma\alpha^2 &= 0
    \end{align}
    which turns out to have the only solution $\alpha = \gamma = 0, \beta = 1$. Therefore, the only $\mathbb{Q}$-automorphism of $\mathbb{Q}(\sqrt[3]{2})$ is the idnetity map. 
  \end{example} 

  Now let's look at how an $F$-automorphism acts on the roots of a polynomial. Let $f(x) = x^n + a_{n-1} x^{n-1} + \ldots + a_1 x + a_0 \in F[x]$, $F \subset K$ a field extension, and suppose $f(x)$ has roots $\alpha_1, \ldots, \alpha_m$ lying in $K$ (and perhaps other roots lying in a further extension). It turns out that any $F$-automorphism of $K$ must permute $\alpha_1, \ldots, \alpha_m$, so the roots stay ``within'' the polynomial. 

  \begin{lemma}[$F$-Automorphims Permute Roots]
    \label{thm:f_auto_permute}
    Let $F \subset K$ and let $\{\alpha_1, \ldots, \alpha_m\} \subset K$ be the roots of $f(x) \in F[x]$. Note that the roots are in $K$, and there may be perhaps other roots lying outside of $K$. 
    \begin{enumerate}
      \item For any $\phi \in \Aut(K/F)$, $\phi(\alpha)$ is also a root of $f(x)$ in $K$.\footnote{However it may not be a permutation!}

      \item There exists a well-defined group homomorphism 
      \begin{equation}
        \sigma: \Aut(K/F) \to S_m, 
      \end{equation}
      where $\sigma(\phi)(i)$ is the unique $j$ such that $\phi(r_i) = r_j$. 

      \item If $K$ is the splitting field of $f(x)$, then $\sigma$ is injective and so $\im(\sigma)$ is a subgroup of $S_m$. 
    \end{enumerate}
  \end{lemma}
  \begin{proof}
    We know that $f(\alpha) = 0$. We claim first that $f(\phi(\alpha)) = 0$ and so $\phi(\alpha)$ is also a root. Now assume that $S = \{\alpha_1, \ldots, \alpha_n\}$ is contained in $K$. We know that $\phi(\alpha_j) \in S$. The map $\phi \mapsto \phi(\alpha_j)$ is indeed a group homomorphism $\Aut(K/F) \to \mathrm{Perm}(S)$. If $K$ is the splitting field, then $K = F[S]$, and so if $\phi \in \Aut(K/F)$ fixes all the $\alpha_j$'s, then it must fix all of $K$. 
  \end{proof} 

  \begin{example}[Recomputing $\Aut(\mathbb{Q}(2^{1/3})/\mathbb{Q})$]
    Therefore, we can get a much simpler solution of the Automorphism group of $\mathbb{Q}(\sqrt[3]{2})$ over $\mathbb{Q}$. Since $\xi = \sqrt[3]{2}$ is a root of the irreducible polynomial $x^3 - 2$, any $\phi \in \Aut(\mathbb{Q}[\xi]/\mathbb{Q})$ must carry $\xi$ to some other root of $x^3 - 2$. The other roots are in the complex plane and not in $\mathbb{Q}(\sqrt[3]{2})$, and so $\phi$ must carry $\xi \mapsto \xi$. Thus, $\phi$ must be the identity. 
  \end{example}

  \begin{example}[Computing $\Aut(\mathbb{Q}(1 + \sqrt{2}) / \mathbb{Q})$]
    Let $f(x) = x^2 - 2x - 1 \in \mathbb{Q}[x]$, which is the minimal polynomial of $1 + \sqrt{2} \in \mathbb{R}$. One root of $f(x)$ is $1 + \sqrt{2} \in \mathbb{Q}(\sqrt{2})$. But we know that there exists a $\phi \in \Aut(\mathbb{Q}(\sqrt{2})/\mathbb{Q})$ that conjugates, and so $1 - \sqrt{2}$ must also be a root. 
  \end{example} 

  \begin{example}[Computing $\Aut(\mathbb{C}/\mathbb{R})$]
    It follows that $\Aut(\mathbb{C}/\mathbb{R})$ is a group of order $2$ generated by complex conjugation. 
  \end{example} 

  By studying how $F$-automorphisms behave, we are able to understand a bit more about the automorphism group. Now we attempt to know more about the order of a automorphism group by finding out how many possible $F$-automorphisms over $K$ can exist. If we let $F \subset K$ be a field extension of degree $n >1$ (i.e. $\dim_F K = n > 1$) and $\alpha \in K$, the set of $n+1$ vectors $\{1, \alpha, \alpha^2, \ldots, \alpha^n\}$ must be linearly dependent over a vector space of dimension $n$. Therefore, there exists some nontrivial linear combination 
  \begin{equation}
    0 = a_n \alpha^n + a_{n-1} \alpha^{n-1} + \ldots + a_1 \alpha + a_0
  \end{equation} 
  which corresponds to some nonzero $f(x) = a_n x^n + \ldots a_0$. Therefore $\alpha$ is a root of a polynomial $f(x) \in F[x]$ of degree at most $n$---therefore a root of an irreducible (i.e. minimal) polynomial of degree at most $n$ (since there could be a smaller linearly dependent set). 

  Now if we assume that $K = F[\alpha]$, let's try to see how many elements $\Aut(K/F)$ could have. 
  \begin{enumerate}
    \item $f(x)$ must have degree exactly $n$ (why?). Since $1, \alpha, \ldots, \alpha^{n-1}$ form a basis of $K$ over $F$, if $\phi \in \Aut(K/F)$, the action of $\phi$ on $K = F[\alpha]$ is completely determined by $\phi(\alpha)$, which is what we saw for the first two examples above. 
    \item On the other hand, $\phi$ must map $\alpha$ to some root of $f(x)$ in $K$, and there are at most $n = [K:F]$ of these. 
  \end{enumerate}
  Thus we have $|\Aut(F[\alpha]/F)| \leq [F[\alpha]:F]$, and it follows that if $F[\alpha] \subset K$, then $|\Aut(F[\alpha]/F)| \leq [K:F]$. But our goal is to do so for arbitrary field extensions, i.e. 
  \begin{equation}
    |\Aut(K/F)| \leq [K:F]
  \end{equation} 
  To do this, we will need to talk about extensions of field isomorphisms. Recall that a field (i.e. a ring) isomorphism $\phi: F \to F^\prime$ induces a ring isomorphism $\Tilde{\phi}: F[x] \to F^\prime [x]$. 


  \begin{theorem}[Extending Isomorphisms of Fields]
    Let $\phi: F \to F^\prime$ be an isomorphism of fields.
    \begin{enumerate}
      \item Let $f(x) \in F[x]$ be an irreducible polynomial with root $\alpha$ in some field extension $K \supset F$.
      \item Let $g(x) = \phi(f(x)) \in F^\prime[x]$ and let $\beta$ be a root of $g(x)$ in some field extension $K^\prime \supset F^\prime$. 
    \end{enumerate}
    Then there is a unique isomorphism  
    \begin{equation}
      \bar{\phi} : F[\alpha] \to F^\prime [\alpha^\prime]
    \end{equation}
    that is an extension of $\phi$ (i.e. behaves the same under $F$) and carries $\alpha$ to $\alpha$. 
  \end{theorem}

  We were interested in the automorphism group of the symmetries of the field. Now we will extend this to bigger fields. 

  \begin{theorem}[Bound on Number of Field Isomorphism Extensions]
    Let $F \subset K_1, F \subset K_2$ be field extensions. Then there are at most $[K_1:F]$ $F$-embeddings $\phi: K_1 \to K_2$. Moreoever, if there are exactly $[K_1:F]$ $F$-embeddings, then for all $F \subset E \subset K$, there are $[E:F]$ $F$-embeddings $\phi: E \to K_2$. 
  \end{theorem}
  \begin{proof}
    By induction on $[K_1:F]$. When $[K_1:F] = 1$. Now suppose the theorem holds for all $F, K_1, K_2$ with $[K_1:F] \leq n-1$. Now choose $F, K_1, K_2$, as in the statement with $[K_1 : F] = n > 1$. Since $[K_1: F] > 1$, we can choose $\alpha \in K_1$, $\alpha \not\in F$. Let $f(x) \in F[x]$ be the minimal polynomial of $\alpha$. We've seen that given the smallest subfield of $K_1$ containing $\alpha$, i.e. $F[\alpha]$, we can use the evaluation homomorphism to state 
    \begin{equation}
      F[\alpha] \simeq \frac{F[x]}{\langle f(x)\rangle}, \qquad F[\alpha] \xleftarrow{\mathrm{ev}_\alpha} \frac{F[x]}{\langle f(x)\rangle} 
    \end{equation}
    Any $F$-embedding of $K_1$ restricst to an $F$-embedding of $F[\alpha]$. By induction, there are at most $[F[\alpha]: F] = m$ $F$-embeddings 
    \begin{equation}
      \phi_1, \ldots, \phi_m : F[\alpha] \to K_2
    \end{equation}
    But by induction, we know that there are at most $[K_1: F[\alpha]]$ $F[\alpha]$-embeddings $K_1 \to K_2$ where $F[\alpha] \subset K_2$ (which we can defined through multiple ways for each injection $\phi_i$). This implies that there are at most 
    \begin{align}
      m \, [K_1: F[\alpha]] & = [F[\alpha]: F]\, [K_1: F[\alpha] \\
                            & = [K_1 : F]
    \end{align}
    $F$-embeddings. Suppose equality holds. Let $F \subset E \subset K_1$. We know 
    \begin{enumerate}
      \item There are at most $[E:F]$ $F$-embeddings $\phi_1, \ldots, \phi_{[E:F]}: E \to K_2$ 
      \item There are at most $[K_1:E]$ $E$-embeddings $K_1 \to K_2$ where here $E \subset K_2$. 
    \end{enumerate}
    This gives at most $[K_2:E] [E:F]$ $F$-embeddings since $[K_2 :E] [E:F] = [K_2 :F]$, and there are exactly $[K_2 : F]$ $F$-embeddings $K_1 \to K_2$, all the $\leq$'s must be $=$'s (otherwise it would be a strict inequality) in that there must be exactly $[E:F]$ $F$-embeddings $E \to K_2$ and exactly $[K_1: E]$ $E$-embeddings $K_1 \to K_2$. 

    Correction. We need to see that there are at most $[F[\alpha]:F]$ $F$-embeddings $F[\alpha] \to K_2$. It's okay to use induction if $[F[\alpha]: F] \subset [K_1:F]$. In general, we showed $F$-embeddings maps 
  \end{proof}

  Now combined with adjoining fields, we get a nice bound on the number of elements in an $F$-automorphism group. 

  \begin{corollary}
    Given field extension $F \subset F[\alpha]$ with $\alpha$ is algebraic over $F$ and minimal polynoimal $f(x) \in F[x]$, there exists a bijection between $\Aut(F(\alpha)/F)$ and the roots $r$ of $f(x)$. It follows that 
    \begin{equation}
      |\Aut(F(\alpha)/F)| \leq [F(\alpha):F]
    \end{equation}
  \end{corollary}
  \begin{proof}
    Therefore, given field extensions $F \subset F(\alpha) \subset K$, by the lemma on bijections, we know that there is a bijection between $\Aut(F(\alpha)/F)$ and the roots $r \in K$ of some irreducible polynomial $f(x) \in F[x]$. But the latter can be at most the degree of $f(x)$, which is precisely 
    \begin{equation}
      [F(\alpha):F] = \dim_F \frac{F[\alpha][x]}{\langle f(x) \rangle}
    \end{equation}
  \end{proof}

\subsection{Galois Extensions and Galois Groups}

  If we reach this bound, then there are some special properties. 

  \begin{definition}[Galois Field Extension]
    A field extension $F \subset K$ is \textbf{Galois} if $|\Aut(K/F)| = [K:F]$---or equivalently, the degree of the minimal polynomial $f(x) \in F[x]$ that splits in $K$. When this is the case, the automorphism group is called the \textbf{Galois Group}, denoted $\mathcal{G}(K/F)$. 
  \end{definition}

  \begin{example}[Simple Examples]
    We review the derived Galois groups above and see if the field extensions are Galois. 
    \begin{enumerate}
      \item $\mathbb{Q}(\sqrt{2})$ is a Galois extension of $\mathbb{Q}$. Since we know that $G = \mathcal{G}(\mathbb{Q}(\sqrt{2})/\mathbb{Q})$ has order $2$, which is the same as the dimension of $\mathbb{Q}(\sqrt{2})$, i.e. the degree of the minimal polynomial $x^2 - 2 \in \mathbb{Q}[x]$. 

      \item $\mathbb{Q}(\sqrt[3]{2})$ is not a Galois extension of $\mathbb{Q}$ since $|\mathcal{G}(\mathbb{Q}[\sqrt[3]{2}]/\mathbb{Q})| = 1$ but $[\mathbb{Q}[\sqrt[3]{2}] : \mathbb{Q}] = 3$. 

      \item Let $K = \mathbb{Q}[\sqrt[3]{2}, i \sqrt{3}]$ is the splitting field of $f(x) = x^3 - 2$. Then $[K:\mathbb{Q}] = 6$. $\mathcal{G}(K/\mathbb{Q}) \simeq S_3$ has order $6$ since we showed 6 $\mathbb{Q}$-automorphisms of $K$, but now we know that there can be no more. 

      \item Let $\alpha = \sqrt[7]{2}$. $\mathbb{Q} \subset \mathbb{Q}[\alpha]$ is not a Galois extension. $x^7 - 2$ is a polynomial with root $\alpha$, and by Eisenstein $x^7 - 2$ is irreducible. So $f(x) = x^7 - 2$ is the minimal polynomial of $\alpha$. This means that the number of $\mathbb{Q}$-embeddings $\mathbb{Q}[\alpha] \to \mathbb{Q}[\alpha]$ is in bijection with the number of roots of $f(x)$ in $\mathbb{Q}[\alpha]$. All $7$ roots of $f(x)$ are $\sqrt[7]{2} e^{2\pi i j/7}$ for $j = 0, \ldots, 6$, which has one real root. So there is $1$ $\mathbb{Q}$-embedding $\mathbb{Q}[\alpha] \to \mathbb{Q}[\alpha]$. But $[\mathbb{Q}[\sqrt[7]{2}] : \mathbb{Q}] = \deg{f(x)} = 7$. Since $\mathbb{Q}[\alpha] \simeq \mathbb{Q}[x]/{\langle x^7 - 2 \rangle}$, which are polynomials of degree at most $6$. 
    \end{enumerate}
  \end{example}

  These examples suggest that $K$ will be a Galois extension of $F$ whenever $K$ is a splitting field of some polynomial $f(x) \in F[x]$. To establish this, we must produce $[K:F]$ $F$-automorphisms of $K$ under these circumstances. 

  \begin{theorem}[Splitting Fields of a Square Free Polynomial are Galois]
    \label{thm:6.6}
    Suppose $F \subset K$ is the splitting field of a polynomial $f(x) \in F[x]$ such that no irreducible factor of $f(x)$ has repeated roots. Then $F \subset K$ is Galois. 
  \end{theorem}
  \begin{proof}
    Since $K$ is the splitting field of $f(x)$ over $F$, we have $F[r_1, \ldots, r_m]$ where the $r_i$ are the distinct roots of $f(x)$. We show by induction that there are $[F[r_1, \ldots, r_j]:F]$ $F$-embeddings of $F[r_1, \ldots, r_j] \to K$. For $j = 1$, $r_1$ is a root of some irreducible factor $f_1 (x)$ of $f(x)$. 
    \begin{equation}
      F[r_1] \simeq \frac{F[x]}{\langle f_1 (x) \rangle}
    \end{equation} 
    and the set of $F$-embeddings $F[r_1] \to K$ is in bijection with the set of roots $\alpha \in K$ of $f_1 (x)$. By hypothesis , $f(x)$ has no repeated roots, which implies that the number of $F$-embeddings $F[r_1] \to K$ is $\deg{f_1 (x)} = [F[r_1]: F]$ which gives the base case. For the inductive step, we know there are 
    \begin{equation}
      [F[r_1, \ldots, r_{j-1}]: F] 
    \end{equation}
    $F$-embeddings $F[r_1, \ldots, r_{j-1}] \xrightarrow{\phi} K$. For each, we will show that there are exactly $[F[r_1, \ldots, r_j]: F[r_1, \ldots, r_{j-1}]]$ extensions of $\phi$ which completes the proof. Because 
    \begin{align}
      F[r_1, \ldots, r_j]: F] & = F[r_1, \ldots, r_j]: F[r_1, \ldots, r_{j-1}]] F[r_1, \ldots, r_{j-1}]: F]
    \end{align} 
    Let $g(x)$ be the minimal polynomial of $r_j$ over $F[r_1, \ldots, r_{j-1}]$. Since $g(r_j) = 0$, $g$ divides one of the irreducible factors of $f(x)$ in $F[r_1, \ldots, r_{j-1}] [x]$ which implies it has no repeated roots. Then 
    \begin{equation}
      F[r_1, \ldots, r_j] = \frac{E[x]}{\langle g(x)\rangle} 
    \end{equation}
    The number of $E$-embeddings is equal to the number of roots in in $g$ which is $\deg{g} = F[r_1, \ldots, r_j: F]$.
  \end{proof} 

  Now we restrict our scope to $\mathbb{Q}$. Note the following. 

  \begin{theorem}[Splitting Fields of $\mathbb{Q}$ are Automatically Galois]
    % Shifrin 6.8 
    \label{thm:q_square_free}
    \label{thm:6.8}
    Any irreducible polynomial in $\mathbb{Q}$ has no repeated roots. Therefore, let $\mathbb{Q} \subset K$ be the splitting field of $f(x) \in Q[x]$. Then $\mathbb{Q} \subset K$ is Galois. 
  \end{theorem} 
  \begin{proof}
    Let $f(x) \in \mathbb{Q}[x]$ be irreducible and $\alpha \in \mathbb{C}$ be its root with multiplicity $k > 1$. Then 
    \begin{equation}
      f(x) = (x - \alpha)^k g(x)
    \end{equation}
    for some $g(x)$. Taking the derivative we have 
    \begin{equation}
      f^\prime(x) = k (x - \alpha)^{k-1} g(x) + (x - \alpha)^k g^\prime (x) = (x - \alpha)^{k-1} \big( k g(x) + (x - \alpha) g^\prime(x) \big)
    \end{equation}
    which also has $\alpha$ as a root. So since $(x - \alpha)$ divides both $f$ and $f^\prime$, it divides $\gcd(f, f^\prime)$. 

    From \ref{thm:q_square_free}, all irreducible factors of $f(x)$ has no repeated roots. Therefore by \ref{thm:6.6} $K$, which is the splitting field of $f(x)$, is Galois. 
  \end{proof}

  What about the converse? There are two steps to proving that for any Galois field extension $F \subset K$, there exists a polynomial $f(x) \in F[x]$ that splits in $K$. 

  \begin{lemma}[Irreducible Polynomial with Root in Galois Extension Splits] 
    % Shifrin 6.9
    \label{thm:6.9}
    Let $F \subset K$ be a Galois extension of fields. Let $f(x) \in F[x]$ be an irreducible polynomial with a root $\alpha \in K$. Then $f(x)$ splits in $K$, i.e. all other roots must be in $K$.  
  \end{lemma} 
  \begin{proof}
    Let $\sigma_j$ for $j = 1, \ldots, n$ be the elements of $\mathcal{G}(K/F)$, and set $\alpha_j = \sigma_j (\alpha)$. Then define 
    \begin{equation}
      h(x) = \prod_{i=1}^n (x - \alpha_i)  \in K[x]
    \end{equation}
    We claim that the coefficients of $h(x)$ are fixed by any element of $\mathcal{G}(K/F)$. This is because by \ref{thm:f_auto_permute}, the coefficients obtained by taking the binomial expansion stay invariant. Therefore, all coefficients of $h(x)$ are in $F$, and so $h(x) \in F[x]$. Since $g(x)$ and $h(x)$ have a common root $\alpha \in K$, and since $g(x)$ is irreducible in $F[x]$, we have $g(x) \mid h(x)$. Therefore $g(x)$ as a factor of $h(x)$ which splits in $K$, also splits in $K[x]$. 
  \end{proof}

  \begin{theorem}[Every Galois Extension is a Splitting Field]
    Let $F \subset K$ be a Galois extension. Then $K$ is the splitting field of a polynomial $f(x) \in F[x]$. 
  \end{theorem}
  \begin{proof}
    Let $\alpha_1, \ldots, \alpha_k$ be a basis for $K$ over $F$, and for each $j = 1, \ldots, k$, let $g_j (x) \in F[x]$ be an irreducible polynomial with root $\alpha_j$. Then by \ref{thm:6.9}, $K$ is the splitting field $g_j(x)$ splits in $K$, which implies that 
    \begin{equation}
      f(x) = g_1 (x) \ldots g_k(x) \in F[x]
    \end{equation}
    also splits in $K$, and it does in no smaller fields since any splitting field must contain all the $\alpha_j$'s. 
  \end{proof}

  Now let's do some practice. In general, 
  \begin{enumerate}
    \item If we are given a field extension, we must prove that it is a Galois extension, then find the order of its Galois group, then compute the elements using the theorems we have above. 
    \item If we are given a polynomial $f(x) \in F[x]$, when we must find that no irreducible factor of it has repeated roots, and so we conclude that $F \subset K$ is Galois. Then we just do the steps in (1). 
  \end{enumerate}

  \begin{example}[Computing Galois Groups]
    Now let's compute the Galois group of $\mathbb{Q}(\sqrt{2}, \sqrt{3})$ over $\mathbb{Q}$. 
    \begin{enumerate}
      \item $\mathbb{Q}(\sqrt{2}, \sqrt{3})$ is the splitting field of $f(x) = (x^2 - 2)(x^2 - 3)$ over $\mathbb{Q}$. Then by \ref{thm:6.8} the field extension $\mathbb{Q} \subset \mathbb{Q}(\sqrt{2}, \sqrt{3})$ is Galois. 

      \item Therefore the Galois group of this extension exists. Since $\sqrt{2}, \sqrt{3}$ are not linearly dependent, we have $[\mathbb{Q}(\sqrt{2}, \sqrt{3}):\mathbb{Q}] = 4$. Thus the Galois group has order $4$. Now it remains to compute the 4 elements. 

      \item Since $\mathbb{Q}(\sqrt{2}, \sqrt{3})$ is a $\mathbb{Q}$-vector space of dimension 4, we know that every element is of the form $a + b \sqrt{2} + c \sqrt{3} + d \sqrt{6}$. Since 
      \begin{equation}
        \phi(a + b \sqrt{2} + c \sqrt{3} + d \sqrt{6}) = a + b \phi(\sqrt{2}) + c \phi(\sqrt{3}) + d \phi(\sqrt{2}) \phi(\sqrt{3}), 
      \end{equation}
      every $\mathbb{Q}$-automorphism is determined by the values $\phi(2), \phi(3)$. 

    \item From \ref{thm:f_auto_permute}, we know that elements $\phi \in \mathcal{G}(\mathbb{Q}(\sqrt{2}, \sqrt{3}) / \mathbb{Q})$ permute the roots of $f(x)$. We immediately see that if $\phi(\sqrt{2}) = \sqrt{3}$
      \begin{equation}
        2 = \phi(2) = \phi(\sqrt{2})^2 = \sqrt{3}^2 = 3
      \end{equation}
      which is a contradiction. The same logic follows for $\sqrt{2} \mapsto \pm \sqrt{3}, \sqrt{3} \mapsto \pm \sqrt{2}$. So it must be the case that $\phi(\sqrt{2}) = \pm \sqrt{2}$ and $\phi(\sqrt{3}) = \pm \sqrt{3}$. 
    \end{enumerate}
    Therefore, we are able to deduce that the 4 automorphisms are. 
    \begin{align}
      \phi_1 (a + b \sqrt{2} + c \sqrt{3} + d \sqrt{6}) & = a + b \sqrt{2} + c \sqrt{3} + d \sqrt{6} \\
      \phi_1 (a + b \sqrt{2} + c \sqrt{3} + d \sqrt{6}) & = a - b \sqrt{2} + c \sqrt{3} - d \sqrt{6} \\
      \phi_1 (a + b \sqrt{2} + c \sqrt{3} + d \sqrt{6}) & = a + b \sqrt{2} - c \sqrt{3} - d \sqrt{6} \\
      \phi_1 (a + b \sqrt{2} + c \sqrt{3} + d \sqrt{6}) & = a - b \sqrt{2} - c \sqrt{3} + d \sqrt{6}
    \end{align}
    which is isormophic to the Klein 4-group. 
  \end{example}

  \begin{example}[Computing Galois Groups of 3 Adjoining Elements]
    
  \end{example}

\subsection{Fundamental Theorem of Galois Theory}

  Now we classify Galois groups by creating a bijection between the intermediate fields and the subgroups of a Galois group. 

  \begin{theorem}[Fundamental Theorem of Galois Theory]
    Let $K$ be a Galois extension of $\mathbb{Q}$, and let $G = \mathcal{G}(K/\mathbb{Q})$. Then there is a bijection between the two sets
    \begin{enumerate}
      \item The intermediate fields $\mathbb{Q} \subset E \subset K$
      \item The set of subgroups $H \subset G$. 
    \end{enumerate} 
    Also, for each intermediate field, we want to find out whether $\mathbb{Q} \subset E$ and/or $E \subset K$ is a Galois extension, hence whether $\mathcal{G}(E/\mathbb{Q})$ or $\mathcal{G}(K/E)$ is Galois. 
    \begin{enumerate}
      \item $K$ is always a Galois extension of $E$ and $\mathcal{G}(K/E) = H$. It is immediate that $[K:E] = |H|$. 
      \item $E$ is a Galois extension of $\mathbb{Q}$ iff $H$ is a normal subgroup of $G$, in which case $\mathcal{G}(E/\mathbb{Q}) \simeq G/H$ and so $[E:\mathbb{Q}] = [G:H]$.\footnote{Note the left is the degree of $E$ over $\mathbb{Q}$ while the right is the number of cosets of $H$ in $G$!} 
    \end{enumerate}
  \end{theorem}
  \begin{proof}
    
  \end{proof}

  Note that in the extreme case when $E = K$, $[K:E] = 1$ and so the Galois group can have order $1$, which must mean that $G = \{e\}$. Furthermore, If $E = \mathbb{Q}$, we have the full group $G = \mathcal{G}(K/\mathbb{Q})$. 

  \begin{corollary}
    $E$ is a Galois extension of $\mathbb{Q}$ if and only if $H$ is a normal subgroup of $G$, in which case $\mathcal{G}(E/\mathbb{Q}) \simeq G/H$. 
  \end{corollary}

\subsection{Cubic Equations}

  The well known discriminant of a quadratic equation 
  \begin{equation}
    f(x) = ax^2 + bx + c
  \end{equation}
  is known in the form $\nabla = b^2 - 4ac$. However, we will present it in a slightly different manner. 

  \begin{definition}
    The \textbf{discriminant} $D(\varphi)$ of a quadratic polynomial
    \begin{equation}
      \varphi = a_0 x^2 + a_1 x + a_2 \in \mathbb{C}[x]
    \end{equation}
    with $c_1, c_2 \in \mathbb{C}$ as its roots is defined
    \begin{equation}
      D(\varphi) = a_1^2 - 4 a_0 a_2 = a_0^2 \bigg( \Big(\frac{a_1}{a_0} \Big)^2 - \frac{4 a_2}{a_0} \bigg) = a_0^2 \big( (c_1 + c_2)^2 - 4 c_1 c_2 \big) = a_0^2 (c_1 - c_2)^2
    \end{equation}
    Clearly, the value of $D(\varphi)$ can tell us three things
    \begin{enumerate}
      \item $c_1, c_2 \in \mathbb{R}, c_1 \neq c_2$. Then $c_1 - c_2$ is a nonzero real number and $D(\varphi) > 0$. 
      \item $c_1 = c_2 \in \mathbb{R}$. Then $c_1 - c_2 = 0$ and $D(\varphi) = 0$. 
      \item $c_1, c_2 \in \mathbb{C}, c_1 = \bar{c}_2$. Then, $c_1 - c_2$ is a nonzero strictly imaginary number and $D(\varphi) < 0$. 
    \end{enumerate}
  \end{definition}

  \begin{definition}
    We can generalize this notion of the discriminant to arbitrary polynomials
    \begin{equation}
      \varphi = a_0 x^n + a_1 x^{n-1} + ... + a_{n-1} x + a_n \in \mathbb{F}[x], \; a_0 \neq 0
    \end{equation}
    The discriminant $D(\varphi)$ of the polynomial above is defined
    \begin{equation}
      D(\varphi) \equiv a_0^{2n-2} \prod_{i>j} (c_i - c_j)^2
    \end{equation}
    The $a_0$ term isn't very important in this formula, since it does not affect whether $D(\varphi)$ is positive, negative, or zero. 
  \end{definition}

  \begin{definition}
    A polynomial 
    \begin{equation}
      \varphi = a_0 x^n + a_1 x^{n-1} + ... + a_{n-1} x + a_n \in \mathbb{F}[x], \; a_0 \neq 0
    \end{equation}
    where $a_1 = 0$ is called \textbf{depressed}. A depressed cubic polynomial is of form
    \begin{equation}
      \varphi = x^3 + p x + q
    \end{equation}
  \end{definition}

  \begin{theorem}
    Every monic (leading coefficeint $=1$) polynomial (and non-monic ones) 
    \begin{equation}
      \varphi = x^n + a_1 x^{n-1} + ... + a_{n-1} x + a_n \in \mathbb{F}[x], \; a_0 \neq 0
    \end{equation}
    can be turned into a depressed polynomial with the change of variable
    \begin{equation}
      x = y - \frac{a_1}{n}
    \end{equation}
    to get the polynomial 
    \begin{equation}
      \psi = y^n + b_2 y^{n-2} + ... + b_{n-1} y + b_n
    \end{equation}
  \end{theorem}

  \begin{lemma}
    A cubic polynomial 
    \begin{equation}
      \varphi = a_0 x^3 + a_1 x^2 + a_2 x + a_3 \in \mathbb{R}[x]
    \end{equation}
    with roots $c_1, c_2, c_3 \in \mathbb{C}$ has discriminant
    \begin{equation}
      D(\varphi) \equiv a_0^4 (c_1 - c_2)^2 (c_1 - c_3)^2 (c_2 - c_3)^2
    \end{equation}
    With a bit of evaluation, it can also be expressed in terms of its coefficients as
    \begin{equation}
      D(\varphi) = a_1^2 a_2^2 - 4a_1^3 a_3 - 4a_0 a_2^3 + 18 a_0 a_1 a_2 a_3 - 27 a_0^2 a_3^2
    \end{equation}
    Again, three possibilities can occur (up to reordering of its roots). 
    \begin{enumerate}
        \item $c_1, c_2, c_3$ are distinct real numbers. Then $D(\varphi) > 0$. 
        \item $c_1, c_2, c_3 \in \mathbb{R}, c_1 = c_2$. Then $D(\varphi) = 0$. 
        \item $c_1 \in \mathbb{R}, c_2 = \bar{c}_3 \not\in \mathbb{R}$. Then $D(\varphi) < 0$. 
    \end{enumerate}
    Furthermore, the cubic formula used to find the roots of the polynomial is 
    \begin{equation}
      c_{1, 2, 3} = \sqrt[3]{-\frac{q}{2} + \sqrt{\frac{p^3}{27} + \frac{q^2}{4}}} + \sqrt[3]{-\frac{q}{2} - \sqrt{\frac{p^3}{27} + \frac{q^2}{4}}}
    \end{equation}
    known as \textbf{Cardano's formula}, after the mathematician Gerolamo Cardano. 
  \end{lemma}




